---
title: "EU Paediatric Clinical Trials"
subtitle: "Summary Report"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    latex_engine: pdflatex
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 6.5
    fig_height: 3.4
    keep_tex: false
classoption: a4paper
header-includes:
  - \usepackage{helvet}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{float}
  - \usepackage{xcolor}
  - \usepackage{geometry}
  - \geometry{top=22mm,bottom=22mm,left=22mm,right=22mm}
  - \definecolor{nordblue}{HTML}{5E81AC}
  - \definecolor{nordgray}{HTML}{4C566A}
  - \setlength{\parskip}{5pt}
  - \setlength{\parindent}{0pt}
params:
  data_path: ""
  filters: !r list()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(stringr)
  library(lubridate)
  library(scales)
  library(knitr)
})

# ── Data ───────────────────────────────────────────────────────────────────────
df <- readRDS(params$data_path)
N  <- nrow(df)
f  <- params$filters

if (N == 0) {
  cat("\\textbf{No trials match the current filter criteria.}")
  knitr::knit_exit()
}

# ── Colour palette ─────────────────────────────────────────────────────────────
BLUE   <- "#5E81AC"; FROST  <- "#81A1C1"; TEAL   <- "#88C0D0"
GREEN  <- "#A3BE8C"; YELLOW <- "#EBCB8B"; ORANGE <- "#D08770"
RED    <- "#BF616A"; PURPLE <- "#B48EAD"; GRAY   <- "#4C566A"
LIGHT  <- "#ECEFF4"; DARK   <- "#2E3440"
NORD8  <- c(BLUE, GREEN, YELLOW, RED, PURPLE, ORANGE, TEAL, FROST)

STATUS_COLS <- c(Ongoing = GREEN, Completed = BLUE, Other = ORANGE)
REG_COLS    <- c(EUCTR = BLUE, CTIS = TEAL)
PIP_COLS    <- c(Yes = GREEN, No = RED, Unknown = ORANGE)

# ── ggplot2 theme ──────────────────────────────────────────────────────────────
theme_rep <- function(base = 9) {
  theme_minimal(base_size = base, base_family = "Helvetica") %+replace% theme(
    plot.title       = element_text(size = base + 1, face = "bold",  color = DARK,  hjust = 0),
    plot.subtitle    = element_text(size = base - 0.5, color = GRAY, hjust = 0),
    axis.title       = element_text(size = base - 0.5, color = GRAY),
    axis.text        = element_text(size = base - 1,   color = GRAY),
    axis.text.y      = element_text(margin = margin(r = 2)),
    panel.grid.major = element_line(color = LIGHT, linewidth = 0.45),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    legend.key.size  = unit(0.32, "cm"),
    legend.text      = element_text(size = base - 1.5, color = GRAY),
    plot.margin      = margin(4, 6, 4, 4),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.text       = element_text(size = base - 0.5, face = "bold", color = DARK)
  )
}

# ── kable helper ───────────────────────────────────────────────────────────────
nk <- function(x, cap = NULL, cn = colnames(x), al = NULL, lw = FALSE) {
  kable(x, booktabs = TRUE, linesep = "", caption = cap,
        col.names = cn, align = al, longtable = lw, escape = TRUE)
}

fmt_pct  <- function(x) sprintf("%.1f%%", x)
fmt_n    <- function(x) format(x, big.mark = ",")
iqr_str  <- function(x) {
  q <- quantile(x, c(0.25, 0.75), na.rm = TRUE)
  sprintf("%.1f\u2013%.1f", q[1], q[2])
}

# ── Pre-compute tables ─────────────────────────────────────────────────────────
status_tab <- df %>%
  mutate(status = coalesce(status, "Unknown")) %>%
  count(status, name = "n") %>%
  mutate(pct = round(n / N * 100, 1)) %>%
  arrange(desc(n))

reg_tab <- df %>%
  count(register, name = "n") %>%
  mutate(pct = round(n / N * 100, 1)) %>%
  arrange(desc(n))

pip_tab <- df %>%
  mutate(has_PIP = coalesce(has_PIP, "Unknown")) %>%
  count(has_PIP, name = "n") %>%
  mutate(pct = round(n / N * 100, 1)) %>%
  arrange(match(has_PIP, c("Yes", "No", "Unknown")))

sponsor_tab <- df %>%
  filter(!is.na(sponsor_type)) %>%
  count(sponsor_type, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(n))

phase_tab <- df %>%
  filter(!is.na(phase), phase != "") %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase != "") %>%
  count(phase, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  arrange(match(phase, c("I", "I-II", "II", "II-III", "III", "III-IV", "IV")))

country_tab <- df %>%
  filter(!is.na(Member_state)) %>%
  separate_rows(Member_state, sep = " / ") %>%
  mutate(Member_state = str_trim(Member_state)) %>%
  filter(Member_state != "") %>%
  count(Member_state, name = "n", sort = TRUE) %>%
  mutate(pct = round(n / N * 100, 1))

organ_tab <- df %>%
  filter(!is.na(MEDDRA_organ_class)) %>%
  separate_rows(MEDDRA_organ_class, sep = " / ") %>%
  mutate(MEDDRA_organ_class = str_trim(MEDDRA_organ_class)) %>%
  filter(MEDDRA_organ_class != "") %>%
  count(MEDDRA_organ_class, name = "n", sort = TRUE) %>%
  mutate(pct = round(n / N * 100, 1))

cond_tab <- df %>%
  filter(!is.na(MEDDRA_term)) %>%
  separate_rows(MEDDRA_term, sep = " / ") %>%
  mutate(MEDDRA_term = str_trim(MEDDRA_term)) %>%
  filter(MEDDRA_term != "") %>%
  count(MEDDRA_term, name = "n", sort = TRUE) %>%
  mutate(pct = round(n / N * 100, 1))

year_reg <- df %>%
  filter(!is.na(submission_date_parsed)) %>%
  mutate(year = year(submission_date_parsed)) %>%
  count(year, register, name = "n") %>%
  arrange(year)

year_total <- year_reg %>%
  group_by(year) %>% summarise(n = sum(n), .groups = "drop") %>%
  arrange(year) %>% mutate(cumulative = cumsum(n))

# Continuous per-trial counts
countries_per_trial <- df %>%
  filter(!is.na(Member_state)) %>%
  rowwise() %>%
  mutate(nc = length(str_trim(str_split(Member_state, " / ")[[1]]))) %>%
  pull(nc)

conds_per_trial <- df %>%
  filter(!is.na(MEDDRA_term)) %>%
  rowwise() %>%
  mutate(nc = length(str_trim(str_split(MEDDRA_term, " / ")[[1]]))) %>%
  pull(nc)

# Top sponsors by name
top_sponsor_tab <- df %>%
  filter(!is.na(sponsor_name)) %>%
  count(sponsor_name, sponsor_type, sort = TRUE) %>%
  slice_head(n = 20) %>%
  mutate(pct = round(n / N * 100, 1))

# Sponsor × register cross-tab
sponsor_reg <- df %>%
  filter(!is.na(sponsor_type)) %>%
  count(sponsor_type, register) %>%
  pivot_wider(names_from = register, values_from = n, values_fill = 0) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>%
  arrange(desc(Total))

# Phase × register
phase_reg <- df %>%
  filter(!is.na(phase), phase != "") %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase != "") %>%
  count(phase, register) %>%
  pivot_wider(names_from = register, values_from = n, values_fill = 0) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>%
  arrange(match(phase, c("I","I-II","II","II-III","III","III-IV","IV")))

# Decision timeline
decision_df <- df %>%
  filter(!is.na(days_to_decision), is.finite(days_to_decision),
         days_to_decision >= 0, days_to_decision < 3650)

```

# Applied Filters {-}

The following filters were active when this report was generated.

```{r filter-table}
fnames <- c("Trial Status", "Source Register", "Submission Date (from)",
            "Submission Date (to)", "MedDRA Organ Class", "Condition / MedDRA Term",
            "Country / Member State", "Trial Phase", "Part of PIP", "Free-text search")
fvals <- c(
  if (length(f$status) == 0)   "All" else paste(f$status, collapse = ", "),
  if (length(f$register) == 0) "All" else paste(f$register, collapse = ", "),
  if (length(f$date_range) >= 1) f$date_range[1] else "—",
  if (length(f$date_range) >= 2) f$date_range[2] else "—",
  if (length(f$organ_class) == 0 || identical(f$organ_class, "All")) "All"
    else paste(f$organ_class, collapse = "; "),
  if (length(f$condition) == 0 || identical(f$condition, "All")) "All"
    else paste(f$condition, collapse = "; "),
  if (length(f$country) == 0 || identical(f$country, "All")) "All"
    else paste(f$country, collapse = "; "),
  if (length(f$phase) == 0 || identical(f$phase, "All")) "All"
    else paste(f$phase, collapse = ", "),
  if (is.null(f$pip) || f$pip == "All") "All" else f$pip,
  if (is.null(f$text_search) || f$text_search == "(none)" || !nzchar(f$text_search)) "(none)"
    else f$text_search
)
nk(data.frame(Filter = fnames, Value = fvals),
   cn = c("Filter", "Value"), al = c("l","l"))
```

\newpage

# Overview

```{r overview-summary}
date_range_str <- if (!all(is.na(df$submission_date_parsed))) {
  paste(format(min(df$submission_date_parsed, na.rm=TRUE), "%d %b %Y"), "to",
        format(max(df$submission_date_parsed, na.rm=TRUE), "%d %b %Y"))
} else "n/a"

n_countries_unique <- country_tab %>% nrow()
n_organs_unique    <- organ_tab %>% nrow()
n_conds_unique     <- cond_tab %>% nrow()

overview_df <- tibble(
  Metric = c("Total trials in selection", "Submission date range",
             "Unique member states", "Unique MedDRA organ classes",
             "Unique MedDRA conditions"),
  Value = c(fmt_n(N), date_range_str,
            as.character(n_countries_unique),
            as.character(n_organs_unique),
            fmt_n(n_conds_unique))
)
nk(overview_df, cap = paste("Key metrics for the", fmt_n(N), "selected trials."),
   cn = c("Metric", "Value"), al = c("l","l"))
```

## Trial Status and Register Distribution

```{r status-register, fig.show='hold', out.width='49%', fig.height=3.2, fig.cap="Left: trial status distribution. Right: source register distribution."}
# Status chart
p1 <- status_tab %>%
  mutate(status = factor(status, levels = status),
         label  = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")) %>%
  ggplot(aes(x = reorder(status, n), y = n, fill = status)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = label), hjust = -0.1, size = 2.8, color = GRAY) +
  scale_fill_manual(values = STATUS_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.22)), labels = comma) +
  coord_flip() +
  labs(title = "Trial Status", x = NULL, y = "Number of trials") +
  theme_rep()
print(p1)

# Register chart
p2 <- reg_tab %>%
  mutate(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")) %>%
  ggplot(aes(x = reorder(register, n), y = n, fill = register)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = label), hjust = -0.1, size = 2.8, color = GRAY) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.22)), labels = comma) +
  coord_flip() +
  labs(title = "Source Register", x = NULL, y = "Number of trials") +
  theme_rep()
print(p2)
```

```{r status-register-tables, fig.show='hold', out.width='49%'}
st <- status_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(st, cn = c("Status", "n", "%"), al = "lrr")

rt <- reg_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(rt, cn = c("Register", "n", "%"), al = "lrr")
```

## Per-trial Continuous Metrics

```{r continuous-table}
cont_df <- tibble(
  Metric   = c("Member states per trial", "Conditions (MedDRA terms) per trial",
               "Submissions per year"),
  Mean_SD  = c(
    sprintf("%.1f \u00b1 %.1f", mean(countries_per_trial), sd(countries_per_trial)),
    sprintf("%.1f \u00b1 %.1f", mean(conds_per_trial),    sd(conds_per_trial)),
    sprintf("%.1f \u00b1 %.1f", mean(year_total$n),       sd(year_total$n))
  ),
  Median   = c(
    sprintf("%.1f", median(countries_per_trial)),
    sprintf("%.1f", median(conds_per_trial)),
    sprintf("%.1f", median(year_total$n))
  ),
  IQR      = c(
    iqr_str(countries_per_trial),
    iqr_str(conds_per_trial),
    iqr_str(year_total$n)
  ),
  Min_Max  = c(
    sprintf("%d\u2013%d", min(countries_per_trial), max(countries_per_trial)),
    sprintf("%d\u2013%d", min(conds_per_trial),    max(conds_per_trial)),
    sprintf("%d\u2013%d", min(year_total$n),        max(year_total$n))
  )
)
nk(cont_df,
   cap  = "Descriptive statistics for continuous per-trial and per-year metrics.",
   cn   = c("Metric", "Mean \u00b1 SD", "Median", "IQR", "Min\u2013Max"),
   al   = c("l","r","r","r","r"))
```

\newpage

# Submission Timeline

```{r timeline-bar, fig.height=3.4, fig.cap="Annual trial submissions by source register."}
year_reg %>%
  ggplot(aes(x = year, y = n, fill = register)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  scale_x_continuous(breaks = pretty_breaks(8)) +
  labs(title = "Submissions per Year", x = NULL, y = "Number of trials") +
  theme_rep()
```

```{r timeline-cumul, fig.height=3.0, fig.cap="Cumulative trial submissions over time."}
year_total %>%
  ggplot(aes(x = year, y = cumulative)) +
  geom_area(fill = BLUE, alpha = 0.18) +
  geom_line(color = BLUE, linewidth = 0.9) +
  geom_point(color = BLUE, size = 1.8) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  scale_x_continuous(breaks = pretty_breaks(8)) +
  labs(title = "Cumulative Submissions", x = NULL, y = "Cumulative trials") +
  theme_rep()
```

```{r timeline-table}
ytbl <- year_total %>%
  mutate(n = fmt_n(n), cumulative = fmt_n(cumulative)) %>%
  rename(Year = year, `Submissions` = n, `Cumulative` = cumulative)
nk(ytbl, cap = "Submissions and cumulative total by year.", al = c("l","r","r"))
```

\newpage

# Geographic Distribution

```{r countries-bar, fig.height=5.5, fig.cap="Top 20 EU member states by number of trials (trials may span multiple countries)."}
country_tab %>%
  head(20) %>%
  mutate(Member_state = factor(Member_state, levels = rev(Member_state))) %>%
  ggplot(aes(x = n, y = Member_state)) +
  geom_col(fill = FROST, width = 0.7) +
  geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
            hjust = -0.08, size = 2.6, color = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2)), labels = comma) +
  labs(title = "Trials by Member State (top 20)", x = "Number of trials", y = NULL) +
  theme_rep()
```

```{r countries-table}
ctbl <- country_tab %>%
  head(20) %>%
  mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(ctbl, cap = "Top 20 member states: number of trial participations and percentage of all selected trials.",
   cn = c("Member State", "n", "% of trials"), al = c("l","r","r"))
```

\newpage

# Therapeutic Areas

## MedDRA Organ Classes

```{r organ-bar, fig.height=4.8, fig.cap="Trials per MedDRA system organ class (top 15). Trials may belong to more than one class."}
organ_tab %>%
  head(15) %>%
  mutate(
    lbl  = str_wrap(MEDDRA_organ_class, 35),
    lbl  = factor(lbl, levels = rev(lbl))
  ) %>%
  ggplot(aes(x = n, y = lbl)) +
  geom_col(fill = PURPLE, width = 0.72) +
  geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
            hjust = -0.08, size = 2.5, color = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25)), labels = comma) +
  labs(title = "MedDRA Organ Classes (top 15)", x = "Number of trials", y = NULL) +
  theme_rep()
```

```{r organ-table}
otbl <- organ_tab %>%
  mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(otbl,
   cap = "All MedDRA system organ classes: number of trials and percentage.",
   cn  = c("MedDRA Organ Class", "n", "% of trials"),
   al  = c("l","r","r"), lw  = TRUE)
```

\newpage

## Conditions (MedDRA Terms)

```{r cond-bar, fig.height=5.2, fig.cap="Top 20 MedDRA conditions. Trials may be classified under more than one condition."}
cond_tab %>%
  head(20) %>%
  mutate(
    lbl = str_trunc(MEDDRA_term, 45),
    lbl = factor(lbl, levels = rev(lbl))
  ) %>%
  ggplot(aes(x = n, y = lbl)) +
  geom_col(fill = TEAL, width = 0.72) +
  geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
            hjust = -0.08, size = 2.5, color = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25)), labels = comma) +
  labs(title = "MedDRA Conditions (top 20)", x = "Number of trials", y = NULL) +
  theme_rep()
```

```{r cond-table}
ctbl2 <- cond_tab %>%
  head(30) %>%
  mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(ctbl2,
   cap = "Top 30 MedDRA conditions: number of trials and percentage.",
   cn  = c("MedDRA Condition", "n", "% of trials"),
   al  = c("l","r","r"))
```

\newpage

# Trial Phases

```{r phase-bar, fig.height=3.0, fig.cap="Phase distribution across all selected trials (trials with multiple phases are counted once per phase)."}
phase_tab %>%
  mutate(phase = factor(phase, levels = phase)) %>%
  ggplot(aes(x = phase, y = n, fill = phase)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n", fmt_pct(pct))),
            vjust = -0.3, size = 2.6, color = GRAY) +
  scale_fill_manual(values = setNames(NORD8, phase_tab$phase)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
  labs(title = "Trial Phase Distribution", x = "Phase", y = "Number of trials") +
  theme_rep()
```

```{r phase-reg-bar, fig.height=3.2, fig.cap="Phase distribution split by source register."}
df %>%
  filter(!is.na(phase), phase != "") %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase != "") %>%
  count(phase, register) %>%
  ggplot(aes(x = phase, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  labs(title = "Phase by Register", x = "Phase", y = "Number of trials") +
  theme_rep()
```

```{r phase-tables, fig.show='hold', out.width='49%'}
# Phase summary
ptbl <- phase_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(ptbl, cn = c("Phase", "n", "%"), al = "lrr")

# Phase × register cross-tab
prtbl <- phase_reg %>%
  mutate(across(where(is.numeric), fmt_n))
nk(prtbl, cap = "Phase by register.", al = c("l", rep("r", ncol(prtbl)-1)))
```

\newpage

# Sponsor Analysis

```{r sponsor-bar, fig.show='hold', out.width='49%', fig.height=3.0, fig.cap="Left: sponsor type distribution. Right: sponsor type by source register."}
p_sp1 <- sponsor_tab %>%
  ggplot(aes(x = reorder(sponsor_type, n), y = n, fill = sponsor_type)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")),
            hjust = -0.1, size = 2.8, color = GRAY) +
  scale_fill_manual(values = c(Academic = GREEN, Industry = BLUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.28)), labels = comma) +
  coord_flip() +
  labs(title = "Sponsor Type", x = NULL, y = "Trials") +
  theme_rep()
print(p_sp1)

p_sp2 <- df %>%
  filter(!is.na(sponsor_type)) %>%
  count(sponsor_type, register) %>%
  ggplot(aes(x = sponsor_type, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  labs(title = "Sponsor Type by Register", x = NULL, y = "Trials") +
  theme_rep()
print(p_sp2)
```

```{r sponsor-table}
sptbl <- sponsor_reg %>%
  mutate(across(where(is.numeric), fmt_n))
nk(sptbl, cap = "Sponsor type by source register.",
   al = c("l", rep("r", ncol(sptbl)-1)))
```

## Top Sponsors by Name

```{r top-sponsor-bar, fig.height=5.5, fig.cap="Top 20 sponsors by number of trials. Colour indicates sponsor type (Academic/Industry)."}
if (nrow(top_sponsor_tab) > 0) {
  p_top_sp <- top_sponsor_tab %>%
    mutate(sponsor_name = factor(sponsor_name, levels = rev(sponsor_name)),
           sponsor_type = coalesce(sponsor_type, "Unknown")) %>%
    ggplot(aes(x = n, y = sponsor_name, fill = sponsor_type)) +
    geom_col(width = 0.65, show.legend = TRUE) +
    geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
              hjust = -0.1, size = 2.6, color = GRAY) +
    scale_fill_manual(values = c(Academic = GREEN, Industry = BLUE, Unknown = GRAY),
                      name = "Sponsor Type") +
    scale_x_continuous(expand = expansion(mult = c(0, 0.25)), labels = comma) +
    labs(title = "Top 20 Sponsors", x = "Trials", y = NULL) +
    theme_rep() +
    theme(legend.position = "bottom")
  print(p_top_sp)
} else {
  cat("No sponsor name data available for this filter selection.")
}
```

```{r top-sponsor-table}
if (nrow(top_sponsor_tab) > 0) {
  tsp_tbl <- top_sponsor_tab %>%
    select(sponsor_name, sponsor_type, n, pct) %>%
    rename(`Sponsor` = sponsor_name, `Type` = sponsor_type,
           `Trials` = n, `%` = pct) %>%
    mutate(Trials = fmt_n(Trials), `%` = fmt_pct(`%`))
  nk(tsp_tbl, cap = "Top 20 sponsors by trial count.",
     al = c("l", "l", "r", "r"))
}
```

\newpage

# Paediatric Investigation Plans (PIP)

```{r pip-bar, fig.show='hold', out.width='49%', fig.height=3.0, fig.cap="Left: overall PIP status. Right: PIP status by register."}
p_pip1 <- pip_tab %>%
  mutate(has_PIP = factor(has_PIP, levels = c("Yes","No","Unknown"))) %>%
  ggplot(aes(x = has_PIP, y = n, fill = has_PIP)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")),
            vjust = -0.3, size = 2.6, color = GRAY) +
  scale_fill_manual(values = PIP_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)), labels = comma) +
  labs(title = "PIP Status", x = NULL, y = "Trials") +
  theme_rep()
print(p_pip1)

p_pip2 <- df %>%
  mutate(has_PIP = coalesce(has_PIP, "Unknown")) %>%
  count(has_PIP, register) %>%
  ggplot(aes(x = has_PIP, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  labs(title = "PIP Status by Register", x = NULL, y = "Trials") +
  theme_rep()
print(p_pip2)
```

```{r pip-table}
ptbl2 <- pip_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct))
nk(ptbl2, cn = c("PIP Status", "n", "% of trials"), al = "lrr")
```

\newpage

# Decision Timeline

```{r decision-check, include=FALSE}
has_decision <- nrow(decision_df) > 0
```

```{r decision-violin, eval=has_decision, fig.height=3.8, fig.cap="Distribution of days from submission date to competent authority decision date, by source register and combined. Box shows IQR and median; whiskers extend to 1.5 × IQR."}
bind_rows(decision_df, mutate(decision_df, register = "All")) %>%
  mutate(register = factor(register, levels = c("EUCTR", "CTIS", "All"))) %>%
  ggplot(aes(x = register, y = days_to_decision, fill = register)) +
  geom_violin(alpha = 0.55, trim = TRUE, width = 0.8) +
  geom_boxplot(width = 0.18, outlier.size = 0.8, outlier.alpha = 0.4,
               colour = DARK, fill = "white") +
  scale_fill_manual(values = c(REG_COLS, All = PURPLE)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Time from Submission to Decision",
       x = "Register", y = "Days") +
  theme_rep() +
  theme(legend.position = "none")
```

```{r decision-table, eval=has_decision}
dec_stats <- bind_rows(decision_df, mutate(decision_df, register = "All")) %>%
  mutate(register = factor(register, levels = c("EUCTR", "CTIS", "All"))) %>%
  group_by(Register = register) %>%
  summarise(
    n         = fmt_n(n()),
    Mean_SD   = sprintf("%.1f \u00b1 %.1f", mean(days_to_decision), sd(days_to_decision)),
    Median    = sprintf("%.0f", median(days_to_decision)),
    IQR       = iqr_str(days_to_decision),
    Min_Max   = sprintf("%d\u2013%d", as.integer(min(days_to_decision)),
                                      as.integer(max(days_to_decision))),
    .groups = "drop"
  )
nk(dec_stats,
   cap = "Descriptive statistics for days from submission to decision, by register.",
   cn  = c("Register", "n", "Mean \u00b1 SD", "Median", "IQR", "Min\u2013Max"),
   al  = c("l","r","r","r","r","r"))
```

```{r decision-missing, eval=!has_decision, results='asis'}
cat("\\textit{No decision date data available for the current filter selection.}")
```

---
*Report generated by the EU Paediatric Clinical Trials dashboard on `r format(Sys.Date(), "%d %B %Y")`.*
*Data sources: EUCTR (EU Clinical Trials Register) and CTIS (Clinical Trials Information System).*
