---
title: "EU Paediatric Clinical Trials"
subtitle: "Summary Report · Dashboard v0.10.3"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 6.5
    fig_height: 3.4
    keep_tex: false
classoption: a4paper
header-includes:
  - \usepackage{fontspec}
  - \IfFontExistsTF{Helvetica}{\setmainfont{Helvetica}[BoldFont=Helvetica-Bold, ItalicFont=Helvetica-Oblique, BoldItalicFont=Helvetica-BoldOblique]}{\IfFontExistsTF{TeX Gyre Heros}{\setmainfont{TeX Gyre Heros}[BoldFont={TeX Gyre Heros Bold}, ItalicFont={TeX Gyre Heros Italic}, BoldItalicFont={TeX Gyre Heros Bold Italic}]}{}}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{float}
  - \usepackage{xcolor}
  - \usepackage{geometry}
  - \geometry{top=22mm,bottom=22mm,left=22mm,right=22mm}
  - \definecolor{nordblue}{HTML}{5E81AC}
  - \definecolor{nordgray}{HTML}{4C566A}
  - \setlength{\parskip}{5pt}
  - \setlength{\parindent}{0pt}
params:
  data_path: ""
  filters: !r list()
  chart_data_path: ""
  chart_type: "bar_stacked"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(stringr)
  library(lubridate)
  library(scales)
  library(knitr)
})

# ── Data ───────────────────────────────────────────────────────────────────────
if (nzchar(params$data_path) && file.exists(params$data_path)) {
  df <- readRDS(params$data_path)
} else {
  cache <- "trials_cache.rds"
  if (!file.exists(cache)) stop("Cache not found. Run rebuild_cache.R first.")
  df <- readRDS(cache)
}
N  <- nrow(df)
f  <- params$filters

if (N == 0) {
  cat("\\textbf{No trials match the current filter criteria.}")
  knitr::knit_exit()
}

# ── Colour palette ─────────────────────────────────────────────────────────────
BLUE   <- "#5E81AC"; FROST  <- "#81A1C1"; TEAL   <- "#88C0D0"
GREEN  <- "#A3BE8C"; YELLOW <- "#EBCB8B"; ORANGE <- "#D08770"
RED    <- "#BF616A"; PURPLE <- "#B48EAD"; GRAY   <- "#4C566A"
LIGHT  <- "#ECEFF4"; DARK   <- "#2E3440"
NORD8  <- c(BLUE, GREEN, YELLOW, RED, PURPLE, ORANGE, TEAL, FROST)

STATUS_COLS <- c(Ongoing = GREEN, Completed = BLUE, Other = ORANGE)
REG_COLS    <- c(EUCTR = BLUE, CTIS = TEAL)
PIP_COLS    <- c(Yes = GREEN, No = RED, Unknown = ORANGE)

# ── ggplot2 theme ──────────────────────────────────────────────────────────────
theme_rep <- function(base = 9) {
  theme_minimal(base_size = base, base_family = "sans") %+replace% theme(
    plot.title       = element_text(size = base + 1, face = "bold", color = DARK, hjust = 0),
    plot.subtitle    = element_text(size = base - 0.5, color = GRAY, hjust = 0),
    axis.title       = element_text(size = base - 0.5, color = GRAY),
    axis.text        = element_text(size = base - 1, color = GRAY),
    axis.text.y      = element_text(margin = margin(r = 2)),
    panel.grid.major = element_line(color = LIGHT, linewidth = 0.45),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    legend.key.size  = unit(0.32, "cm"),
    legend.text      = element_text(size = base - 1.5, color = GRAY),
    plot.margin      = margin(4, 6, 4, 4),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.text       = element_text(size = base - 0.5, face = "bold", color = DARK)
  )
}

# ── kable helper ───────────────────────────────────────────────────────────────
nk <- function(x, cap = NULL, cn = colnames(x), al = NULL, lw = FALSE) {
  kable(x, booktabs = TRUE, linesep = "", caption = cap,
        col.names = cn, align = al, longtable = lw, escape = TRUE)
}

fmt_pct <- function(x) sprintf("%.1f%%", x)
fmt_n   <- function(x) format(x, big.mark = ",")
iqr_str <- function(x) {
  q <- quantile(x, c(0.25, 0.75), na.rm = TRUE)
  sprintf("%.1f\u2013%.1f", q[1], q[2])
}

# ── Pre-compute summary tables ─────────────────────────────────────────────────
status_tab <- df %>%
  mutate(status = coalesce(status, "Unknown")) %>%
  count(status, name = "n") %>%
  mutate(pct = round(n / N * 100, 1)) %>%
  arrange(desc(n))

reg_tab <- df %>%
  count(register, name = "n") %>%
  mutate(pct = round(n / N * 100, 1)) %>%
  arrange(desc(n))

pip_tab <- df %>%
  mutate(has_PIP = coalesce(has_PIP, "Unknown")) %>%
  count(has_PIP, name = "n") %>%
  mutate(pct = round(n / N * 100, 1)) %>%
  arrange(match(has_PIP, c("Yes", "No", "Unknown")))

orphan_tab <- if ("is_orphan" %in% names(df)) {
  df %>%
    mutate(is_orphan = coalesce(is_orphan, "Unknown")) %>%
    count(is_orphan, name = "n") %>%
    mutate(pct = round(n / N * 100, 1)) %>%
    arrange(match(is_orphan, c("Yes", "No", "Unknown")))
} else {
  NULL
}

sponsor_tab <- df %>%
  filter(!is.na(sponsor_type)) %>%
  count(sponsor_type, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(n))

phase_tab <- df %>%
  filter(!is.na(phase), phase != "") %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase %in% c("Phase I","Phase II","Phase III","Phase IV")) %>%
  count(phase, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  arrange(match(phase, c("Phase I","Phase II","Phase III","Phase IV")))

country_tab <- df %>%
  filter(!is.na(Member_state)) %>%
  separate_rows(Member_state, sep = " / ") %>%
  mutate(Member_state = str_trim(Member_state)) %>%
  filter(Member_state != "") %>%
  count(Member_state, name = "n", sort = TRUE) %>%
  mutate(pct = round(n / N * 100, 1))

organ_tab <- df %>%
  filter(!is.na(MEDDRA_organ_class)) %>%
  separate_rows(MEDDRA_organ_class, sep = " / ") %>%
  mutate(MEDDRA_organ_class = str_trim(MEDDRA_organ_class)) %>%
  filter(MEDDRA_organ_class != "") %>%
  count(MEDDRA_organ_class, name = "n", sort = TRUE) %>%
  mutate(pct = round(n / N * 100, 1))

cond_tab <- df %>%
  filter(!is.na(MEDDRA_term)) %>%
  separate_rows(MEDDRA_term, sep = " / ") %>%
  mutate(MEDDRA_term = str_trim(MEDDRA_term)) %>%
  filter(MEDDRA_term != "") %>%
  count(MEDDRA_term, name = "n", sort = TRUE) %>%
  mutate(pct = round(n / N * 100, 1))

year_reg <- df %>%
  filter(!is.na(submission_date_parsed)) %>%
  mutate(year = year(submission_date_parsed)) %>%
  count(year, register, name = "n") %>%
  arrange(year)

year_total <- year_reg %>%
  group_by(year) %>% summarise(n = sum(n), .groups = "drop") %>%
  arrange(year) %>% mutate(cumulative = cumsum(n))

countries_per_trial <- df %>%
  filter(!is.na(Member_state)) %>%
  rowwise() %>%
  mutate(nc = length(str_trim(str_split(Member_state, " / ")[[1]]))) %>%
  pull(nc)

conds_per_trial <- df %>%
  filter(!is.na(MEDDRA_term)) %>%
  rowwise() %>%
  mutate(nc = length(str_trim(str_split(MEDDRA_term, " / ")[[1]]))) %>%
  pull(nc)

top_sponsor_tab <- df %>%
  filter(!is.na(sponsor_name)) %>%
  count(sponsor_name, sponsor_type, sort = TRUE) %>%
  slice_head(n = 20) %>%
  mutate(pct = round(n / N * 100, 1))

sponsor_reg <- df %>%
  filter(!is.na(sponsor_type)) %>%
  count(sponsor_type, register) %>%
  pivot_wider(names_from = register, values_from = n, values_fill = 0) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>%
  arrange(desc(Total))

phase_reg <- df %>%
  filter(!is.na(phase), phase != "") %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase %in% c("Phase I","Phase II","Phase III","Phase IV")) %>%
  count(phase, register) %>%
  pivot_wider(names_from = register, values_from = n, values_fill = 0) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>%
  arrange(match(phase, c("Phase I","Phase II","Phase III","Phase IV")))

decision_df <- df %>%
  filter(!is.na(days_to_decision), is.finite(days_to_decision),
         days_to_decision >= 0, days_to_decision < 3650)

# Results posting (completed trials)
has_results_col <- "has_results" %in% names(df)
results_base <- df %>% filter(status == "Completed", !is.na(decision_date))
```

# Applied Filters {-}

The following filters were active when this report was generated.

```{r filter-table}
fnames <- c(
  "Trial Status", "Source Register",
  "Submission Date (from)", "Submission Date (to)",
  "Sponsor / Company", "Country / Member State",
  "Trial Phase", "Part of PIP", "Orphan Designation",
  "MedDRA Organ Class", "Condition / MedDRA Term",
  "Free-text search"
)
fvals <- c(
  if (length(f$status)   == 0) "All" else paste(f$status,   collapse = ", "),
  if (length(f$register) == 0) "All" else paste(f$register, collapse = ", "),
  if (length(f$date_range) >= 1) f$date_range[1] else "\u2014",
  if (length(f$date_range) >= 2) f$date_range[2] else "\u2014",
  if (length(f$sponsor) == 0 || identical(f$sponsor, "All")) "All"
    else paste(f$sponsor, collapse = "; "),
  if (length(f$country) == 0 || identical(f$country, "All")) "All"
    else paste(f$country, collapse = "; "),
  if (length(f$phase) == 0 || identical(f$phase, "All")) "All"
    else paste(f$phase, collapse = ", "),
  if (is.null(f$pip)    || f$pip    == "All") "All" else f$pip,
  if (is.null(f$orphan) || f$orphan == "All") "All" else f$orphan,
  if (length(f$organ_class) == 0 || identical(f$organ_class, "All")) "All"
    else paste(f$organ_class, collapse = "; "),
  if (length(f$condition) == 0 || identical(f$condition, "All")) "All"
    else paste(f$condition, collapse = "; "),
  if (is.null(f$text_search) || f$text_search == "(none)" || !nzchar(f$text_search))
    "(none)" else f$text_search
)
nk(data.frame(Filter = fnames, Value = fvals),
   cn = c("Filter", "Value"), al = c("l","l"))
```

\newpage

# Overview

```{r overview-summary}
date_range_str <- if (!all(is.na(df$submission_date_parsed))) {
  paste(format(min(df$submission_date_parsed, na.rm = TRUE), "%d %b %Y"), "to",
        format(max(df$submission_date_parsed, na.rm = TRUE), "%d %b %Y"))
} else "n/a"

n_pip_yes     <- sum(df$has_PIP == "Yes",  na.rm = TRUE)
n_ongoing     <- sum(df$status  == "Ongoing",   na.rm = TRUE)
n_completed   <- sum(df$status  == "Completed", na.rm = TRUE)

overview_df <- tibble(
  Metric = c(
    "Total trials in selection", "Submission date range",
    "Ongoing", "Completed",
    "Part of a PIP",
    "Unique member states", "Unique MedDRA organ classes",
    "Unique MedDRA conditions"
  ),
  Value = c(
    fmt_n(N), date_range_str,
    paste0(fmt_n(n_ongoing),   " (", round(n_ongoing   / N * 100, 1), "%)"),
    paste0(fmt_n(n_completed), " (", round(n_completed / N * 100, 1), "%)"),
    paste0(fmt_n(n_pip_yes),   " (", round(n_pip_yes   / N * 100, 1), "%)"),
    as.character(nrow(country_tab)),
    as.character(nrow(organ_tab)),
    fmt_n(nrow(cond_tab))
  )
)
nk(overview_df, cap = paste("Key metrics for the", fmt_n(N), "selected trials."),
   cn = c("Metric", "Value"), al = c("l","l"))
```

## Trial Status and Register Distribution

```{r status-register, fig.show='hold', out.width='49%', fig.height=3.2, fig.cap="Left: trial status distribution. Right: source register distribution."}
p1 <- status_tab %>%
  mutate(status = factor(status, levels = status),
         label  = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")) %>%
  ggplot(aes(x = reorder(status, n), y = n, fill = status)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = label), hjust = -0.1, size = 2.8, color = GRAY) +
  scale_fill_manual(values = STATUS_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.22)), labels = comma) +
  coord_flip() +
  labs(title = "Trial Status", x = NULL, y = "Number of trials") +
  theme_rep()
print(p1)

p2 <- reg_tab %>%
  mutate(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")) %>%
  ggplot(aes(x = reorder(register, n), y = n, fill = register)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = label), hjust = -0.1, size = 2.8, color = GRAY) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.22)), labels = comma) +
  coord_flip() +
  labs(title = "Source Register", x = NULL, y = "Number of trials") +
  theme_rep()
print(p2)
```

```{r status-register-tables, fig.show='hold', out.width='49%'}
nk(status_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cn = c("Status", "n", "%"), al = "lrr")
nk(reg_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cn = c("Register", "n", "%"), al = "lrr")
```

## Per-trial Descriptive Statistics

```{r continuous-table}
cont_df <- tibble(
  Metric  = c("Member states per trial", "Conditions (MedDRA terms) per trial",
              "Submissions per year"),
  Mean_SD = c(
    sprintf("%.1f \u00b1 %.1f", mean(countries_per_trial), sd(countries_per_trial)),
    sprintf("%.1f \u00b1 %.1f", mean(conds_per_trial),     sd(conds_per_trial)),
    sprintf("%.1f \u00b1 %.1f", mean(year_total$n),        sd(year_total$n))
  ),
  Median  = c(
    sprintf("%.1f", median(countries_per_trial)),
    sprintf("%.1f", median(conds_per_trial)),
    sprintf("%.1f", median(year_total$n))
  ),
  IQR     = c(
    iqr_str(countries_per_trial),
    iqr_str(conds_per_trial),
    iqr_str(year_total$n)
  ),
  Min_Max = c(
    sprintf("%d\u2013%d", min(countries_per_trial), max(countries_per_trial)),
    sprintf("%d\u2013%d", min(conds_per_trial),     max(conds_per_trial)),
    sprintf("%d\u2013%d", min(year_total$n),         max(year_total$n))
  )
)
nk(cont_df,
   cap = "Descriptive statistics for continuous per-trial and per-year metrics.",
   cn  = c("Metric", "Mean \u00b1 SD", "Median", "IQR", "Min\u2013Max"),
   al  = c("l","r","r","r","r"))
```

\newpage

# Submission Timeline

```{r timeline-bar, fig.height=3.4, fig.cap="Annual trial submissions by source register. CTIS bars appear from 2023 onward; pre-2023 CTIS records (migrated from EudraCT) are counted as EUCTR."}
year_reg %>%
  ggplot(aes(x = year, y = n, fill = register)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  scale_x_continuous(breaks = pretty_breaks(8)) +
  labs(title = "Submissions per Year", x = NULL, y = "Number of trials") +
  theme_rep()
```

```{r timeline-cumul, fig.height=3.0, fig.cap="Cumulative trial submissions over time."}
year_total %>%
  ggplot(aes(x = year, y = cumulative)) +
  geom_area(fill = BLUE, alpha = 0.18) +
  geom_line(color = BLUE, linewidth = 0.9) +
  geom_point(color = BLUE, size = 1.8) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  scale_x_continuous(breaks = pretty_breaks(8)) +
  labs(title = "Cumulative Submissions", x = NULL, y = "Cumulative trials") +
  theme_rep()
```

```{r timeline-table}
ytbl <- year_total %>%
  mutate(n = fmt_n(n), cumulative = fmt_n(cumulative)) %>%
  rename(Year = year, Submissions = n, Cumulative = cumulative)
nk(ytbl, cap = "Submissions and cumulative total by year.", al = c("l","r","r"))
```

\newpage

# Geographic Distribution

```{r countries-bar, fig.height=5.5, fig.cap="Top 20 EU member states by number of trial participations. A trial spanning multiple countries is counted once per country."}
country_tab %>%
  head(20) %>%
  mutate(Member_state = factor(Member_state, levels = rev(Member_state))) %>%
  ggplot(aes(x = n, y = Member_state)) +
  geom_col(fill = FROST, width = 0.7) +
  geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
            hjust = -0.08, size = 2.6, color = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2)), labels = comma) +
  labs(title = "Trials by Member State (top 20)", x = "Number of trials", y = NULL) +
  theme_rep()
```

```{r countries-table}
nk(country_tab %>% head(20) %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cap = "Top 20 member states: trial participations and percentage of all selected trials.",
   cn  = c("Member State", "n", "% of trials"), al = c("l","r","r"))
```

\newpage

# Therapeutic Areas

## MedDRA Organ Classes

```{r organ-bar, fig.height=4.8, fig.cap="Trials per MedDRA system organ class (top 15). Trials may belong to more than one class."}
organ_tab %>%
  head(15) %>%
  mutate(lbl = str_wrap(MEDDRA_organ_class, 35),
         lbl = factor(lbl, levels = rev(lbl))) %>%
  ggplot(aes(x = n, y = lbl)) +
  geom_col(fill = PURPLE, width = 0.72) +
  geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
            hjust = -0.08, size = 2.5, color = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25)), labels = comma) +
  labs(title = "MedDRA Organ Classes (top 15)", x = "Number of trials", y = NULL) +
  theme_rep()
```

```{r organ-table}
nk(organ_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cap = "All MedDRA system organ classes: number of trials and percentage.",
   cn  = c("MedDRA Organ Class", "n", "% of trials"),
   al  = c("l","r","r"), lw = TRUE)
```

\newpage

## Conditions (MedDRA Terms)

```{r cond-bar, fig.height=5.2, fig.cap="Top 20 MedDRA conditions. Trials may be classified under more than one condition."}
cond_tab %>%
  head(20) %>%
  mutate(lbl = str_trunc(MEDDRA_term, 45),
         lbl = factor(lbl, levels = rev(lbl))) %>%
  ggplot(aes(x = n, y = lbl)) +
  geom_col(fill = TEAL, width = 0.72) +
  geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
            hjust = -0.08, size = 2.5, color = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25)), labels = comma) +
  labs(title = "MedDRA Conditions (top 20)", x = "Number of trials", y = NULL) +
  theme_rep()
```

```{r cond-table}
nk(cond_tab %>% head(30) %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cap = "Top 30 MedDRA conditions: number of trials and percentage.",
   cn  = c("MedDRA Condition", "n", "% of trials"), al = c("l","r","r"))
```

\newpage

# Trial Phases

```{r phase-dist, fig.height=3.0, fig.cap="Phase distribution across all selected trials. Trials with combined phase designations (e.g. Phase I/II) are counted once per constituent phase."}
phase_tab %>%
  ggplot(aes(x = phase, y = n, fill = phase)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")),
            vjust = -0.3, size = 2.6, color = GRAY) +
  scale_fill_manual(values = setNames(NORD8[seq_along(phase_tab$phase)], phase_tab$phase)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.22)), labels = comma) +
  labs(title = "Phase Distribution", x = NULL, y = "Number of trials") +
  theme_rep()
```

```{r phase-by-register, fig.show='hold', out.width='49%', fig.height=3.2, fig.cap="Left: phase by source register. Right: phase by trial status."}
df %>%
  filter(!is.na(phase), phase != "") %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase %in% c("Phase I","Phase II","Phase III","Phase IV")) %>%
  count(phase, register) %>%
  ggplot(aes(x = phase, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10)), labels = comma) +
  labs(title = "Phase by Register", x = NULL, y = "Number of trials") +
  theme_rep()

df %>%
  filter(!is.na(phase), phase != "", !is.na(status)) %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase %in% c("Phase I","Phase II","Phase III","Phase IV")) %>%
  count(phase, status) %>%
  ggplot(aes(x = phase, y = n, fill = status)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = STATUS_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10)), labels = comma) +
  labs(title = "Phase by Status", x = NULL, y = "Number of trials") +
  theme_rep()
```

```{r phase-by-sponsor, fig.height=3.2, fig.cap="Phase distribution by sponsor type (Academic vs Industry)."}
df %>%
  filter(!is.na(phase), phase != "", !is.na(sponsor_type)) %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(phase %in% c("Phase I","Phase II","Phase III","Phase IV")) %>%
  count(phase, sponsor_type) %>%
  ggplot(aes(x = phase, y = n, fill = sponsor_type)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = c(Academic = GREEN, Industry = BLUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10)), labels = comma) +
  labs(title = "Phase by Sponsor Type", x = NULL, y = "Number of trials") +
  theme_rep()
```

```{r phase-tables, fig.show='hold', out.width='49%'}
nk(phase_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cn = c("Phase", "n", "%"), al = "lrr")
nk(phase_reg %>% mutate(across(where(is.numeric), fmt_n)),
   cap = "Phase by register.", al = c("l", rep("r", ncol(phase_reg) - 1)))
```

## Completion Rate by Authorisation Cohort

```{r completion-cohort, fig.height=3.4, fig.cap="Percentage of trials authorised in each calendar year that have subsequently reached completed status, by source register. Cohorts with fewer than five trials are excluded. More recent cohorts naturally show lower completion rates owing to shorter observation windows."}
cohort_df <- df %>%
  filter(!is.na(decision_date), !is.na(status)) %>%
  mutate(auth_year = year(decision_date)) %>%
  filter(auth_year >= 2004, auth_year < year(Sys.Date())) %>%
  group_by(auth_year, register) %>%
  summarise(n_total     = n(),
            n_completed = sum(status == "Completed", na.rm = TRUE),
            .groups     = "drop") %>%
  filter(n_total >= 5) %>%
  mutate(pct_completed = n_completed / n_total * 100)

if (nrow(cohort_df) > 0) {
  cohort_df %>%
    ggplot(aes(x = auth_year, y = pct_completed,
               colour = register, group = register)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 1.8) +
    scale_colour_manual(values = REG_COLS) +
    scale_x_continuous(breaks = pretty_breaks(8)) +
    scale_y_continuous(limits = c(0, 105),
                       labels = function(x) paste0(x, "%")) +
    labs(title = "Completion Rate by Authorisation Cohort",
         x = "Authorisation Year", y = "% Completed") +
    theme_rep()
} else {
  cat("\\textit{Insufficient data to render completion cohort chart.}")
}
```

```{r completion-cohort-table}
if (nrow(cohort_df) > 0) {
  nk(cohort_df %>%
       mutate(pct_completed = sprintf("%.1f%%", pct_completed)) %>%
       select(auth_year, register, n_total, n_completed, pct_completed) %>%
       rename(Year = auth_year, Register = register,
              Authorised = n_total, Completed = n_completed, `% Completed` = pct_completed),
     cap = "Completion rates by authorisation year and register (cohorts with \u2265 5 trials).",
     al  = c("l","l","r","r","r"))
}
```

\newpage

# Sponsor Analysis

```{r sponsor-type, fig.show='hold', out.width='49%', fig.height=3.0, fig.cap="Left: sponsor type distribution. Right: sponsor type by source register."}
sponsor_tab %>%
  ggplot(aes(x = reorder(sponsor_type, n), y = n, fill = sponsor_type)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")),
            hjust = -0.1, size = 2.8, color = GRAY) +
  scale_fill_manual(values = c(Academic = GREEN, Industry = BLUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.28)), labels = comma) +
  coord_flip() +
  labs(title = "Sponsor Type", x = NULL, y = "Trials") +
  theme_rep()

df %>%
  filter(!is.na(sponsor_type)) %>%
  count(sponsor_type, register) %>%
  ggplot(aes(x = sponsor_type, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10)), labels = comma) +
  labs(title = "Sponsor Type by Register", x = NULL, y = "Trials") +
  theme_rep()
```

```{r sponsor-reg-table}
nk(sponsor_reg %>% mutate(across(where(is.numeric), fmt_n)),
   cap = "Sponsor type by source register.",
   al  = c("l", rep("r", ncol(sponsor_reg) - 1)))
```

## Top Sponsors

```{r top-sponsor-bar, fig.height=5.5, fig.cap="Top 20 sponsors by trial count. Sponsor names are normalised (legal suffixes stripped, canonical brand names applied). Colour indicates sponsor type."}
if (nrow(top_sponsor_tab) > 0) {
  top_sponsor_tab %>%
    mutate(sponsor_name = factor(sponsor_name, levels = rev(sponsor_name)),
           sponsor_type = coalesce(sponsor_type, "Unknown")) %>%
    ggplot(aes(x = n, y = sponsor_name, fill = sponsor_type)) +
    geom_col(width = 0.65) +
    geom_text(aes(label = paste0(fmt_n(n), " (", fmt_pct(pct), ")")),
              hjust = -0.1, size = 2.6, color = GRAY) +
    scale_fill_manual(values = c(Academic = GREEN, Industry = BLUE, Unknown = GRAY),
                      name = "Sponsor Type") +
    scale_x_continuous(expand = expansion(mult = c(0, 0.25)), labels = comma) +
    labs(title = "Top 20 Sponsors", x = "Trials", y = NULL) +
    theme_rep()
} else {
  cat("\\textit{No sponsor name data available for this filter selection.}")
}
```

```{r top-sponsor-table}
if (nrow(top_sponsor_tab) > 0) {
  nk(top_sponsor_tab %>%
       select(sponsor_name, sponsor_type, n, pct) %>%
       rename(Sponsor = sponsor_name, Type = sponsor_type,
              Trials = n, `%` = pct) %>%
       mutate(Trials = fmt_n(Trials), `%` = fmt_pct(`%`)),
     cap = "Top 20 sponsors by trial count.",
     al  = c("l","l","r","r"))
}
```

\newpage

# Regulatory Status

## Paediatric Investigation Plans (PIP)

A Paediatric Investigation Plan (PIP) is a development plan aimed at ensuring that medicines are researched and authorised for use in children. PIP involvement is recorded in the registry at submission.

```{r pip-bar, fig.show='hold', out.width='49%', fig.height=3.0, fig.cap="Left: overall PIP status. Right: PIP status by source register."}
pip_tab %>%
  mutate(has_PIP = factor(has_PIP, levels = c("Yes","No","Unknown"))) %>%
  ggplot(aes(x = has_PIP, y = n, fill = has_PIP)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")),
            vjust = -0.3, size = 2.6, color = GRAY) +
  scale_fill_manual(values = PIP_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)), labels = comma) +
  labs(title = "PIP Status", x = NULL, y = "Trials") +
  theme_rep()

df %>%
  mutate(has_PIP = coalesce(has_PIP, "Unknown")) %>%
  count(has_PIP, register) %>%
  ggplot(aes(x = has_PIP, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10)), labels = comma) +
  labs(title = "PIP Status by Register", x = NULL, y = "Trials") +
  theme_rep()
```

```{r pip-table}
nk(pip_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cn = c("PIP Status", "n", "% of trials"), al = "lrr")
```

## Orphan Designation

Orphan designation is granted for medicinal products intended for rare diseases (prevalence \(\leq\) 5 per 10,000 in the EU). The designation is derived from EUCTR DIMP field D.2.5 and CTIS orphan designation numbers (EU/3/\ldots format).

```{r orphan-check, include=FALSE}
has_orphan_col <- !is.null(orphan_tab)
```

```{r orphan-bar, eval=has_orphan_col, fig.show='hold', out.width='49%', fig.height=3.0, fig.cap="Left: orphan designation status. Right: orphan designation by source register."}
orphan_tab %>%
  mutate(is_orphan = factor(is_orphan, levels = c("Yes","No","Unknown"))) %>%
  ggplot(aes(x = is_orphan, y = n, fill = is_orphan)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = paste0(fmt_n(n), "\n(", fmt_pct(pct), ")")),
            vjust = -0.3, size = 2.6, color = GRAY) +
  scale_fill_manual(values = PIP_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)), labels = comma) +
  labs(title = "Orphan Designation", x = NULL, y = "Trials") +
  theme_rep()

df %>%
  mutate(is_orphan = coalesce(is_orphan, "Unknown")) %>%
  count(is_orphan, register) %>%
  ggplot(aes(x = is_orphan, y = n, fill = register)) +
  geom_col(position = position_dodge(0.72), width = 0.65) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10)), labels = comma) +
  labs(title = "Orphan Designation by Register", x = NULL, y = "Trials") +
  theme_rep()
```

```{r orphan-table, eval=has_orphan_col}
nk(orphan_tab %>% mutate(n = fmt_n(n), pct = fmt_pct(pct)),
   cn = c("Orphan Designation", "n", "% of trials"), al = "lrr")
```

```{r orphan-missing, eval=!has_orphan_col, results='asis'}
cat("\\textit{Orphan designation data requires a cache rebuild (added in v0.7.0). Run rebuild\\_cache.R to activate.}")
```

\newpage

# Result Reporting

This section shows which completed trials have posted results to the registry and which have not. Results are sourced directly from EUCTR (`endPoints.endPoint.readyForValues`) and CTIS (`resultsFirstReceived`). Trial end dates are not reliably available across registers; authorization date is shown for reference only and is not used as a compliance deadline.

```{r results-summary}
n_completed_total <- nrow(results_base)

if (has_results_col) {
  n_posted    <- sum(results_base$has_results, na.rm = TRUE)
  n_not_posted <- n_completed_total - n_posted
  pct_posted  <- if (n_completed_total > 0) round(n_posted / n_completed_total * 100, 1) else 0

  res_summary <- tibble(
    Metric = c("Completed trials", "Results posted", "No results posted",
               "Results posting rate"),
    Value  = c(fmt_n(n_completed_total),
               paste0(fmt_n(n_posted),     " (", pct_posted, "%)"),
               paste0(fmt_n(n_not_posted), " (", round(100 - pct_posted, 1), "%)"),
               paste0(pct_posted, "%"))
  )
} else {
  res_summary <- tibble(
    Metric = "Results data status",
    Value  = "Not available — rebuild cache to activate (added in v0.7.0)"
  )
}
nk(res_summary, cn = c("Metric", "Value"), al = c("l","l"))
```

```{r results-by-year, eval=has_results_col, fig.height=3.4, fig.cap="Completed trials by authorization year. Green = results confirmed posted in the registry; red = no results posted."}
results_base %>%
  mutate(auth_year     = year(decision_date),
         results_status = ifelse(has_results, "Results posted", "No results posted"),
         results_status = factor(results_status,
           levels = c("Results posted", "No results posted"))) %>%
  filter(auth_year >= 2004) %>%
  count(auth_year, results_status) %>%
  ggplot(aes(x = auth_year, y = n, fill = results_status)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = c("Results posted" = GREEN, "No results posted" = RED)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  scale_x_continuous(breaks = pretty_breaks(8)) +
  labs(title = "Result Reporting by Authorization Year",
       x = "Authorization Year", y = "Completed Trials") +
  theme_rep()
```

```{r results-by-sponsor, eval=has_results_col, fig.height=3.0, fig.cap="Results posting status for completed trials by sponsor type."}
results_base %>%
  filter(!is.na(sponsor_type)) %>%
  mutate(results_status = ifelse(has_results, "Results posted", "No results posted"),
         results_status = factor(results_status,
           levels = c("Results posted", "No results posted"))) %>%
  count(sponsor_type, results_status) %>%
  ggplot(aes(x = sponsor_type, y = n, fill = results_status)) +
  geom_col(width = 0.65) +
  scale_fill_manual(values = c("Results posted" = GREEN, "No results posted" = RED)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)), labels = comma) +
  labs(title = "Result Reporting by Sponsor Type",
       x = NULL, y = "Completed Trials") +
  theme_rep()
```


\newpage

# Decision Timeline

```{r decision-check, include=FALSE}
has_decision <- nrow(decision_df) > 0
```

```{r decision-violin, eval=has_decision, fig.height=3.8, fig.cap="Distribution of days from submission date to competent authority decision date, by source register. Box shows IQR and median; whiskers extend to 1.5 x IQR."}
bind_rows(decision_df, mutate(decision_df, register = "All")) %>%
  mutate(register = factor(register, levels = c("EUCTR", "CTIS", "All"))) %>%
  ggplot(aes(x = register, y = days_to_decision, fill = register)) +
  geom_violin(alpha = 0.55, trim = TRUE, width = 0.8) +
  geom_boxplot(width = 0.18, outlier.size = 0.8, outlier.alpha = 0.4,
               colour = DARK, fill = "white") +
  scale_fill_manual(values = c(REG_COLS, All = PURPLE)) +
  scale_y_log10(labels = comma) +
  labs(title = "Time from Submission to Decision",
       x = "Register", y = "Days (log10 scale)") +
  theme_rep() +
  theme(legend.position = "none")
```

```{r decision-table, eval=has_decision}
nk(bind_rows(decision_df, mutate(decision_df, register = "All")) %>%
     mutate(register = factor(register, levels = c("EUCTR", "CTIS", "All"))) %>%
     group_by(Register = register) %>%
     summarise(
       n       = fmt_n(n()),
       Mean_SD = sprintf("%.1f \u00b1 %.1f", mean(days_to_decision), sd(days_to_decision)),
       Median  = sprintf("%.0f", median(days_to_decision)),
       IQR     = iqr_str(days_to_decision),
       Min_Max = sprintf("%d\u2013%d", as.integer(min(days_to_decision)),
                                       as.integer(max(days_to_decision))),
       .groups = "drop"),
   cap = "Days from submission to decision, by register.",
   cn  = c("Register", "n", "Mean \u00b1 SD", "Median", "IQR", "Min\u2013Max"),
   al  = c("l","r","r","r","r","r"))
```

```{r decision-missing, eval=!has_decision, results='asis'}
cat("\\textit{No decision date data available for the current filter selection.}")
```

\newpage

# Custom Chart Builder

```{r custom-chart-setup, include=FALSE}
EXPLORE_LABELS <- c(
  "year"               = "Year of Submission",
  "status"             = "Status",
  "register"           = "Register",
  "phase"              = "Phase",
  "sponsor_type"       = "Sponsor Type",
  "has_PIP"            = "PIP Status",
  "is_orphan"          = "Orphan Designation",
  "MEDDRA_organ_class" = "Organ Class (MedDRA SOC)",
  "MEDDRA_term"        = "Condition (MedDRA term)",
  "Member_state"       = "Country / Member State"
)

chart_ed <- if (nzchar(params$chart_data_path) && file.exists(params$chart_data_path)) {
  readRDS(params$chart_data_path)
} else {
  list(data = data.frame(), x_var = "submission_year", grp = "None")
}
chart_df      <- chart_ed$data
chart_x_var   <- chart_ed$x_var
chart_grp     <- chart_ed$grp
chart_type    <- params$chart_type
chart_x_lbl   <- unname(EXPLORE_LABELS[chart_x_var])
chart_grp_lbl <- if (!is.null(chart_grp) && chart_grp != "None")
                   unname(EXPLORE_LABELS[chart_grp]) else NULL
has_custom    <- nrow(chart_df) > 0

if (chart_x_var == "year" && has_custom) {
  chart_df <- chart_df %>% mutate(x_val = as.integer(x_val))
}
```

```{r custom-chart-note, eval=!has_custom, results='asis'}
cat("\\textit{No data available for the custom chart selection.}")
```

```{r custom-chart-plot, eval=has_custom, fig.height=3.6}
y_lbl <- if (chart_type == "bar_pct") "Percentage of Trials (%)" else "Number of Trials"

if (is.null(chart_grp) || chart_grp == "None") {
  p <- ggplot(chart_df, aes(x = x_val, y = n))
  if (chart_type == "line") {
    p <- p + geom_line(colour = BLUE, linewidth = 0.8) +
             geom_point(colour = BLUE, size = 1.5)
  } else {
    p <- p + geom_col(fill = BLUE)
  }
  p <- p + labs(x = chart_x_lbl, y = y_lbl)
} else {
  n_grps <- n_distinct(chart_df$grp_val)
  pal    <- setNames(
    colorRampPalette(c(BLUE, FROST, ORANGE, GREEN, PURPLE, RED, YELLOW))(n_grps),
    sort(unique(chart_df$grp_val)))

  if (chart_type == "line") {
    p <- ggplot(chart_df, aes(x = x_val, y = n, colour = grp_val, group = grp_val)) +
      geom_line(linewidth = 0.8) + geom_point(size = 1.5) +
      scale_colour_manual(values = pal, name = chart_grp_lbl)
  } else {
    pos <- switch(chart_type,
      bar_grouped = position_dodge(width = 0.8),
      bar_pct     = "fill",
      "stack")
    p <- ggplot(chart_df, aes(x = x_val, y = n, fill = grp_val)) +
      geom_col(position = pos) +
      scale_fill_manual(values = pal, name = chart_grp_lbl)
  }
  p <- p + labs(x = chart_x_lbl, y = y_lbl)
}

if (!is.null(chart_x_var) && chart_x_var == "year") {
  p <- p + scale_x_continuous(breaks = scales::breaks_width(1))
}
if (chart_type == "bar_pct") {
  p <- p + scale_y_continuous(labels = scales::percent)
}
p + theme_rep()
```

```{r custom-chart-stats, eval=has_custom}
grand_total <- sum(chart_df$n)

if (is.null(chart_grp) || chart_grp == "None") {
  nk(data.frame(
       Statistic = c("Total", "Mean", "Median", "SD", "Min", "Max"),
       Value     = c(fmt_n(grand_total),
                     sprintf("%.1f", mean(chart_df$n)),
                     sprintf("%.1f", median(chart_df$n)),
                     sprintf("%.1f", sd(chart_df$n)),
                     fmt_n(min(chart_df$n)),
                     fmt_n(max(chart_df$n))),
       stringsAsFactors = FALSE),
     cap = paste0("Summary statistics for trial counts by ", chart_x_lbl, "."),
     al  = c("l","r"))
} else {
  nk(chart_df %>%
       group_by(!!as.name(chart_grp_lbl) := grp_val) %>%
       summarise(
         Total    = fmt_n(sum(n)),
         `% Total` = sprintf("%.1f%%", sum(n) / grand_total * 100),
         Mean     = sprintf("%.1f", mean(n)),
         Median   = sprintf("%.1f", median(n)),
         SD       = sprintf("%.1f", sd(n)),
         Min      = fmt_n(min(n)),
         Max      = fmt_n(max(n)),
         .groups  = "drop") %>%
       arrange(desc(as.numeric(gsub(",", "", Total)))),
     cap = paste0("Summary statistics by ", chart_grp_lbl,
                  " for trial counts grouped by ", chart_x_lbl, "."),
     al  = c("l", rep("r", 7)))
}
```

---
*Report generated by the EU Paediatric Clinical Trials dashboard on `r format(Sys.Date(), "%d %B %Y")`.*
*Data sources: EUCTR (EU Clinical Trials Register) and CTIS (Clinical Trials Information System).*
*Search query: all trials with at least one subject under 18 in the EU.*
