---
title: "Data Preprocessing Report"
subtitle: "EU Paediatric Clinical Trials - Pipeline Audit"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    self_contained: true
    fig_width: 9
    fig_height: 4.5
params:
  cache_path: "../trials_cache.rds"
  log_dir:    "../data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 120
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(stringr)
  library(lubridate)
  library(scales)
  library(knitr)
  library(kableExtra)
  library(readr)
  library(DT)
})

has_pkg <- function(pkg) requireNamespace(pkg, quietly = TRUE)

# Palette
BLUE   <- "#5E81AC"; FROST  <- "#81A1C1"; TEAL   <- "#88C0D0"
GREEN  <- "#A3BE8C"; YELLOW <- "#EBCB8B"; ORANGE <- "#D08770"
RED    <- "#BF616A"; PURPLE <- "#B48EAD"; GRAY   <- "#4C566A"
LIGHT  <- "#ECEFF4"; DARK   <- "#2E3440"; PALE <- "#F8FAFC"
NORD8  <- c(BLUE, GREEN, YELLOW, RED, PURPLE, ORANGE, TEAL, FROST)

REG_COLS <- c(EUCTR = BLUE, CTIS = TEAL)
STATUS_COLS <- c(Ongoing = GREEN, Completed = BLUE, Withdrawn = ORANGE, Other = RED)

theme_pre <- function(base = 11) {
  theme_minimal(base_size = base, base_family = "sans") %+replace% theme(
    plot.title       = element_text(size = base + 2, face = "bold", color = DARK, hjust = 0),
    plot.subtitle    = element_text(size = base, color = GRAY, hjust = 0, margin = margin(t = 2, b = 6)),
    axis.title       = element_text(size = base, color = GRAY),
    axis.text        = element_text(size = base - 1, color = GRAY),
    axis.text.y      = element_text(margin = margin(r = 4)),
    panel.grid.major = element_line(color = LIGHT, linewidth = 0.45),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    legend.key.size  = unit(0.4, "cm"),
    legend.text      = element_text(size = base - 1, color = GRAY),
    plot.margin      = margin(8, 12, 8, 8),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.text       = element_text(size = base, face = "bold", color = DARK)
  )
}

fmt_n <- function(x) {
  if (length(x) == 0 || all(is.na(x))) return("NA")
  format(round(as.numeric(x)), big.mark = ",", scientific = FALSE, trim = TRUE)
}
fmt_pct <- function(x, digits = 1) {
  ifelse(is.na(x) | is.nan(x) | is.infinite(x), "NA", sprintf(paste0("%.", digits, "f%%"), x))
}
safe_div <- function(num, den) ifelse(is.na(den) | den == 0, NA_real_, num / den)
co_chr <- function(x, default = NA_character_) ifelse(is.na(x) | x == "", default, as.character(x))
has_cols <- function(x, cols) !is.null(x) && all(cols %in% names(x))
changed_flag <- function(x) if ("changed" %in% names(x)) x$changed %in% TRUE else rep(FALSE, nrow(x))
first_existing <- function(paths) {
  paths <- unique(paths[nzchar(paths)])
  hit <- paths[file.exists(paths)]
  if (length(hit) == 0) NA_character_ else normalizePath(hit[[1]], mustWork = TRUE)
}

find_project_root <- function() {
  starts <- unique(c(getwd(), dirname(normalizePath(knitr::current_input() %||% ".", mustWork = FALSE))))
  starts <- starts[nzchar(starts)]
  for (start in starts) {
    cur <- normalizePath(start, mustWork = FALSE)
    for (i in seq_len(8)) {
      if (file.exists(file.path(cur, "app.R")) && dir.exists(file.path(cur, "data"))) return(cur)
      parent <- dirname(cur)
      if (identical(parent, cur)) break
      cur <- parent
    }
  }
  normalizePath(getwd(), mustWork = TRUE)
}

`%||%` <- function(x, y) if (is.null(x) || length(x) == 0 || is.na(x) || !nzchar(x)) y else x

project_root <- find_project_root()
rmd_dir <- dirname(normalizePath(knitr::current_input() %||% "rmarkdown/preprocessing.Rmd", mustWork = FALSE))

resolve_file <- function(path, fallback = NULL) {
  if (is.null(path) || is.na(path) || !nzchar(path)) path <- fallback
  first_existing(c(
    path,
    file.path(getwd(), path),
    file.path(rmd_dir, path),
    file.path(project_root, path),
    file.path(project_root, basename(path)),
    fallback,
    file.path(project_root, fallback %||% "")
  ))
}

resolve_dir <- function(path, fallback = "data") {
  candidates <- unique(c(
    path,
    file.path(getwd(), path),
    file.path(rmd_dir, path),
    file.path(project_root, path),
    file.path(project_root, basename(path)),
    file.path(project_root, fallback)
  ))
  hit <- candidates[dir.exists(candidates)]
  if (length(hit) == 0) NA_character_ else normalizePath(hit[[1]], mustWork = TRUE)
}

cache_file <- resolve_file(params$cache_path, "trials_cache.rds")
log_dir <- resolve_dir(params$log_dir, "data")
db_path <- resolve_file("data/trials.sqlite", file.path("data", "trials.sqlite"))

if (is.na(cache_file)) stop("Cache not found. Tried param cache_path and project-root trials_cache.rds.")

t0 <- proc.time()
df <- readRDS(cache_file)
load_sec <- round((proc.time() - t0)[["elapsed"]], 1)
df <- tibble::as_tibble(df)

required_defaults <- list(
  DIMP_inn_name = NA_character_,
  transition_eudract_number = NA_character_,
  trial_identifiers = if ("CT_number" %in% names(df)) df$CT_number else NA_character_,
  analysis_register = if ("register" %in% names(df)) df$register else NA_character_,
  sponsor_label = if ("sponsor_name" %in% names(df)) df$sponsor_name else NA_character_,
  canonical_trial_id = if ("_id" %in% names(df)) df$`_id` else NA_character_
)
for (nm in names(required_defaults)) {
  if (!nm %in% names(df)) df[[nm]] <- required_defaults[[nm]]
}

read_log <- function(name) {
  p <- file.path(log_dir, paste0(name, "_normalisation_log.csv"))
  if (!is.na(log_dir) && file.exists(p)) readr::read_csv(p, show_col_types = FALSE) else NULL
}
read_optional_csv <- function(...) {
  p <- file.path(...)
  if (file.exists(p)) readr::read_csv(p, show_col_types = FALSE) else NULL
}

log_country  <- read_log("country")
log_meddra   <- read_log("meddra_term")
log_organ    <- read_log("organ_class")
log_phase    <- read_log("phase")
log_status_c <- read_log("status_category")
log_status_d <- read_log("status_display")
# Substance v2. The v1 matcher and its config (2_substance_alias_index.csv,
# substance_llm_*.csv, 3_substance_review_queue.csv, the match_score /
# match_source / match_reason columns) were retired with the rewrite.
#
# data/substance_normalisation_log.csv is deliberately NOT read: it still exists
# on disk and its schema is close enough that a naive audit would render a
# plausible page describing a pipeline that no longer produces the app's labels
# — the same trap documented for sponsors above.
sub_log      <- read_optional_csv(log_dir, "substance_normalisation_log_v2.csv")
spon_labels  <- read_optional_csv(log_dir, "trial_sponsor_labels.csv")
sub_labels   <- read_optional_csv(log_dir, "trial_substance_labels.csv")
spon_raw     <- read_optional_csv(log_dir, "trial_sponsors_raw.csv")
sub_raw      <- read_optional_csv(log_dir, "trial_substances_raw.csv")
sub_v2_dir   <- file.path(project_root, "config", "substance_norm_v2")
sub_queue    <- read_optional_csv(sub_v2_dir, "E_review_queue.csv")
sub_registry <- read_optional_csv(sub_v2_dir, "registry.csv")
sub_assign   <- read_optional_csv(sub_v2_dir, "assignments.csv")
sub_rejected <- read_optional_csv(log_dir, "substance_rejected.csv")

# ── Sponsor normalisation v2 artefacts ────────────────────────────────────────
# The deterministic matcher and its config (2_sponsor_alias_index.csv,
# final_sponsor_canonical_map.csv, 3_sponsor_review_queue.csv, the match_score /
# match_source / match_reason columns) were retired with the v2 rewrite; the
# scripts now live in LEGACY/. Nothing below reads them.
#
# Reading the OLD data/sponsor_normalisation_log.csv here would be worse than
# reading nothing: it still exists on disk, and its schema is close enough that
# a naive audit renders a plausible-looking page describing a pipeline that no
# longer produces the app's labels.
v2_dir       <- file.path(project_root, "config", "sponsor_norm_v2")
spon_log     <- read_optional_csv(log_dir, "sponsor_normalisation_log_v2.csv")
spon_blocks  <- read_optional_csv(log_dir, "sponsor_blocks.csv")
spon_registry <- read_optional_csv(v2_dir, "registry.csv")
spon_assign  <- read_optional_csv(v2_dir, "assignments.csv")
spon_queue   <- read_optional_csv(v2_dir, "E_review_queue.csv")
spon_ledger  <- read_optional_csv(project_root, "config", "review_ledger", "review_decisions.csv")

waterfall_plot <- function(wf, title, subtitle = NULL) {
  wf <- wf |>
    mutate(step = factor(step, levels = step))
  ggplot(wf, aes(step, n, fill = type)) +
    geom_col(width = 0.58, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.25, color = DARK) +
    geom_text(
      aes(label = ifelse(!is.na(delta) & delta > 0, paste0("-", fmt_n(delta)), "")),
      y = 0,
      vjust = 1.3,
      size = 3.05,
      color = RED
    ) +
    scale_fill_manual(values = c(start = BLUE, step = RED, end = GREEN, label = TEAL), na.value = GRAY) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0.08, 0.13))) +
    labs(title = title, subtitle = subtitle, x = NULL, y = "Rows / strings remaining") +
    theme_pre() +
    theme(axis.text.x = element_text(angle = 16, hjust = 1))
}

dt_tbl <- function(x, caption = NULL, page = 20, filter = "none") {
  DT::datatable(
    x,
    rownames = FALSE,
    filter = filter,
    options = list(pageLength = page, scrollX = TRUE, autoWidth = TRUE),
    caption = caption
  )
}

nk <- function(x, cap = NULL, cn = colnames(x)) {
  knitr::kable(x, booktabs = TRUE, linesep = "", caption = cap, col.names = cn, escape = FALSE) |>
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left",
      font_size = 13
    )
}

kpi_box <- function(value, label, colour = BLUE) {
  sprintf(
    '<div class="kpi" style="border-left-color:%s"><div class="kpi-value">%s</div><div class="kpi-label">%s</div></div>',
    colour, value, label
  )
}

kpi_row <- function(...) {
  paste0('<div class="kpi-row">', paste0(..., collapse = ""), '</div>')
}

notice <- function(title, text, type = "note") {
  cls <- paste("audit-notice", type)
  sprintf('<div class="%s"><strong>%s</strong><br>%s</div>', cls, title, text)
}

make_title_key <- function(title) {
  title |>
    as.character() |>
    tolower() |>
    str_replace_all("[^a-z0-9 ]", " ") |>
    str_squish() |>
    substr(1, 80)
}

is_euctr_id <- function(x) str_detect(coalesce(as.character(x), ""), "^[0-9]{4}-[0-9]{6}-[0-9]{2}-[A-Z]")
is_ctis_id <- function(x) !is_euctr_id(x)
```

```{css, echo=FALSE}
body {
  color: #2E3440;
}
.main-container {
  max-width: 1180px;
}
h1, h2, h3 {
  letter-spacing: 0;
}
.kpi-row {
  display: flex;
  flex-wrap: wrap;
  gap: 12px;
  margin: 12px 0 18px;
}
.kpi {
  background: #F8FAFC;
  border: 1px solid #E5E9F0;
  border-left: 5px solid #5E81AC;
  border-radius: 6px;
  min-width: 150px;
  padding: 10px 14px;
}
.kpi-value {
  font-size: 1.45rem;
  font-weight: 700;
  line-height: 1.15;
  color: #2E3440;
}
.kpi-label {
  margin-top: 3px;
  font-size: 0.88rem;
  color: #4C566A;
}
.audit-notice {
  border-radius: 6px;
  padding: 12px 14px;
  margin: 12px 0 16px;
  border: 1px solid #D8DEE9;
  background: #F8FAFC;
}
.audit-notice.warn {
  border-color: #EBCB8B;
  background: #FFF8E6;
}
.audit-notice.fail {
  border-color: #BF616A;
  background: #FFF0F0;
}
.section-note {
  color: #4C566A;
  max-width: 880px;
}
.dataTables_wrapper {
  margin-bottom: 22px;
}
```

# 1. Dataset Overview

```{r overview-kpis, results='asis'}
n_tot <- nrow(df)
n_cols <- ncol(df)
reg_counts <- df |> count(register, name = "n")
n_euctr <- reg_counts$n[match("EUCTR", reg_counts$register)] %||% 0
n_ctis  <- reg_counts$n[match("CTIS", reg_counts$register)] %||% 0
n_analysis_euctr <- sum(df$analysis_register == "EUCTR", na.rm = TRUE)
n_analysis_ctis  <- sum(df$analysis_register == "CTIS", na.rm = TRUE)

cat(kpi_row(
  kpi_box(fmt_n(n_tot), "trials in final cache", BLUE),
  kpi_box(fmt_n(n_euctr), "source EUCTR", FROST),
  kpi_box(fmt_n(n_ctis), "source CTIS", TEAL),
  kpi_box(fmt_n(n_cols), "columns", GREEN),
  kpi_box(paste0(load_sec, " s"), "cache load time", PURPLE)
))
```

The report audits the cache at `r cache_file`. Logs are read from `r log_dir`; the SQLite source database is `r ifelse(is.na(db_path), "not available", db_path)`.

## Register Mix

```{r overview-registers, fig.height=3.6}
df |>
  count(register, name = "n") |>
  ggplot(aes(register, n, fill = register)) +
  geom_col(width = 0.58, show.legend = FALSE) +
  geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.5, color = DARK) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.14))) +
  labs(title = "Trials by source register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r overview-analysis-registers, fig.height=3.6}
df |>
  count(analysis_register, name = "n") |>
  mutate(analysis_register = coalesce(analysis_register, "Missing")) |>
  ggplot(aes(analysis_register, n, fill = analysis_register)) +
  geom_col(width = 0.58, show.legend = FALSE) +
  geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.5, color = DARK) +
  scale_fill_manual(values = c(REG_COLS, Missing = GRAY), na.value = GRAY) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.14))) +
  labs(title = "Trials by analysis register", subtitle = "Migrated CTIS records with EudraCT transition numbers are analysed as EUCTR-era.", x = NULL, y = "Trials") +
  theme_pre()
```

## Key Field Completeness

```{r overview-key-completeness}
key_cols <- c(
  "_id", "CT_number", "Full_title", "register", "analysis_register",
  "transition_eudract_number", "canonical_trial_id", "Member_state",
  "age_group", "sponsor_label", "sponsor_type", "phase", "status",
  "submission_date_parsed", "decision_date", "days_to_decision",
  "has_PIP", "has_results", "is_orphan", "substance_label"
)
key_cols <- intersect(key_cols, names(df))

tibble(
  Column = key_cols,
  Type = vapply(df[key_cols], function(x) class(x)[1], character(1)),
  `Non-missing` = vapply(df[key_cols], function(x) sum(!(is.na(x) | (is.character(x) & x == ""))), integer(1)),
  `Complete (%)` = round(`Non-missing` / nrow(df) * 100, 1)
) |>
  arrange(`Complete (%)`) |>
  dt_tbl(caption = "Completeness of key fields in the final cache", page = 25)
```

# 2. Deduplication Pipeline

The dedup audit compares raw SQLite row identifiers with the final cache. EUCTR country variants and CTIS amendments are counted directly from raw IDs; cross-register and final removals are the remaining difference between the deduplicated raw universe and the cache.

```{r dedup-compute}
db_available <- !is.na(db_path) && file.exists(db_path) && has_pkg("DBI") && has_pkg("RSQLite")

if (db_available) {
  con <- DBI::dbConnect(RSQLite::SQLite(), db_path)
  all_ids <- DBI::dbGetQuery(con, 'SELECT "_id" FROM trials')[["_id"]]
  trans_df <- DBI::dbGetQuery(con, '
    SELECT "_id",
      json_extract(json, "$.x5_trial_status") AS trial_status,
      json_extract(json, "$.p_end_of_trial_status") AS end_status,
      json_extract(json, "$.a3_full_title_of_the_trial") AS title
    FROM trials
    WHERE lower(json_extract(json, "$.x5_trial_status")) LIKE "%transition%"
       OR lower(json_extract(json, "$.p_end_of_trial_status")) LIKE "%transition%"
  ')
  DBI::dbDisconnect(con)
} else {
  all_ids <- character(0)
  trans_df <- tibble(`_id` = character(), trial_status = character(), end_status = character(), title = character())
}

euctr_all <- all_ids[is_euctr_id(all_ids)]
ctis_all <- all_ids[is_ctis_id(all_ids)]
euctr_base_all <- str_replace(euctr_all, "-[A-Z]{2,3}$", "")
ctis_base_all <- str_replace(ctis_all, "-[0-9]{2}$", "")

n_raw <- length(all_ids)
n_euctr_base <- n_distinct(euctr_base_all)
n_ctis_base <- n_distinct(ctis_base_all)
n_ctis_amended <- if (length(ctis_base_all) > 0) sum(table(ctis_base_all) > 1) else 0
n_step1 <- n_euctr_base + length(ctis_all)
n_step2 <- n_euctr_base + n_ctis_base
n_final <- nrow(df)
n_drop_country <- n_raw - n_step1
n_drop_ctis <- n_step1 - n_step2
n_drop_final <- n_step2 - n_final
n_migrated <- sum(df$register == "CTIS" & df$analysis_register == "EUCTR", na.rm = TRUE)

wf <- tibble(
  step = factor(
    c("Raw DB rows", "After EUCTR country collapse", "After CTIS latest amendment", "After cross-register/final dedup", "Final cache"),
    levels = c("Raw DB rows", "After EUCTR country collapse", "After CTIS latest amendment", "After cross-register/final dedup", "Final cache")
  ),
  remaining = c(n_raw, n_step1, n_step2, n_final, n_final),
  removed = c(NA_integer_, n_drop_country, n_drop_ctis, n_drop_final, 0L),
  type = c("start", "step", "step", "step", "end")
)

wf_ok <- db_available && n_raw > 0 && all(!is.na(wf$remaining)) && all(diff(wf$remaining) <= 0) && all(wf$remaining >= 0)
```

## Waterfall

```{r dedup-waterfall, fig.height=4.6, results='asis'}
if (!db_available) {
  cat(notice("SQLite database unavailable", "The waterfall needs data/trials.sqlite. The final cache can still be audited, but raw-row dedup counts are skipped.", "warn"))
} else if (!wf_ok) {
  cat(notice("Waterfall counts are inconsistent", "The report will show the raw stage table instead of a plot because counts are not monotonic.", "fail"))
  print(nk(wf, cap = "Raw waterfall stage counts"))
} else {
  ggplot(wf, aes(step, remaining, fill = type)) +
    geom_col(width = 0.58, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(remaining)), vjust = -0.35, size = 3.4, color = DARK) +
    geom_text(
      data = wf |> filter(!is.na(removed) & removed > 0),
      aes(label = paste0("-", fmt_n(removed))),
      y = 0,
      vjust = 1.3,
      size = 3.2,
      color = RED,
      inherit.aes = TRUE
    ) +
    scale_fill_manual(values = c(start = BLUE, step = RED, end = GREEN)) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0.08, 0.12))) +
    labs(
      title = "Deduplication waterfall: rows remaining after each step",
      subtitle = "Red labels at the baseline show rows removed at that stage.",
      x = NULL,
      y = "Rows remaining"
    ) +
    theme_pre() +
    theme(axis.text.x = element_text(angle = 18, hjust = 1))
}
```

```{r dedup-table}
dedup_table <- tibble(
  Step = c("Raw DB", "EUCTR country collapse", "CTIS latest amendment", "Cross-register / final dedup", "Analysis label"),
  Logic = c(
    "All raw records fetched from SQLite",
    "Strip EUCTR member-state suffix and keep one base trial",
    "Strip CTIS amendment suffix and keep the latest version",
    "Prefer CTIS over duplicate EUCTR records and apply final cache exclusions",
    "Set analysis_register to EUCTR for migrated CTIS records"
  ),
  `Rows removed` = c("NA", fmt_n(n_drop_country), fmt_n(n_drop_ctis), fmt_n(n_drop_final), "0"),
  `Rows remaining` = c(fmt_n(n_raw), fmt_n(n_step1), fmt_n(n_step2), fmt_n(n_final), fmt_n(n_final))
)
nk(dedup_table, cap = "Deduplication stages")
```

## EUCTR Country Variants

```{r dedup-euctr-table}
if (!db_available) {
  cat("Raw DB not available.")
} else {
  cache_euctr_base <- df |>
    filter(is_euctr_id(`_id`)) |>
    transmute(base = str_replace(`_id`, "-[A-Z]{2,3}$", "")) |>
    pull(base)

  euctr_variant_tbl <- tibble(raw_id = euctr_all) |>
    mutate(base = str_replace(raw_id, "-[A-Z]{2,3}$", ""),
           country = str_extract(raw_id, "[A-Z]{2,3}$")) |>
    count(base, country, name = "n") |>
    group_by(base) |>
    summarise(
      `Countries in raw DB` = n_distinct(country),
      `Country variants` = paste(sort(unique(country)), collapse = ", "),
      `Kept in cache` = base %in% cache_euctr_base,
      .groups = "drop"
    ) |>
    filter(`Kept in cache`) |>
    arrange(desc(`Countries in raw DB`)) |>
    slice_head(n = 20) |>
    rename(`Base trial ID` = base)
  dt_tbl(euctr_variant_tbl, caption = "EUCTR base trials with the most country-specific raw rows", page = 20)
}
```

```{r dedup-euctr-dist, fig.height=3.5}
if (db_available && length(euctr_all) > 0) {
  tibble(raw_id = euctr_all) |>
    mutate(base = str_replace(raw_id, "-[A-Z]{2,3}$", "")) |>
    count(base, name = "country_rows") |>
    count(country_rows, name = "trials") |>
    filter(country_rows <= 30) |>
    ggplot(aes(country_rows, trials)) +
    geom_col(fill = BLUE, width = 0.8) +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = pretty_breaks()) +
    labs(title = "Distribution of EUCTR country rows per trial", x = "Country-specific rows", y = "Trials") +
    theme_pre()
}
```

## CTIS Amendment Versions

```{r dedup-ctis-table}
if (!db_available) {
  cat("Raw DB not available.")
} else {
  ctis_version_tbl <- tibble(raw_id = ctis_all) |>
    mutate(base = str_replace(raw_id, "-[0-9]{2}$", ""),
           version = suppressWarnings(as.integer(str_extract(raw_id, "[0-9]{2}$")))) |>
    group_by(base) |>
    filter(n() > 1) |>
    summarise(
      `Versions in DB` = n(),
      `Latest version` = max(version, na.rm = TRUE),
      `Latest raw ID` = raw_id[which.max(version)],
      `Latest kept in cache` = `Latest raw ID` %in% df$`_id`,
      `All versions` = paste(sort(raw_id), collapse = " -> "),
      .groups = "drop"
    ) |>
    arrange(desc(`Versions in DB`)) |>
    slice_head(n = 25) |>
    rename(`Base CTIS ID` = base)
  dt_tbl(ctis_version_tbl, caption = paste0("CTIS trials with multiple amendment versions (", fmt_n(n_ctis_amended), " base trials)"), page = 25)
}
```

```{r dedup-ctis-dist, fig.height=3.4}
if (db_available && length(ctis_all) > 0) {
  tibble(raw_id = ctis_all) |>
    mutate(base = str_replace(raw_id, "-[0-9]{2}$", "")) |>
    count(base, name = "versions") |>
    count(versions, name = "trials") |>
    ggplot(aes(versions, trials)) +
    geom_col(fill = TEAL, width = 0.7) +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = pretty_breaks()) +
    labs(title = "Distribution of CTIS amendment versions", x = "Versions in raw DB", y = "Base trials") +
    theme_pre()
}
```

## Cross-Register Checks

```{r dedup-cross-register}
ctis_by_tk <- df |>
  filter(register == "CTIS") |>
  mutate(title_key_check = make_title_key(Full_title)) |>
  filter(!is.na(title_key_check), nchar(title_key_check) >= 20) |>
  select(ctis_id = `_id`, title_key_check, ctis_title = Full_title, ctis_status = status)

euctr_by_tk <- df |>
  filter(register == "EUCTR") |>
  mutate(title_key_check = make_title_key(Full_title)) |>
  filter(!is.na(title_key_check), nchar(title_key_check) >= 20) |>
  select(euctr_id = `_id`, title_key_check, euctr_title = Full_title, euctr_status = status)

cross_reg_tbl <- ctis_by_tk |>
  inner_join(euctr_by_tk, by = "title_key_check") |>
  filter(ctis_id != euctr_id) |>
  select(
    `EUCTR ID` = euctr_id,
    `CTIS ID` = ctis_id,
    `Shared title key` = title_key_check,
    `EUCTR status` = euctr_status,
    `CTIS status` = ctis_status
  ) |>
  slice_head(n = 100)

trans_matched <- tibble()
if (db_available && nrow(trans_df) > 0) {
  ctis_transition <- df |>
    filter(register == "CTIS", !is.na(transition_eudract_number), transition_eudract_number != "") |>
    separate_rows(transition_eudract_number, sep = " / ") |>
    mutate(transition_eudract_number = str_trim(transition_eudract_number)) |>
    distinct(transition_eudract_number, ctis_id = `_id`, ctis_title = Full_title, ctis_status = status)

  trans_matched <- trans_df |>
    as_tibble() |>
    rename(euctr_id = `_id`) |>
    mutate(
      euctr_base = str_replace(euctr_id, "-[A-Z]{2,3}$", ""),
      title_key_check = make_title_key(title)
    ) |>
    distinct(euctr_base, title_key_check, trial_status, end_status) |>
    left_join(ctis_transition, by = c("euctr_base" = "transition_eudract_number")) |>
    left_join(ctis_by_tk, by = "title_key_check") |>
    mutate(
      matched_ctis_id = coalesce(ctis_id.x, ctis_id.y),
      match_method = case_when(
        !is.na(ctis_id.x) ~ "transition EudraCT number",
        !is.na(ctis_id.y) ~ "title key",
        TRUE ~ NA_character_
      )
    ) |>
    distinct(euctr_base, .keep_all = TRUE)
}
n_resid_dupes <- nrow(cross_reg_tbl)
n_unmatched_transitioned <- if (nrow(trans_matched) > 0) sum(is.na(trans_matched$matched_ctis_id)) else NA_integer_
```

```{r dedup-cross-table}
if (nrow(cross_reg_tbl) > 0) {
  dt_tbl(cross_reg_tbl, caption = "Potential residual cross-register duplicates in the final cache", page = 25)
} else {
  cat("No residual EUCTR/CTIS title-key duplicates were detected in the final cache.")
}
```

```{r dedup-transition-table}
if (!db_available) {
  cat("Raw DB not available.")
} else if (nrow(trans_matched) == 0) {
  cat("No transitioned EUCTR rows were found in the raw DB status fields.")
} else {
  trans_matched |>
    mutate(
      `EUCTR status` = coalesce(end_status, trial_status),
      `CTIS match` = coalesce(matched_ctis_id, "No CTIS match found")
    ) |>
    select(`EUCTR base ID` = euctr_base, `EUCTR status`, `CTIS match`, `Match method` = match_method) |>
    arrange(`CTIS match` == "No CTIS match found", `EUCTR base ID`) |>
    dt_tbl(caption = "EUCTR rows marked as transitioned and their CTIS match status", page = 25)
}
```

# 3. Register, Age, Phase, Status, and Dates

## Age Group

```{r age-kpis, results='asis'}
age_counts <- df |> count(age_group, name = "n") |> mutate(pct = n / sum(n) * 100)
n_paediatric <- sum(df$age_group %in% c("Paediatric", "Paediatric & Adult"), na.rm = TRUE)
n_adult <- sum(df$age_group %in% c("Adult", "Paediatric & Adult"), na.rm = TRUE)
cat(kpi_row(
  kpi_box(fmt_n(n_paediatric), "eligible under 18", BLUE),
  kpi_box(fmt_n(n_adult), "eligible 18 or older", GREEN),
  kpi_box(fmt_n(sum(df$age_group == "Unknown", na.rm = TRUE)), "unknown age group", RED)
))
```

```{r age-chart, fig.height=3.8}
df |>
  count(register, age_group, name = "n") |>
  mutate(age_group = coalesce(age_group, "Unknown")) |>
  ggplot(aes(register, n, fill = age_group)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = ifelse(n > max(n) * 0.05, fmt_n(n), "")), position = position_stack(vjust = 0.5), size = 3, color = "white") +
  scale_fill_manual(values = c(Paediatric = BLUE, Adult = ORANGE, `Paediatric & Adult` = GREEN, Unknown = GRAY), na.value = GRAY) +
  scale_y_continuous(labels = comma) +
  labs(title = "Age-group classification by register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r age-table}
age_counts |>
  mutate(`Trials` = fmt_n(n), `% of cache` = fmt_pct(pct)) |>
  select(`Age group` = age_group, Trials, `% of cache`) |>
  nk(cap = "Adult / paediatric classification")
```

## Phase

```{r phase-chart, fig.height=4}
df |>
  mutate(phase = coalesce(phase, "Unknown")) |>
  count(register, phase, name = "n") |>
  ggplot(aes(phase, n, fill = register)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.62) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma) +
  labs(title = "Trial phase by register", x = NULL, y = "Trials") +
  theme_pre() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r phase-log-table}
if (!is.null(log_phase) && has_cols(log_phase, c("raw", "normalised", "n_trials"))) {
  log_phase |>
    arrange(desc(n_trials)) |>
    select(Register = register, Raw = raw, Normalised = normalised, `N trials` = n_trials) |>
    slice_head(n = 30) |>
    dt_tbl(caption = "Most common raw phase values and parsed labels", page = 30)
} else {
  cat("Phase normalisation log not available.")
}
```

## Status

```{r status-chart, fig.height=3.8}
df |>
  mutate(status = coalesce(status, "Other")) |>
  count(register, status, name = "n") |>
  ggplot(aes(register, n, fill = status)) +
  geom_col(width = 0.62) +
  scale_fill_manual(values = STATUS_COLS, na.value = GRAY) +
  scale_y_continuous(labels = comma) +
  labs(title = "Derived trial status by register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r status-log-table}
if (!is.null(log_status_c) && has_cols(log_status_c, c("raw", "normalised", "n_trials"))) {
  log_status_c |>
    filter(register == "CTIS" | changed_flag(log_status_c)) |>
    arrange(register, desc(n_trials)) |>
    select(Register = register, Raw = raw, Normalised = normalised, `N trials` = n_trials) |>
    slice_head(n = 40) |>
    dt_tbl(caption = "Status category mapping audit", page = 25)
} else {
  cat("Status normalisation log not available.")
}
```

## Dates and Decision Lag

```{r date-kpis, results='asis'}
n_decision_missing <- sum(is.na(df$decision_date))
n_completed_no_decision <- sum(df$status == "Completed" & is.na(df$decision_date), na.rm = TRUE)
median_days <- median(df$days_to_decision, na.rm = TRUE)
cat(kpi_row(
  kpi_box(fmt_n(sum(!is.na(df$submission_date_parsed))), "parsed submission dates", BLUE),
  kpi_box(fmt_n(sum(!is.na(df$decision_date))), "decision dates", GREEN),
  kpi_box(fmt_n(n_completed_no_decision), "completed with no decision date", RED),
  kpi_box(fmt_n(median_days), "median days to decision", PURPLE)
))
```

```{r year-chart, fig.height=4.2}
df |>
  filter(!is.na(year), year >= 1990, year <= year(Sys.Date())) |>
  count(year, analysis_register, name = "n") |>
  ggplot(aes(year, n, fill = analysis_register)) +
  geom_col(width = 0.9) +
  scale_fill_manual(values = REG_COLS, na.value = GRAY) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Trials by submission/analysis year", x = NULL, y = "Trials") +
  theme_pre()
```

```{r days-chart, fig.height=4}
df |>
  filter(!is.na(days_to_decision), days_to_decision >= 0, days_to_decision <= 3650) |>
  mutate(log_days = log10(days_to_decision + 1)) |>
  ggplot(aes(log_days, fill = register)) +
  geom_histogram(bins = 45, alpha = 0.72, position = "identity") +
  scale_fill_manual(values = REG_COLS) +
  scale_x_continuous(labels = function(x) fmt_n(10^x - 1)) +
  labs(title = "Days from submission to decision", subtitle = "X-axis is log-scaled for readability.", x = "Days to decision", y = "Trials") +
  theme_pre()
```

# 4. Field Normalisation Audits

## Country

```{r country-kpis, results='asis'}
if (!is.null(log_country) && has_cols(log_country, c("raw", "normalised", "n_trials"))) {
  n_country_rows <- sum(log_country$n_trials, na.rm = TRUE)
  n_country_raw <- n_distinct(log_country$raw)
  n_country_norm <- n_distinct(log_country$normalised, na.rm = TRUE)
  n_country_changed <- sum(log_country$n_trials[changed_flag(log_country)], na.rm = TRUE)
  cat(kpi_row(
    kpi_box(fmt_n(n_country_rows), "country entries", BLUE),
    kpi_box(fmt_n(n_country_raw), "unique raw values", FROST),
    kpi_box(fmt_n(n_country_norm), "unique normalised values", GREEN),
    kpi_box(fmt_n(n_country_changed), "entries changed", ORANGE)
  ))
} else {
  cat(notice("Country log unavailable", "No country normalisation log was found for this render.", "warn"))
}
```

```{r country-chart, fig.height=5}
if ("Member_state" %in% names(df)) {
  df |>
    filter(!is.na(Member_state), Member_state != "") |>
    separate_rows(Member_state, sep = " / ") |>
    mutate(Member_state = str_squish(Member_state)) |>
    filter(Member_state != "") |>
    count(Member_state, sort = TRUE, name = "n") |>
    slice_head(n = 20) |>
    mutate(Member_state = reorder(Member_state, n)) |>
    ggplot(aes(n, Member_state, fill = n)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    scale_fill_gradient(low = FROST, high = BLUE) +
    labs(title = "Top countries in final cache", x = "Trials", y = NULL) +
    theme_pre()
}
```

## MedDRA Terms

```{r meddra-kpis, results='asis'}
if (!is.null(log_meddra) && has_cols(log_meddra, c("raw", "normalised", "n_trials"))) {
  cat(kpi_row(
    kpi_box(fmt_n(sum(log_meddra$n_trials, na.rm = TRUE)), "MedDRA term entries", BLUE),
    kpi_box(fmt_n(n_distinct(log_meddra$raw)), "unique raw values", FROST),
    kpi_box(fmt_n(n_distinct(log_meddra$normalised, na.rm = TRUE)), "unique normalised values", GREEN),
    kpi_box(fmt_n(sum(log_meddra$n_trials[changed_flag(log_meddra)], na.rm = TRUE)), "entries changed", ORANGE)
  ))
} else {
  cat(notice("MedDRA term log unavailable", "No MedDRA term normalisation log was found for this render.", "warn"))
}
```

```{r meddra-chart, fig.height=5}
if ("MEDDRA_term" %in% names(df)) {
  df |>
    filter(!is.na(MEDDRA_term), MEDDRA_term != "") |>
    separate_rows(MEDDRA_term, sep = " / ") |>
    mutate(MEDDRA_term = str_squish(MEDDRA_term)) |>
    count(MEDDRA_term, sort = TRUE, name = "n") |>
    slice_head(n = 20) |>
    mutate(MEDDRA_term = reorder(MEDDRA_term, n)) |>
    ggplot(aes(n, MEDDRA_term, fill = n)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    scale_fill_gradient(low = FROST, high = BLUE) +
    labs(title = "Most common MedDRA terms", x = "Trials", y = NULL) +
    theme_pre()
}
```

## Organ Class

```{r organ-kpis, results='asis'}
if (!is.null(log_organ) && has_cols(log_organ, c("raw", "normalised", "n_trials"))) {
  cat(kpi_row(
    kpi_box(fmt_n(sum(log_organ$n_trials, na.rm = TRUE)), "organ-class entries", BLUE),
    kpi_box(fmt_n(n_distinct(log_organ$raw)), "unique raw values", FROST),
    kpi_box(fmt_n(n_distinct(log_organ$normalised, na.rm = TRUE)), "unique normalised values", GREEN),
    kpi_box(fmt_n(sum(log_organ$n_trials[changed_flag(log_organ)], na.rm = TRUE)), "entries changed", ORANGE)
  ))
} else {
  cat(notice("Organ-class log unavailable", "No organ-class normalisation log was found for this render.", "warn"))
}
```

```{r organ-chart, fig.height=5}
if ("MEDDRA_organ_class" %in% names(df)) {
  df |>
    filter(!is.na(MEDDRA_organ_class), MEDDRA_organ_class != "") |>
    separate_rows(MEDDRA_organ_class, sep = " / ") |>
    mutate(MEDDRA_organ_class = str_squish(MEDDRA_organ_class)) |>
    count(MEDDRA_organ_class, sort = TRUE, name = "n") |>
    slice_head(n = 18) |>
    mutate(MEDDRA_organ_class = reorder(MEDDRA_organ_class, n)) |>
    ggplot(aes(n, MEDDRA_organ_class, fill = n)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    scale_fill_gradient(low = FROST, high = BLUE) +
    labs(title = "Most common MedDRA organ classes", x = "Trials", y = NULL) +
    theme_pre()
}
```

## Sponsor Names

Sponsor labels come from a **model-built canonical registry** (v2), which replaced
the deterministic matcher and its ~16,500 committed alias rows. Raw strings are
blocked offline, each block is named once with every variant visible, leftover
strings are assigned by picking from a candidate list, duplicate entries are
merged, and one script emits the app-facing labels.

The label the app displays is resolved by an explicit precedence chain, and the
top of it is not the pipeline:

| Rank | Source | Where |
|---|---|---|
| 1 | **Human curation** | `config/review_ledger/review_decisions.csv` |
| 2 | Pipeline canonical | `data/trial_sponsor_labels.csv` |
| 3 | Raw sponsor string | fallback when neither exists |

The ledger is read on every app load and never enters `trials_cache.rds`, so a
curator's decision is live on the next start without a cache rebuild.

```{r sponsor-diagram, out.width="100%", fig.cap="Sponsor normalisation v2. Blue is deterministic, violet is model-driven, green is what the app reads."}
# Regenerate after editing the .dot:
#   dot -Tpng -Gdpi=110 rmarkdown/sponsor_normalisation_pipeline.dot \
#     -o rmarkdown/sponsor_normalisation_pipeline.png
spon_diagram <- first_existing(c(
  "sponsor_normalisation_pipeline.png",
  file.path(project_root, "rmarkdown", "sponsor_normalisation_pipeline.png")
))
# An illustration must never take down the report. On the server this chunk
# aborted the whole render with "Cannot find the file(s)" — knitr resolves
# include_graphics against the render directory, which is not where the .png
# lives when rebuild_cache.R renders to www/. Everything below this point is the
# actual audit; losing it to a missing picture is a bad trade.
if (!is.na(spon_diagram)) {
  tryCatch(knitr::include_graphics(spon_diagram, rel_path = FALSE),
           error = function(e) {
             message("pipeline diagram skipped: ", conditionMessage(e))
             invisible(NULL)
           })
}
```

```{r sponsor-compute}
spon_has_v2 <- !is.null(spon_log) && has_cols(spon_log, c("raw_sponsor", "sponsor_clean", "n_trials"))

spon_live <- if (!is.null(spon_registry) && "canonical" %in% names(spon_registry)) {
  if ("merged_into" %in% names(spon_registry)) {
    spon_registry |> filter(is.na(merged_into) | merged_into == "")
  } else spon_registry
} else NULL

# Trial rows, not distinct strings: a canonical attached to one rare string and
# one attached to 1,600 trials are not the same thing, and every chart below is
# weighted by impact for that reason.
spon_by_canon <- if (spon_has_v2) {
  spon_log |>
    filter(!is.na(sponsor_clean), sponsor_clean != "") |>
    group_by(sponsor_clean) |>
    summarise(n_trials = sum(n_trials, na.rm = TRUE), .groups = "drop") |>
    arrange(desc(n_trials))
} else tibble()

spon_n_strings   <- if (spon_has_v2) nrow(spon_log) else NA_integer_
spon_n_trials    <- if (spon_has_v2) sum(spon_log$n_trials, na.rm = TRUE) else NA_integer_
spon_n_canon     <- if (nrow(spon_by_canon)) nrow(spon_by_canon) else NA_integer_
spon_n_live      <- if (!is.null(spon_live)) nrow(spon_live) else NA_integer_
spon_n_merged    <- if (!is.null(spon_registry) && "merged_into" %in% names(spon_registry)) {
  sum(!is.na(spon_registry$merged_into) & spon_registry$merged_into != "")
} else NA_integer_
spon_queue_rows  <- if (!is.null(spon_queue)) nrow(spon_queue) else NA_integer_
spon_queue_trials <- if (!is.null(spon_queue) && "n_trials" %in% names(spon_queue)) {
  sum(spon_queue$n_trials, na.rm = TRUE)
} else NA_integer_
spon_unlabelled  <- if (spon_has_v2) sum(is.na(spon_log$sponsor_clean) | spon_log$sponsor_clean == "") else NA_integer_

# Human decisions are the top of the precedence chain, so their absence is a
# finding in its own right, not an empty section.
spon_human_n <- if (!is.null(spon_ledger) && "domain" %in% names(spon_ledger)) {
  sum(spon_ledger$domain %in% "sponsor", na.rm = TRUE)
} else 0L
```

```{r sponsor-kpis, results='asis'}
if (spon_has_v2) {
  cat(kpi_row(
    kpi_box(fmt_n(spon_n_strings), "distinct raw strings", BLUE),
    kpi_box(fmt_n(spon_n_trials), "trial rows labelled", FROST),
    kpi_box(fmt_n(spon_n_canon), "canonical sponsors", GREEN),
    kpi_box(fmt_n(spon_queue_rows), "awaiting human review", ORANGE)
  ))
} else {
  cat(notice(
    "Sponsor v2 artefacts unavailable",
    paste("No <code>data/sponsor_normalisation_log_v2.csv</code> was found.",
          "Run <code>E_emit.R</code> to regenerate it."),
    "warn"))
}
```

```{r sponsor-coverage, results='asis'}
if (spon_has_v2) {
  collapse_ratio <- safe_div(spon_n_strings, spon_n_canon)
  msg <- sprintf(
    paste0("<strong>%s</strong> distinct raw strings collapse to <strong>%s</strong> canonical ",
           "sponsors — a %.1fx reduction. %s string(s) carry no canonical."),
    fmt_n(spon_n_strings), fmt_n(spon_n_canon), collapse_ratio, fmt_n(spon_unlabelled))
  if (!is.na(spon_n_merged) && spon_n_merged > 0) {
    msg <- paste0(msg, sprintf(" Consolidation merged a further <strong>%s</strong> duplicate entries away.",
                               fmt_n(spon_n_merged)))
  }
  cat(notice("Collapse", msg, if (isTRUE(spon_unlabelled > 0)) "warn" else "note"))
}
```

### Who sponsors these trials

```{r sponsor-top, fig.height=5.2}
if (nrow(spon_by_canon) > 0) {
  spon_by_canon |>
    slice_head(n = 15) |>
    mutate(sponsor_clean = reorder(sponsor_clean, n_trials)) |>
    ggplot(aes(n_trials, sponsor_clean)) +
    geom_col(fill = BLUE, width = 0.68) +
    geom_text(aes(label = fmt_n(n_trials)), hjust = -0.12, size = 3.1, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.14))) +
    labs(
      title = "Top 15 sponsors by trial rows",
      subtitle = "One canonical per organisation — national subsidiaries and legal variants are folded into the brand.",
      x = "Trial rows", y = NULL) +
    theme_pre()
}
```

```{r sponsor-concentration, fig.height=3.9}
# The long tail is the point. A bar chart of the top N hides that most
# canonicals are attached to a single trial, which is what actually determines
# how much curation effort the registry will ever need.
if (nrow(spon_by_canon) > 20) {
  conc <- spon_by_canon |>
    mutate(rank = row_number(), cum_share = cumsum(n_trials) / sum(n_trials))
  marks <- conc |> filter(rank %in% c(10, 100, 1000))
  ggplot(conc, aes(rank, cum_share)) +
    geom_line(color = BLUE, linewidth = 0.9) +
    geom_point(data = marks, color = ORANGE, size = 2.2) +
    geom_text(data = marks, aes(label = sprintf("top %s: %s", fmt_n(rank), fmt_pct(100 * cum_share, 0))),
              hjust = -0.08, vjust = 1.4, size = 3.1, color = GRAY) +
    scale_x_log10(labels = comma) +
    scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
    labs(
      title = "Trial coverage by sponsor rank",
      subtitle = sprintf("%s of %s canonicals carry exactly one trial row.",
                         fmt_n(sum(spon_by_canon$n_trials == 1)), fmt_n(nrow(spon_by_canon))),
      x = "Sponsor rank (log scale)", y = "Cumulative share of trial rows") +
    theme_pre()
}
```

```{r sponsor-type, fig.height=3.6}
if (spon_has_v2 && "sponsor_type" %in% names(spon_log)) {
  spon_log |>
    mutate(sponsor_type = coalesce(as.character(sponsor_type), "unknown")) |>
    count(sponsor_type, wt = n_trials, name = "n") |>
    filter(n > 0) |>
    mutate(sponsor_type = reorder(sponsor_type, n)) |>
    ggplot(aes(n, sponsor_type)) +
    geom_col(fill = TEAL, width = 0.66) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.12, size = 3.1, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    labs(
      title = "Trial rows by sponsor type",
      subtitle = "Assigned during minting; the split between commercial and academic sponsorship is a property of the corpus, not of the pipeline.",
      x = "Trial rows", y = NULL) +
    theme_pre()
}
```

### How confident is the registry

```{r sponsor-confidence, fig.height=3.6}
# Weighted by trial rows, because a shaky label on a 500-trial sponsor matters
# and a shaky label on a one-off does not.
if (spon_has_v2 && "confidence" %in% names(spon_log)) {
  bands <- spon_log |>
    mutate(band = cut(suppressWarnings(as.numeric(confidence)),
                      breaks = c(0, 0.5, 0.7, 0.75, 0.9, 1),
                      labels = c("<= 0.50", "0.50-0.70", "0.70-0.75", "0.75-0.90", "> 0.90"),
                      include.lowest = TRUE)) |>
    count(band, wt = n_trials, name = "n") |>
    filter(!is.na(band))
  if (nrow(bands) > 0) {
    ggplot(bands, aes(band, n, fill = band)) +
      geom_col(width = 0.62, show.legend = FALSE) +
      geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.1, color = DARK) +
      scale_fill_manual(values = c("<= 0.50" = RED, "0.50-0.70" = ORANGE, "0.70-0.75" = YELLOW,
                                   "0.75-0.90" = FROST, "> 0.90" = GREEN)) +
      scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
      labs(
        title = "Trial rows by assignment confidence",
        subtitle = "Rows below 0.75 are routed to the review queue; high-impact rows below 0.90 are too.",
        x = "Model confidence", y = "Trial rows") +
      theme_pre()
  }
}
```

### What still needs a human

```{r sponsor-review-note, results='asis'}
if (!is.null(spon_queue)) {
  share <- safe_div(spon_queue_trials, spon_n_trials) * 100
  cat(notice(
    "Review queue",
    sprintf(paste0("<strong>%s</strong> raw strings carrying <strong>%s</strong> trial rows (%s of all rows) ",
                   "are queued for confirmation. These are proposals, not decisions — nothing here changes a ",
                   "displayed label until a reviewer acts on it."),
            fmt_n(spon_queue_rows), fmt_n(spon_queue_trials), fmt_pct(share)),
    "note"))
  cat(notice(
    "Human decisions recorded",
    if (spon_human_n > 0) {
      sprintf("<strong>%s</strong> sponsor decision(s) in the review ledger. These override the pipeline wherever they exist.", fmt_n(spon_human_n))
    } else {
      paste("<strong>None yet.</strong> Every label on this page is model output.",
            "The precedence chain is wired and tested, so the first curated decision takes effect",
            "on the next app start — but nothing has been human-verified so far.")
    },
    if (spon_human_n > 0) "note" else "warn"))
}
```

```{r sponsor-queue-table}
if (!is.null(spon_queue) && nrow(spon_queue) > 0) {
  cols <- intersect(c("raw_sponsor", "proposed", "confidence", "n_trials", "review_reason"), names(spon_queue))
  spon_queue |>
    arrange(desc(.data[["n_trials"]])) |>
    slice_head(n = 15) |>
    select(all_of(cols)) |>
    mutate(across(any_of("confidence"), ~ sprintf("%.2f", suppressWarnings(as.numeric(.x))))) |>
    mutate(across(any_of("n_trials"), fmt_n)) |>
    nk(cap = "Highest-impact rows awaiting review — curate these first")
}
```

## Substance Names

Substance normalisation was rewritten as v2. The premise of the rewrite is that
substances, unlike sponsors, already have a canonical vocabulary: ChEMBL's
`pref_name` plus the EMA medicines report. That vocabulary is applied first and
deterministically, and a model is asked only about what it could not place.

| Pass | What it does | Model |
|---|---|---|
| `A_resolve` | exact match against ChEMBL + EPAR, placebo rule, junk filter | none |
| `B_assign` | pick-from-list over retrieved candidates | Sonnet 5 |
| `C_mint` | name what B could not place, cluster-at-a-time | Opus 5 / Sonnet 5 |
| `D_consolidate` | salt rollup (deterministic) + merge partition | Opus 5 |
| `E_emit` | per-trial labels, review queue, regression diff | none |

The canonical is the **INN base**: `Methotrexate` absorbs `methotrexate sodium`,
`Metoject` and `Methotrexat 10mg Tabletten`, with the salt and brand kept in
separate registry columns. Where ChEMBL states a USAN that differs from the INN,
the INN is used — this is an EU trial corpus.

```{r substance-audit-compute}
sub_has_v2 <- !is.null(sub_log) &&
  has_cols(sub_log, c("raw_substance", "substance_clean", "match_status", "n_trials"))

sub_raw_pairs  <- if (!is.null(sub_raw) && has_cols(sub_raw, c("_id", "raw_substance"))) nrow(sub_raw) else NA_integer_
sub_raw_unique <- if (!is.null(sub_raw) && "raw_substance" %in% names(sub_raw)) n_distinct(sub_raw$raw_substance, na.rm = TRUE) else NA_integer_
sub_label_trials <- if (!is.null(sub_labels) && "_id" %in% names(sub_labels)) n_distinct(sub_labels$`_id`) else NA_integer_
sub_queue_rows <- if (!is.null(sub_queue)) nrow(sub_queue) else NA_integer_

sub_pairs_accepted <- if (sub_has_v2) sum(sub_log$n_trials[sub_log$match_status == "accepted"], na.rm = TRUE) else NA_integer_
sub_pairs_rejected <- if (sub_has_v2) sum(sub_log$n_trials[sub_log$match_status == "rejected"], na.rm = TRUE) else NA_integer_
sub_pairs_unknown  <- if (sub_has_v2) sum(sub_log$n_trials[sub_log$match_status == "unknown"], na.rm = TRUE) else NA_integer_

sub_n_canon <- if (!is.null(sub_registry) && has_cols(sub_registry, c("canonical", "merged_into"))) {
  sum(is.na(sub_registry$merged_into) | sub_registry$merged_into == "")
} else if (sub_has_v2) n_distinct(sub_log$substance_clean[sub_log$match_status == "accepted"]) else NA_integer_

sub_n_displayed <- if (sub_has_v2) {
  n_distinct(sub_log$substance_clean[sub_log$match_status == "accepted" & !is.na(sub_log$substance_clean)])
} else NA_integer_

sub_wf <- tibble(
  step = c("Raw trial-substance pairs", "Not a substance (filtered or judged)",
           "Unresolved", "Accepted pairs", "Trials with a label"),
  n = c(sub_raw_pairs, sub_pairs_rejected, sub_pairs_unknown,
        sub_pairs_accepted, sub_label_trials),
  type = c("start", "step", "step", "step", "end")
) |>
  mutate(delta = lag(n) - n)

# Per-trial view. The v2 log is keyed on the raw STRING, not the trial, so the
# trial dimension comes back through the raw export.
active_by_trial <- if (sub_has_v2 && !is.null(sub_raw) && has_cols(sub_raw, c("_id", "raw_substance"))) {
  sub_raw |>
    inner_join(
      sub_log |>
        filter(match_status == "accepted", !is.na(substance_clean), substance_clean != "") |>
        select(raw_substance, substance_clean),
      by = "raw_substance"
    ) |>
    distinct(`_id`, substance_clean) |>
    left_join(df |> select(`_id`, register, year, age_group), by = "_id")
} else {
  tibble()
}
```

```{r substance-kpis, results='asis'}
if (sub_has_v2) {
  cat(kpi_row(
    kpi_box(fmt_n(sub_raw_pairs), "raw trial-substance pairs", BLUE),
    kpi_box(fmt_n(sub_raw_unique), "distinct raw strings", FROST),
    kpi_box(fmt_n(sub_n_displayed), "substances displayed", GREEN),
    kpi_box(fmt_n(sub_label_trials), "trials with a label", PURPLE),
    kpi_box(fmt_n(sub_queue_rows), "awaiting human review", ORANGE)
  ))
} else {
  cat(notice(
    "Substance v2 artefacts unavailable",
    paste("No <code>data/substance_normalisation_log_v2.csv</code> was found.",
          "Run <code>E_emit.R</code> in helper_scripts/substance_norm_pipeline_v2/."),
    "warn"))
}
```

```{r substance-waterfall, fig.height=4.3}
if (sub_has_v2 && all(!is.na(sub_wf$n))) {
  waterfall_plot(
    sub_wf,
    "Substance normalisation waterfall",
    "Pairs judged not-a-substance are removed deliberately, not lost."
  )
}
```

```{r substance-waterfall-table}
if (sub_has_v2) {
  sub_wf |>
    transmute(Step = step, Count = fmt_n(n),
              `Change from previous step` = ifelse(is.na(delta), "NA", fmt_n(delta))) |>
    nk(cap = "Substance normalisation stage counts")
}
```

```{r substance-status-chart, fig.height=3.5}
if (sub_has_v2) {
  sub_log |>
    count(match_status, wt = n_trials, name = "trial_weighted_n") |>
    mutate(match_status = factor(match_status, levels = c("accepted", "rejected", "unknown"))) |>
    ggplot(aes(match_status, trial_weighted_n, fill = match_status)) +
    geom_col(width = 0.62, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), vjust = -0.35, size = 3.1, color = DARK) +
    scale_fill_manual(values = c(accepted = GREEN, rejected = ORANGE, unknown = RED), na.value = GRAY) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    labs(title = "Substance outcomes weighted by trial count",
         subtitle = "'rejected' is a positive judgement that the string names no substance.",
         x = NULL, y = "Trial occurrences") +
    theme_pre()
}
```

```{r substance-decided-by-chart, fig.height=3.2}
# Who decided, and therefore how much of the corpus needed a model at all.
if (sub_has_v2 && "decided_by" %in% names(sub_log)) {
  sub_log |>
    filter(match_status == "accepted") |>
    mutate(decided_by = coalesce(as.character(decided_by), "not recorded")) |>
    count(decided_by, wt = n_trials, name = "trial_weighted_n", sort = TRUE) |>
    mutate(decided_by = reorder(decided_by, trial_weighted_n)) |>
    ggplot(aes(trial_weighted_n, decided_by)) +
    geom_col(fill = BLUE, width = 0.6) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), hjust = -0.1, size = 3.1, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    labs(title = "Who resolved each substance",
         subtitle = "registry = exact ChEMBL/EPAR match, no model involved.",
         x = "Trial occurrences", y = NULL) +
    theme_pre()
}
```

```{r substance-type-chart, fig.height=4}
if (sub_has_v2 && "substance_type" %in% names(sub_log)) {
  sub_log |>
    filter(match_status == "accepted", !is.na(substance_type), substance_type != "") |>
    count(substance_type, wt = n_trials, name = "trial_weighted_n", sort = TRUE) |>
    slice_head(n = 12) |>
    mutate(substance_type = reorder(substance_type, trial_weighted_n)) |>
    ggplot(aes(trial_weighted_n, substance_type)) +
    geom_col(fill = FROST, width = 0.62) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), hjust = -0.1, size = 3.1, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    labs(title = "Substance type", x = "Trial occurrences", y = NULL) +
    theme_pre()
}
```

```{r substance-top-chart, fig.height=5.5}
if (nrow(active_by_trial) > 0) {
  active_by_trial |>
    count(substance_clean, register, name = "n_trials") |>
    group_by(substance_clean) |>
    mutate(total = sum(n_trials)) |>
    ungroup() |>
    slice_max(total, n = 40, with_ties = FALSE) |>
    mutate(substance_clean = reorder(substance_clean, total)) |>
    ggplot(aes(n_trials, substance_clean, fill = register)) +
    geom_col() +
    scale_fill_manual(values = REG_COLS) +
    scale_x_continuous(labels = comma) +
    labs(title = "Top active substances by trial count",
         subtitle = "Canonical is the INN base; salts and brands are folded in.",
         x = "Trials", y = NULL) +
    theme_pre()
}
```

```{r substance-register-age, fig.height=4.8}
if (nrow(active_by_trial) > 0) {
  top_subs <- active_by_trial |>
    count(substance_clean, name = "total") |>
    slice_max(total, n = 12, with_ties = FALSE) |>
    pull(substance_clean)

  active_by_trial |>
    filter(substance_clean %in% top_subs) |>
    mutate(age_group = coalesce(age_group, "Unknown")) |>
    count(substance_clean, age_group, name = "n_trials") |>
    group_by(substance_clean) |>
    mutate(pct = n_trials / sum(n_trials)) |>
    ungroup() |>
    mutate(substance_clean = factor(substance_clean, levels = rev(top_subs))) |>
    ggplot(aes(pct, substance_clean, fill = age_group)) +
    geom_col() +
    scale_x_continuous(labels = percent) +
    scale_fill_manual(values = c(Paediatric = BLUE, Adult = ORANGE, `Paediatric & Adult` = GREEN, Unknown = GRAY), na.value = GRAY) +
    labs(title = "Age-group mix for top substances", x = "% of trials", y = NULL) +
    theme_pre()
}
```

```{r substance-rejected-table}
# The filter's output is auditable on purpose. If a real substance appears here,
# the judgement was wrong and nothing downstream would ever have asked about it.
if (!is.null(sub_rejected) && nrow(sub_rejected) > 0) {
  sub_rejected |>
    arrange(desc(n_trials)) |>
    select(any_of(c("raw_substance", "n_trials", "reason"))) |>
    dt_tbl(caption = "Strings rejected as not naming a substance", page = 15, filter = "top")
} else {
  cat("No rejected-string audit file available.")
}
```

```{r substance-queue-table}
if (!is.null(sub_queue) && nrow(sub_queue) > 0) {
  sub_queue |>
    arrange(desc(n_trials)) |>
    select(any_of(c("raw_substance", "proposed", "confidence", "n_trials",
                    "review_reason", "reason", "channel", "model_id"))) |>
    dt_tbl(caption = "Substance review queue — highest impact first", page = 25, filter = "top")
} else {
  cat("Substance review queue not available.")
}
```

```{r substance-log-table}
if (sub_has_v2) {
  sub_log |>
    arrange(match_status, desc(n_trials)) |>
    select(any_of(c("raw_substance", "substance_clean", "substance_salt",
                    "substance_type", "match_status", "confidence",
                    "decided_by", "n_trials"))) |>
    dt_tbl(caption = "Full substance normalisation log", page = 25, filter = "top")
} else {
  cat("Substance log not available.")
}
```

# 5. PIP, Results, and Orphan Flags

```{r boolean-flag-charts, fig.height=4}
flag_cols <- intersect(c("has_PIP", "has_results", "is_orphan"), names(df))
if (length(flag_cols) > 0) {
  flag_df <- df |>
    select(register, all_of(flag_cols)) |>
    pivot_longer(-register, names_to = "flag", values_to = "value", values_transform = list(value = as.character)) |>
    mutate(value = coalesce(as.character(value), "Unknown"))

  ggplot(flag_df |> count(flag, register, value, name = "n"), aes(register, n, fill = value)) +
    geom_col(width = 0.62) +
    facet_wrap(~ flag, scales = "free_y") +
    scale_y_continuous(labels = comma) +
    labs(title = "PIP, results, and orphan-derived flags", x = NULL, y = "Trials") +
    theme_pre()
}
```

```{r flag-tables}
if (length(flag_cols) > 0) {
  df |>
    select(register, all_of(flag_cols)) |>
    pivot_longer(-register, names_to = "Flag", values_to = "Value", values_transform = list(Value = as.character)) |>
    mutate(Value = coalesce(as.character(Value), "Unknown")) |>
    count(Flag, register, Value, name = "Trials") |>
    group_by(Flag, register) |>
    mutate(`% within register` = fmt_pct(Trials / sum(Trials) * 100), Trials = fmt_n(Trials)) |>
    ungroup() |>
    dt_tbl(caption = "Flag distributions by register", page = 30)
}
```

# 6. Completeness

```{r completeness-table}
comp_df <- tibble(
  Column = names(df),
  Type = vapply(df, function(x) class(x)[1], character(1)),
  `Non-missing` = vapply(df, function(x) sum(!(is.na(x) | (is.character(x) & x == ""))), integer(1)),
  `Complete (%)` = round(`Non-missing` / nrow(df) * 100, 1)
) |>
  arrange(`Complete (%)`)

dt_tbl(comp_df, caption = paste0("Completeness of all ", ncol(df), " cache columns"), page = 25, filter = "top")
```

```{r missingness-heatmap, fig.height=8, fig.width=10}
heat_cols <- intersect(c(
  "CT_number", "Full_title", "MEDDRA_term", "MEDDRA_organ_class", "Member_state",
  "sponsor_label", "sponsor_type", "phase", "status", "has_PIP", "has_results",
  "is_orphan", "submission_date_parsed", "decision_date", "start_date",
  "days_to_decision", "year", "DIMP_product_name", "substance_label"
), names(df))

if (length(heat_cols) > 0) {
  miss_df <- df |>
    select(register, all_of(heat_cols)) |>
    group_by(register) |>
    summarise(across(all_of(heat_cols), ~ mean(is.na(.x) | (is.character(.x) & .x == "")) * 100), .groups = "drop") |>
    pivot_longer(-register, names_to = "column", values_to = "pct_missing")

  ggplot(miss_df, aes(register, column, fill = pct_missing)) +
    geom_tile(color = "white", linewidth = 0.45) +
    geom_text(aes(label = sprintf("%.0f%%", pct_missing)), size = 2.8, color = DARK) +
    scale_fill_gradient2(low = GREEN, mid = YELLOW, high = RED, midpoint = 50, name = "% missing") +
    labs(title = "Missingness by register", x = NULL, y = NULL) +
    theme_pre() +
    theme(legend.position = "right", axis.text.y = element_text(size = 8.5))
}
```

# 7. Data Quality Issues

```{r issue-compute, include=FALSE}
count_blank <- function(x) sum(is.na(x) | (is.character(x) & x == ""), na.rm = TRUE)

unmapped_status_codes <- if (!is.null(log_status_c) && has_cols(log_status_c, c("register", "raw", "normalised", "n_trials"))) {
  log_status_c |>
    filter(register == "CTIS", !is.na(raw), str_detect(as.character(raw), "^[0-9]+$"), !(as.character(raw) %in% as.character(1:6))) |>
    arrange(desc(n_trials))
} else tibble()
n_ctis_unmapped_status <- if (nrow(unmapped_status_codes) > 0) sum(unmapped_status_codes$n_trials, na.rm = TRUE) else 0

country_fail_examples <- if (!is.null(log_country) && has_cols(log_country, c("raw", "normalised", "n_trials"))) {
  log_country |>
    filter(!is.na(raw), raw != "", is.na(normalised) | normalised == "") |>
    arrange(desc(n_trials)) |>
    slice_head(n = 30)
} else tibble()
n_country_fail <- if (nrow(country_fail_examples) > 0) sum(country_fail_examples$n_trials, na.rm = TRUE) else 0

organ_unresolved_examples <- if (!is.null(log_organ) && has_cols(log_organ, c("raw", "normalised", "n_trials"))) {
  log_organ |>
    filter(!is.na(normalised), str_detect(as.character(normalised), "^[0-9]+$")) |>
    arrange(desc(n_trials)) |>
    slice_head(n = 30)
} else tibble()
n_organ_unresolved <- if (nrow(organ_unresolved_examples) > 0) sum(organ_unresolved_examples$n_trials, na.rm = TRUE) else 0

ct_dupes_examples <- df |>
  filter(!is.na(canonical_trial_id), canonical_trial_id != "") |>
  add_count(canonical_trial_id) |>
  filter(n > 1) |>
  arrange(canonical_trial_id) |>
  select(any_of(c("_id", "register", "canonical_trial_id", "CT_number", "transition_eudract_number", "sponsor_label", "status")))
n_ct_dupes <- n_distinct(ct_dupes_examples$canonical_trial_id)

junk_pat <- regex("\\bGmbH\\b|\\bInc\\.?\\b|\\bLtd\\.?\\b|\\bS\\.A\\.\\b|\\bAG\\b|\\bB\\.V\\.\\b|\\bLLC\\b|\\bPLC\\b", ignore_case = FALSE)
sponsor_junk_ex <- df |>
  filter(!is.na(sponsor_label), str_detect(sponsor_label, junk_pat)) |>
  count(sponsor_label, sort = TRUE) |>
  slice_head(n = 30)
n_sponsor_junk <- sum(!is.na(df$sponsor_label) & str_detect(df$sponsor_label, junk_pat))

days_extreme_examples <- df |>
  filter(!is.na(days_to_decision), days_to_decision > 3650) |>
  arrange(desc(days_to_decision)) |>
  select(any_of(c("_id", "register", "sponsor_label", "submission_date_parsed", "decision_date", "days_to_decision"))) |>
  slice_head(n = 30)
n_days_extreme <- nrow(days_extreme_examples)

year_implaus_ex <- df |>
  filter(!is.na(year), year < 1990 | year > year(Sys.Date())) |>
  select(any_of(c("_id", "register", "submission_date", "submission_date_parsed", "year", "sponsor_label"))) |>
  slice_head(n = 30)
n_year_implaus <- nrow(year_implaus_ex)

issue_tbl <- tibble(
  Severity = c("High", "High", "High", "High", "High", "Medium", "Medium", "Medium", "Medium", "Low", "Low", "Low"),
  Category = c("Deduplication", "Deduplication", "Deduplication", "Status", "Dates", "Country", "Organ class", "Identifiers", "Sponsor", "Status", "Dates", "Dates"),
  Issue = c(
    "Residual cross-register duplicates by shared title key",
    "Transitioned EUCTR trials without a CTIS match in current fallbacks",
    "Duplicate canonical trial IDs in final cache",
    "CTIS numeric status codes outside the current mapping",
    "Completed trials with no decision_date",
    "Country strings normalised to missing",
    "CTIS organ-class codes still stored as raw numbers",
    "Canonical trial IDs appearing more than once",
    "Sponsor labels still contain legal-entity tokens",
    "Trials with status 'Other'",
    "days_to_decision greater than 10 years",
    "Submission year outside [1990, current year]"
  ),
  `N affected` = c(
    fmt_n(n_resid_dupes),
    fmt_n(n_unmatched_transitioned),
    fmt_n(n_ct_dupes),
    fmt_n(n_ctis_unmapped_status),
    fmt_n(sum(df$status == "Completed" & is.na(df$decision_date), na.rm = TRUE)),
    fmt_n(n_country_fail),
    fmt_n(n_organ_unresolved),
    fmt_n(n_ct_dupes),
    fmt_n(n_sponsor_junk),
    fmt_n(sum(df$status == "Other", na.rm = TRUE)),
    fmt_n(n_days_extreme),
    fmt_n(n_year_implaus)
  ),
  Suggested_action = c(
    "Review title-only matches and require corroborating identifiers before automated collapse.",
    "Refresh CTIS data and inspect missing transition EudraCT identifiers.",
    "Review canonical_trial_id derivation for transition aliases and latest-version preference.",
    "Extend CTIS status-code mapping if these codes are valid source values.",
    "Add or audit decision-date fallbacks for completed trials.",
    "Add unresolved country strings to the country lookup or preserve raw values for review.",
    "Extend the MedDRA SOC lookup for unresolved numeric CTIS codes.",
    "Review duplicate canonical identities and remove duplicate real-world trials.",
    "Add sponsor aliases for fallback labels that still include legal suffixes.",
    "Review raw status values and extend status category patterns where warranted.",
    "Treat implausible lag values as missing or audit source dates.",
    "Set implausible parsed years to missing after date parsing."
  )
)
```

## Summary

```{r issues-summary}
issue_tbl |>
  dt_tbl(caption = "Data-quality issues detected during preprocessing audit", page = 20, filter = "top") |>
  DT::formatStyle(
    "Severity",
    backgroundColor = DT::styleEqual(c("High", "Medium", "Low"), c("#FFEBEB", "#FFF3E0", "#FFFDE7"))
  )
```

## Detail Tables

```{r issue-detail-tabs, results='asis'}
cat("The tables below show examples for issues where example rows are available.")
```

```{r issue-residual-dupes}
if (nrow(cross_reg_tbl) > 0) {
  dt_tbl(cross_reg_tbl, caption = "Residual cross-register duplicate candidates", page = 25)
} else {
  cat("No residual cross-register duplicate candidates.")
}
```

```{r issue-unmapped-status}
if (nrow(unmapped_status_codes) > 0) {
  unmapped_status_codes |>
    select(Register = register, `Raw code` = raw, Normalised = normalised, `N trials` = n_trials) |>
    dt_tbl(caption = "Unmapped CTIS status codes", page = 20)
} else {
  cat("No unmapped CTIS status codes were detected.")
}
```

```{r issue-country-fail}
if (nrow(country_fail_examples) > 0) {
  country_fail_examples |>
    select(Register = register, `Raw country` = raw, `N trials` = n_trials) |>
    dt_tbl(caption = "Country strings normalised to missing", page = 20)
} else {
  cat("No country strings normalised to missing.")
}
```

```{r issue-organ-unresolved}
if (nrow(organ_unresolved_examples) > 0) {
  organ_unresolved_examples |>
    select(Register = register, Raw = raw, Normalised = normalised, `N trials` = n_trials) |>
    dt_tbl(caption = "Unresolved numeric organ-class values", page = 20)
} else {
  cat("No unresolved numeric organ-class values.")
}
```

```{r issue-canonical-dupes}
if (nrow(ct_dupes_examples) > 0) {
  dt_tbl(ct_dupes_examples, caption = "Duplicate canonical trial IDs", page = 25)
} else {
  cat("No duplicate canonical trial IDs.")
}
```

```{r issue-sponsor-junk}
if (nrow(sponsor_junk_ex) > 0) {
  sponsor_junk_ex |>
    rename(`Sponsor label` = sponsor_label, `N trials` = n) |>
    dt_tbl(caption = "Sponsor labels with legal suffix tokens", page = 20)
} else {
  cat("No sponsor labels with legal suffix tokens.")
}
```

```{r issue-days-extreme}
if (nrow(days_extreme_examples) > 0) {
  days_extreme_examples |>
    mutate(days_to_decision = fmt_n(days_to_decision)) |>
    dt_tbl(caption = "Extreme days_to_decision values", page = 20)
} else {
  cat("No days_to_decision values above 10 years.")
}
```

```{r issue-year-implaus}
if (nrow(year_implaus_ex) > 0) {
  dt_tbl(year_implaus_ex, caption = "Implausible submission years", page = 20)
} else {
  cat("No implausible submission years.")
}
```

---

*Generated by `rmarkdown/preprocessing.Rmd` at `r format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")`.*
