---
title: "Data Preprocessing Report"
subtitle: "EU Paediatric Clinical Trials — Pipeline Audit"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    self_contained: true
    fig_width: 9
    fig_height: 4.5
params:
  cache_path: "../trials_cache.rds"
  log_dir:    "../data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(stringr)
  library(lubridate)
  library(scales)
  library(knitr)
  library(kableExtra)
  library(readr)
  library(DT)
  library(plotly)
})

# ── Colour palette (Nord) ──────────────────────────────────────────────────────
BLUE   <- "#5E81AC"; FROST  <- "#81A1C1"; TEAL   <- "#88C0D0"
GREEN  <- "#A3BE8C"; YELLOW <- "#EBCB8B"; ORANGE <- "#D08770"
RED    <- "#BF616A"; PURPLE <- "#B48EAD"; GRAY   <- "#4C566A"
LIGHT  <- "#ECEFF4"; DARK   <- "#2E3440"
NORD8  <- c(BLUE, GREEN, YELLOW, RED, PURPLE, ORANGE, TEAL, FROST)

STATUS_COLS <- c(Ongoing = GREEN, Completed = BLUE, Withdrawn = ORANGE, Other = RED)
REG_COLS    <- c(EUCTR = BLUE, CTIS = TEAL)

# ── ggplot theme ───────────────────────────────────────────────────────────────
theme_pre <- function(base = 11) {
  theme_minimal(base_size = base, base_family = "sans") %+replace% theme(
    plot.title       = element_text(size = base + 1, face = "bold", color = DARK, hjust = 0),
    plot.subtitle    = element_text(size = base - 0.5, color = GRAY, hjust = 0),
    axis.title       = element_text(size = base - 0.5, color = GRAY),
    axis.text        = element_text(size = base - 1,   color = GRAY),
    axis.text.y      = element_text(margin = margin(r = 2)),
    panel.grid.major = element_line(color = LIGHT, linewidth = 0.45),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    legend.key.size  = unit(0.4, "cm"),
    legend.text      = element_text(size = base - 1.5, color = GRAY),
    plot.margin      = margin(6, 8, 6, 6),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.text       = element_text(size = base - 0.5, face = "bold", color = DARK)
  )
}

# ── Helpers ────────────────────────────────────────────────────────────────────
fmt_n   <- function(x) format(x, big.mark = ",")
fmt_pct <- function(x) sprintf("%.1f%%", x)

kpi_box <- function(val, label, colour)
  sprintf('<div style="background:#ECEFF4;border-left:4px solid %s;border-radius:4px;padding:10px 18px;min-width:140px;"><b>%s</b><br><small>%s</small></div>',
          colour, val, label)

nk <- function(x, cap = NULL, cn = colnames(x)) {
  kable(x, booktabs = TRUE, linesep = "", caption = cap, col.names = cn, escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, position = "left", font_size = 13)
}

clean_active_substance_label <- function(x) {
  x <- str_squish(as.character(x))
  x <- str_remove_all(x, regex("®|™", ignore_case = TRUE))
  x <- str_remove(x, regex("^placebo\\s+(matching|for|of|to)?\\s*", ignore_case = TRUE))
  x <- str_remove_all(x, regex("\\b\\d+[,.]?\\d*\\s*(mg|mcg|micrograms?|microgram|µg|ug|g|ml|l|%|units?|i\\.?u\\.?)\\b",
                               ignore_case = TRUE))
  x <- str_remove(x, regex("\\s+\\b[A-Z][A-Z0-9&.-]{2,}\\b$", ignore_case = FALSE))
  x <- str_remove(x, regex("\\s+\\b(accord|teva|viatris|aristo|normon|altan|efg|jenapharm|galen|sandoz|mylan|krka|zentiva|aurobindo|fresenius|hospira|stada|ratiopharm|cinfa|arrow|acis|pharma|allergan|allergan units?|aerosphere|kwikpen)\\b.*$",
                           ignore_case = TRUE))
  x <- str_remove_all(x, regex("\\b(etexilate|pentahydrate|hydrochloride|hydrobromide|sulfate|sulphate|phosphate|acetate|maleate|fumarate|tartrate|citrate|mesylate|tosylate|besylate|monohydrate|dihydrate|trihydrate|sesquihydrate|sodium|potassium|calcium|magnesium|zinc|aluminum|aluminium|monosodium|disodium)\\b",
                               ignore_case = TRUE))
  x <- str_remove(x, regex("[,;:-]?\\s*\\b(concentrate|solution|suspension|powder|film-coated|hard capsules?|capsules?|tablets?|injection|infusion|pre-filled syringes?|vials?|oral|intravenous|subcutaneous|injectable|inyectable|injektionslösung|injektionsloesung|iniettabile|solución|solucion|poudre|polvere|polvo|solvant|solvente|comprimidos|recubiertos|pel[ií]cula|jeringas?|precargadas|zachte|pressurised inhalation)\\b.*$",
                           ignore_case = TRUE))
  x <- str_remove(x, regex("\\s+\\d+[,.]?\\d*$", ignore_case = TRUE))
  x <- str_remove(x, regex("\\s+(and|\\+)$", ignore_case = TRUE))
  x <- str_remove_all(x, regex("\\b(of|w|w/v|v/v)\\b", ignore_case = TRUE))
  x <- str_remove_all(x, "[[:punct:]]+$")
  str_squish(str_to_sentence(str_to_lower(x)))
}

is_exploratory_substance <- function(x) {
  key <- str_to_lower(coalesce(as.character(x), ""))
  nchar(key) >= 3 &
    str_detect(key, "[[:alpha:]]") &
    !str_detect(key, regex("^\\d+[,.]?\\d*\\s*%?\\s*(w|w/v|v/v)?$|^dose$|^the placebo|\\bplacebo\\b|not yet established|not available|not applicable|^na$|^n/a$|^none$|^ml\\b|^mg\\b|micrograms?|concentrate|solution|solución|solucion|injectable|inyectable|injection|infusion|injektionslösung|iniettabile|comprimidos|recubiertos|pel[ií]cula|capsules?|tablets?|poudre|polvere|polvo|solvant|solvente|sodium chloride|nacl|magnesium stearate|kwikpen",
                           ignore_case = TRUE))
}

# ── Load data ──────────────────────────────────────────────────────────────────
cache_file <- params$cache_path
log_dir    <- params$log_dir

if (!file.exists(cache_file))
  stop("Cache not found at: ", cache_file, ". Run rebuild_cache.R first.")

t0 <- proc.time()
df <- readRDS(cache_file)
load_sec <- round((proc.time() - t0)["elapsed"], 1)
if (!"DIMP_inn_name" %in% names(df)) df$DIMP_inn_name <- NA_character_
if (!"transition_eudract_number" %in% names(df)) df$transition_eudract_number <- NA_character_
if (!"trial_identifiers" %in% names(df)) df$trial_identifiers <- df$CT_number
if (!"canonical_trial_id" %in% names(df)) {
  df$canonical_trial_id <- ifelse(
    df$register == "CTIS" & !is.na(df$transition_eudract_number),
    str_trim(str_split_fixed(df$transition_eudract_number, " / ", 2)[, 1]),
    ifelse(str_detect(df$`_id`, "^\\d{4}-\\d{6}-\\d{2}-[A-Z]"),
           str_replace(df$`_id`, "-[A-Z]{2,3}$", ""), df$`_id`)
  )
}
if (!"analysis_register" %in% names(df)) df$analysis_register <- df$register
if (!"sponsor_label" %in% names(df)) {
  slabels_path <- file.path(log_dir, "trial_sponsor_labels.csv")
  if (file.exists(slabels_path)) {
    sl <- readr::read_csv(slabels_path, show_col_types = FALSE,
                          col_types = readr::cols(`_id` = readr::col_character(),
                                                  sponsor_clean = readr::col_character()))
    df <- dplyr::left_join(df, dplyr::select(sl, `_id`, sponsor_clean), by = "_id")
    df$sponsor_label <- dplyr::coalesce(df$sponsor_clean, df$sponsor_name)
  } else {
    df$sponsor_label <- df$sponsor_name
  }
}

load_log <- function(name) {
  p <- file.path(log_dir, paste0(name, "_normalisation_log.csv"))
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}

log_country  <- load_log("country")
log_meddra   <- load_log("meddra_term")
log_organ    <- load_log("organ_class")
log_phase    <- load_log("phase")
log_status_c <- load_log("status_category")
log_status_d <- load_log("status_display")
```

---

# 1 · Dataset Overview {.tabset}

The cache was loaded in **`r load_sec` s** and contains **`r fmt_n(nrow(df))`** deduplicated trials across **`r ncol(df)`** columns.

```{r overview-cards}
reg_counts <- df %>% count(register)
n_tot   <- nrow(df)
n_euctr <- reg_counts$n[reg_counts$register == "EUCTR"]
n_ctis  <- reg_counts$n[reg_counts$register == "CTIS"]
```

<div style="display:flex;gap:24px;margin:12px 0 20px;">
<div style="background:#5E81AC;color:#fff;border-radius:8px;padding:16px 28px;min-width:160px;text-align:center;">
  <div style="font-size:2.2em;font-weight:700;">`r fmt_n(n_tot)`</div>
  <div style="font-size:0.95em;opacity:0.9;">Total trials</div>
</div>
<div style="background:#81A1C1;color:#fff;border-radius:8px;padding:16px 28px;min-width:160px;text-align:center;">
  <div style="font-size:2.2em;font-weight:700;">`r fmt_n(n_euctr)`</div>
  <div style="font-size:0.95em;opacity:0.9;">EUCTR</div>
</div>
<div style="background:#88C0D0;color:#fff;border-radius:8px;padding:16px 28px;min-width:160px;text-align:center;">
  <div style="font-size:2.2em;font-weight:700;">`r fmt_n(n_ctis)`</div>
  <div style="font-size:0.95em;opacity:0.9;">CTIS</div>
</div>
<div style="background:#A3BE8C;color:#fff;border-radius:8px;padding:16px 28px;min-width:160px;text-align:center;">
  <div style="font-size:2.2em;font-weight:700;">`r ncol(df)`</div>
  <div style="font-size:0.95em;opacity:0.9;">Columns</div>
</div>
</div>

## Register Distribution

```{r reg-bar, fig.height=3.5}
df %>%
  count(register) %>%
  ggplot(aes(register, n, fill = register)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = fmt_n(n)), vjust = -0.4, size = 3.5, color = DARK) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Trials by register", x = NULL, y = "Count") +
  theme_pre()
```

## Column Inventory

```{r col-inventory}
col_df <- tibble(
  Column  = names(df),
  Type    = sapply(df, function(x) class(x)[1]),
  NonMiss = sapply(df, function(x) sum(!is.na(x))),
  Pct     = round(NonMiss / nrow(df) * 100, 1)
) %>% arrange(Pct)

datatable(col_df,
  colnames  = c("Column", "Type", "Non-missing (n)", "Complete (%)"),
  rownames  = FALSE,
  filter    = "top",
  options   = list(pageLength = 20, scrollX = TRUE),
  caption   = "All columns in the final cache, sorted by completeness (ascending)"
)
```

---

# 2 · Age Group Coverage {.tabset}

`age_group` classifies each trial as **Paediatric**, **Adult**, **Paediatric & Adult**, or **Unknown**. EUCTR records are derived from the age eligibility flags (`f11_trial_has_subjects_under_18`, `f12_adults_1864_years`, `f13_elderly_65_years`). CTIS records are derived from `ageGroup`, where `"0-17 years"` marks paediatric eligibility and `"18-64 years"` / `"65+ years"` mark adult eligibility.

```{r age-group-stats}
age_levels <- c("Paediatric", "Adult", "Paediatric & Adult", "Unknown")

age_df <- df %>%
  mutate(
    age_group = replace_na(age_group, "Unknown"),
    age_group = factor(age_group, levels = age_levels)
  )

age_summary <- age_df %>%
  count(age_group, name = "Trials", .drop = FALSE) %>%
  mutate(
    `% of cache` = fmt_pct(Trials / sum(Trials) * 100),
    `Included by < 18 filter` = if_else(
      age_group %in% c("Paediatric", "Paediatric & Adult"),
      "Yes", "No"
    ),
    `Included by ≥ 18 filter` = if_else(
      age_group %in% c("Adult", "Paediatric & Adult"),
      "Yes", "No"
    )
  )

n_paediatric_filter <- age_df %>%
  filter(age_group %in% c("Paediatric", "Paediatric & Adult")) %>%
  nrow()
n_adult_filter <- age_df %>%
  filter(age_group %in% c("Adult", "Paediatric & Adult")) %>%
  nrow()
n_both <- age_df %>% filter(age_group == "Paediatric & Adult") %>% nrow()
```

<div style="display:flex;gap:16px;flex-wrap:wrap;margin:10px 0 16px;">
<div style="background:#ECEFF4;border-left:4px solid #5E81AC;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_paediatric_filter)`</b><br><small>shown by &lt; 18 years filter</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #A3BE8C;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_adult_filter)`</b><br><small>shown by ≥ 18 years filter</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #B48EAD;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_both)`</b><br><small>eligible for both filters</small>
</div>
</div>

## Classification Summary

```{r age-group-summary}
age_summary %>%
  mutate(Trials = fmt_n(Trials)) %>%
  nk(cap = "Adult / paediatric classification in the final cache")
```

## Register Split

```{r age-group-register, fig.height=3.8}
age_df %>%
  count(register, age_group, .drop = FALSE) %>%
  ggplot(aes(register, n, fill = age_group)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c(
    `Paediatric` = BLUE,
    `Adult` = GREEN,
    `Paediatric & Adult` = PURPLE,
    `Unknown` = GRAY
  ), drop = FALSE) +
  labs(title = "Age group classification by register",
       subtitle = "Paediatric & Adult trials are included in both sidebar age filters",
       x = NULL, y = "Proportion of trials") +
  theme_pre()
```

---

# 3 · Field Extraction & Column Mapping

`prepare_trial_data()` pulls separate field lists for EUCTR and CTIS from the SQLite database, then unites them into unified columns via `dplyr::unite()`.

```{r field-map}
field_map <- tribble(
  ~`Output column`,        ~`EUCTR source field`,                                             ~`CTIS source field`,
  "CT_number",             "a2_eudract_number",                                               "applicationInfo.ctNumber",
  "transition_eudract_number", NA_character_,                                                 "authorizedApplication.eudraCt.eudraCtCode",
  "Full_title",            "a3_full_title_of_the_trial",                                      "clinicalTrialIdentifiers.fullTitle",
  "DIMP_product_name",     "dimp.d31_product_name",                                           "products.productName",
  "DIMP_inn_name",         "dimp.d38_imp_identification_details.d38_inn__proposed_inn",       NA_character_,
  "MEDDRA_term",           "e12_meddra_classification.e12_term",                              "meddraConditionTerms.termName",
  "MEDDRA_organ_class",    "e12_meddra_classification.e12_system_organ_class",                "meddraConditionTerms.organClass",
  "Member_state",          "a1_member_state_concerned",                                       "memberStatesConcerned",
  "age_group",             "f11/f12/f13 age eligibility flags",                               "ageGroup",
  "PIP_status",            "a7_trial_is_part_of_a_paediatric_investigation_plan",             "scientificAdviceAndPip.paediatricInvestigationPlan",
  "trial_status_raw",      "x5_trial_status",                                                 "applicationInfo.trialStatus",
  "submission_date",       "x6_date_on_which_…_first_entered_in_the_eudract_database",        "applicationInfo.submissionDate",
  "decision_date",         "n_date_of_competent_authority_decision",                          "applicationInfo.decisionDate",
  "sponsor_name",          "b1_sponsor.b11_name_of_sponsor",                                  "sponsors.organisation.name",
  "sponsor_type",          "b1_sponsor.b31_and_b32_status_of_the_sponsor",                   "sponsors.commercial",
  "phase",                 "e71/e72/e73/e74_phase_i–iv (boolean flags)",                     "trialPhase (text)",
  "has_results",           "endPoints.endPoint.readyForValues (non-NA)",                      "resultsFirstReceived (TRUE)",
  "is_orphan",             "dimp.d25_…_orphan_drug_…_community (Yes/No text)",               "products.orphanDrugDesigNumber (EU/3/… code)",
  "start_date",            "n_date_of_competent_authority_decision / recruitmentStartDate",   "estimatedRecruitmentStartDate"
)

nk(field_map, cap = "Raw source fields unified into output columns")
```

**All `unite()` calls use `na.rm = TRUE` and `remove = TRUE`** — so the EUCTR value is kept when the CTIS field is NA, and vice versa.  List-typed columns are first collapsed by `deep_flatten_col()`, which recursively unlists nested JSON, removes NA/NULL/numeric-only tokens, and joins unique values with `" / "`.

---

# 4 · Deduplication Pipeline {.tabset}

The raw database may contain multiple rows per trial: EUCTR stores one row per **member state** (e.g., 2020-001234-56-BE, -DE, -FR), and CTIS stores multiple **amendment versions** (e.g., …-00, -01, -02). Cross-register overlaps exist when a trial appears in both registers (EudraCT trials that transitioned to CTIS). The pipeline resolves all of these in six sequential steps.

```{r dedup-db}
library(DBI)
library(RSQLite)

db_path <- "data/trials.sqlite"
has_db  <- file.exists(db_path)

is_euctr_id <- function(x) str_detect(x, "^\\d{4}-\\d{6}-\\d{2}-[A-Z]")

if (has_db) {
  con     <- dbConnect(RSQLite::SQLite(), db_path)
  all_ids <- dbGetQuery(con, 'SELECT "_id" FROM trials')[["_id"]]

  # EUCTR "transitioned" status — also fetch titles for later title-key matching
  trans_df <- dbGetQuery(con, '
    SELECT "_id",
      json_extract(json, "$.x5_trial_status")        AS trial_status,
      json_extract(json, "$.p_end_of_trial_status")  AS end_status,
      json_extract(json, "$.a3_full_title_of_the_trial") AS title
    FROM trials
    WHERE lower(json_extract(json, "$.x5_trial_status"))       LIKE "%transition%"
       OR lower(json_extract(json, "$.p_end_of_trial_status")) LIKE "%transition%"
  ')
  transitioned_ids <- trans_df[["_id"]]

  # EUCTR per-country raw IDs
  euctr_raw_ids <- all_ids[is_euctr_id(all_ids)]

  # CTIS raw IDs
  ctis_raw_ids <- all_ids[!is_euctr_id(all_ids)]

  dbDisconnect(con)
} else {
  all_ids          <- character(0)
  trans_df         <- data.frame(`_id`=character(), trial_status=character(),
                                  end_status=character(), title=character(),
                                  check.names=FALSE)
  transitioned_ids <- character(0)
  euctr_raw_ids    <- character(0)
  ctis_raw_ids     <- character(0)
}

# title_key helper (mirrors app.R logic)
make_title_key <- function(title) {
  tt <- tolower(as.character(title))
  tt <- str_replace_all(tt, "[^a-z0-9 ]", " ")
  tt <- str_squish(tt)
  substr(tt, 1, 80)
}

cache_ids <- df[["_id"]]
cache_euctr_base <- str_replace(cache_ids[is_euctr_id(cache_ids)], "-[A-Z]{2,3}$", "")

# ── Step counts ────────────────────────────────────────────────────────────────
euctr_all  <- all_ids[is_euctr_id(all_ids)]
ctis_all   <- all_ids[!is_euctr_id(all_ids)]

# EUCTR base IDs (strip 2–3-letter country code)
euctr_base_all   <- str_replace(euctr_all,  "-[A-Z]{2,3}$", "")
n_euctr_base     <- n_distinct(euctr_base_all)
avg_countries    <- round(length(euctr_all) / n_euctr_base, 1)

# CTIS base IDs (strip 2-digit version suffix)
ctis_base_all    <- str_replace(ctis_all,   "-\\d{2}$", "")
n_ctis_base      <- n_distinct(ctis_base_all)
n_ctis_amended   <- sum(table(ctis_base_all) > 1)   # bases with >1 version

# Waterfall steps computed from first principles so the chart always decreases.
# This is an audit summary, not a replay of every app-side keep/drop decision:
#   Step 1: unique EUCTR base trials + ALL CTIS records (country-variant collapse only)
#   Step 2: unique EUCTR base + unique CTIS base (CTIS amendment collapse)
#   Step 3+: final cache (cross-register dedup + any other drops)
# This avoids the DB-minus-cache diff approach which inflates n_drop_ctis_versions
# whenever the DB and cache are built at different times.
n_raw                  <- length(all_ids)
n_step1                <- n_euctr_base + length(ctis_all)   # after EUCTR country dedup
n_step2                <- n_euctr_base + n_ctis_base        # after CTIS amendment dedup
n_final                <- nrow(df)

n_drop_country_variant <- n_raw    - n_step1
n_drop_ctis_versions   <- n_step1  - n_step2
n_drop_cross_register  <- n_step2  - n_final    # cross-register + any other drops

# Pre-2023 CTIS labelled as EUCTR-era for analysis charts (source register unchanged)
n_pre2023 <- df %>%
  filter((register == "CTIS" & analysis_register == "EUCTR") |
         (register == "EUCTR" & str_detect(`_id`, "^\\d{4}-\\d{6}-\\d{2}-\\d"))) %>%
  mutate(yr = year(submission_date_parsed)) %>%
  filter(!is.na(yr) & yr < 2023) %>%
  nrow()
```

## Waterfall

```{r dedup-waterfall, fig.height=4.5}
wf <- tibble(
  step  = c(
    "Raw DB rows",
    "− EUCTR country variants",
    "− CTIS older amendments",
    "− Cross-register duplicates",
    "= Final cache"
  ),
  delta = c(
    n_raw,
    -n_drop_country_variant,
    -n_drop_ctis_versions,
    -n_drop_cross_register,
    NA
  ),
  cumul = c(
    n_raw,
    n_raw - n_drop_country_variant,
    n_raw - n_drop_country_variant - n_drop_ctis_versions,
    n_final,
    n_final
  ),
  type  = c("start", "drop", "drop", "drop", "end")
)

ggplot(wf, aes(x = forcats::fct_inorder(step), y = cumul, fill = type)) +
  geom_col(width = 0.55) +
  geom_text(aes(label = fmt_n(cumul)), vjust = -0.4, size = 3.5, color = DARK) +
  scale_fill_manual(values = c(start = BLUE, drop = RED, end = GREEN),
                    guide  = "none") +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Deduplication waterfall: rows removed at each step",
       subtitle = "Blue = raw start; red = rows removed; green = final",
       x = NULL, y = "Cumulative rows remaining") +
  theme_pre() +
  theme(axis.text.x = element_text(size = 9, angle = 15, hjust = 1))
```

```{r dedup-summary-table}
tribble(
  ~Step, ~Logic, ~`Rows removed`, ~`Rows remaining`,
  "Raw DB",                 "All records fetched from SQLite",
    "—", fmt_n(n_raw),
  "1. EUCTR per-country collapse", paste0("One row per member state → keep 1 per base ID; ", fmt_n(n_euctr_base), " unique EUCTR trials (avg ", avg_countries, " countries each)"),
    fmt_n(n_drop_country_variant), fmt_n(n_step1),
  "2. CTIS latest amendment",     paste0("Keep highest version suffix per base ID; ", fmt_n(n_ctis_amended), " base trials had ≥ 2 versions"),
    fmt_n(n_drop_ctis_versions), fmt_n(n_step2),
  "3–5. Cross-register / final dedup",    "EUCTR rows dropped when same trial found in CTIS (by CT#, title_key, or 'transitioned' status), plus any final cache exclusions",
    fmt_n(n_drop_cross_register), fmt_n(n_final),
  "6. Pre-2023 CTIS analysis label", "Source register stays CTIS; analysis_register changes to EUCTR for migrated EudraCT-era trials; row count unchanged",
    paste0("0 rows (", fmt_n(n_pre2023), " labelled)"), fmt_n(n_final)
) %>%
  nk(cap = "Deduplication steps with real row counts")
```

## EUCTR Country Variants {.tabset}

Each EUCTR trial protocol was submitted per member state, generating one row per country in the DB. The pipeline collapses these to a single row and aggregates country lists.

```{r dedup-euctr-before-after}
# Show top 12 trials with most country variants; compare raw DB vs cache
euctr_variant_tbl <- tibble(raw_id = euctr_raw_ids) %>%
  mutate(base = str_replace(raw_id, "-[A-Z]{2,3}$", ""),
         country = str_extract(raw_id, "[A-Z]{2,3}$")) %>%
  group_by(base) %>%
  summarise(
    n_countries   = n(),
    country_list  = paste(sort(country), collapse = ", "),
    kept_in_cache = base %in% cache_euctr_base,
    .groups = "drop"
  ) %>%
  filter(kept_in_cache) %>%
  arrange(desc(n_countries)) %>%
  slice_head(n = 15) %>%
  select(`Base trial ID` = base,
         `# countries in DB` = n_countries,
         `Country variants` = country_list)

datatable(euctr_variant_tbl,
  rownames = FALSE,
  options  = list(pageLength = 15, scrollX = TRUE),
  caption  = "Top 15 EUCTR trials by number of per-country rows in raw DB (collapsed to 1 row in cache)"
)
```

```{r dedup-euctr-dist, fig.height=3.5}
tibble(raw_id = euctr_raw_ids) %>%
  mutate(base = str_replace(raw_id, "-[A-Z]{2,3}$", "")) %>%
  count(base, name = "n_countries") %>%
  count(n_countries, name = "n_trials") %>%
  filter(n_countries <= 30) %>%
  ggplot(aes(n_countries, n_trials)) +
  geom_col(fill = BLUE, width = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "Distribution of per-country DB rows per EUCTR trial",
       subtitle = "Most trials are submitted in 1–5 member states",
       x = "Country variants (rows in raw DB)", y = "Number of trials") +
  theme_pre()
```

## CTIS Amendment Versions {.tabset}

CTIS assigns a version suffix (`-00`, `-01`, `-02`, …) to each amendment. Only the latest version is retained.

```{r dedup-ctis-versions}
ctis_version_tbl <- tibble(raw_id = ctis_all) %>%
  mutate(base    = str_replace(raw_id, "-\\d{2}$", ""),
         version = as.integer(str_extract(raw_id, "\\d{2}$"))) %>%
  group_by(base) %>%
  filter(n() > 1) %>%                           # only amended trials
	  summarise(
	    versions     = paste(sort(raw_id), collapse = " → "),
	    n_versions   = n(),
	    max_version  = max(version),
	    latest_id    = raw_id[which.max(version)],
	    kept_id      = raw_id[raw_id %in% cache_ids][1],
	    latest_kept  = latest_id %in% cache_ids,
	    .groups = "drop"
	  ) %>%
  arrange(desc(n_versions)) %>%
  slice_head(n = 20) %>%
  select(`Base CTIS ID` = base,
	         `# versions` = n_versions,
	         `Latest version kept` = max_version,
	         `Latest kept?` = latest_kept,
	         `Kept ID` = kept_id,
	         `All versions (old → new)` = versions)

datatable(ctis_version_tbl,
  rownames = FALSE,
  options  = list(pageLength = 20, scrollX = TRUE),
  caption  = paste0("CTIS trials with multiple amendment versions (", fmt_n(n_ctis_amended), " total). Only latest kept in cache.")
)
```

```{r dedup-ctis-version-dist, fig.height=3}
tibble(raw_id = ctis_all) %>%
  mutate(base = str_replace(raw_id, "-\\d{2}$", "")) %>%
  count(base, name = "n_versions") %>%
  count(n_versions, name = "n_trials") %>%
  ggplot(aes(n_versions, n_trials)) +
  geom_col(fill = TEAL, width = 0.7) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 1:10) +
  labs(title = "Distribution of amendment versions per CTIS trial",
       x = "Amendment versions in raw DB", y = "Number of base trials") +
  theme_pre()
```

## Cross-Register Duplicates {.tabset}

When the same trial appears in both EUCTR and CTIS, the EUCTR copy is dropped in favour of the CTIS record. This covers four overlap types: (a) CTIS transition applications that carry the old EudraCT number, (b) "Trial now transitioned" status in EUCTR, (c) same CT number base in both registers, (d) matching title_key.

```{r dedup-transitioned}
# Match transitioned EUCTR → CTIS via the same fallbacks as app.R:
# transition EudraCT number, CTIS base ID, then title_key.
n_transitioned <- n_distinct(transitioned_ids)

ctis_by_tk <- df %>%
  filter(register == "CTIS") %>%
  mutate(tk = make_title_key(Full_title)) %>%
  filter(!is.na(tk) & nchar(tk) >= 20) %>%
  select(ctis_id = `_id`, tk, ctis_title = Full_title,
         ctis_status = status, ctis_phase = phase)

ctis_by_transition <- df %>%
  filter(register == "CTIS", !is.na(transition_eudract_number)) %>%
  separate_rows(transition_eudract_number, sep = " / ") %>%
  mutate(transition_eudract_number = str_trim(transition_eudract_number)) %>%
  filter(transition_eudract_number != "") %>%
  distinct(transition_eudract_number, ctis_id_transition = `_id`,
           ctis_title_transition = Full_title, ctis_status_transition = status)

ctis_by_base <- df %>%
  filter(register == "CTIS") %>%
  mutate(ctis_base = str_replace(`_id`, "-\\d{2}$", "")) %>%
  distinct(ctis_base, ctis_id_base = `_id`,
           ctis_title_base = Full_title, ctis_status_base = status)

trans_base <- trans_df %>%
  as_tibble() %>%
  rename(euctr_id = `_id`) %>%
  mutate(
    euctr_base = str_replace(euctr_id, "-[A-Z]{2,3}$", ""),
    tk         = make_title_key(title)
  ) %>%
  distinct(euctr_base, tk, trial_status, end_status)

trans_matched <- trans_base %>%
  left_join(ctis_by_transition, by = c("euctr_base" = "transition_eudract_number")) %>%
  left_join(ctis_by_base, by = c("euctr_base" = "ctis_base")) %>%
  left_join(ctis_by_tk, by = "tk") %>%
  mutate(
    ctis_id = coalesce(ctis_id_transition, ctis_id_base, ctis_id),
    ctis_title = coalesce(ctis_title_transition, ctis_title_base, ctis_title),
    ctis_status = coalesce(ctis_status_transition, ctis_status_base, ctis_status),
    match_method = case_when(
      !is.na(ctis_id_transition) ~ "CTIS transition EudraCT number",
      !is.na(ctis_id_base)       ~ "CTIS base ID",
      !is.na(ctis_id)            ~ "title_key",
      TRUE                       ~ NA_character_
    )
  ) %>%
  distinct(euctr_base, .keep_all = TRUE)

n_matched   <- sum(!is.na(trans_matched$ctis_id))
n_unmatched <- sum( is.na(trans_matched$ctis_id))

trans_matched %>%
  arrange(!is.na(ctis_id), euctr_base) %>%
  #slice_head(n = 800) %>%
  mutate(
	    euctr_status_disp = coalesce(end_status, trial_status),
	    ctis_id_disp      = coalesce(ctis_id, "— no CTIS match found by current fallbacks"),
	    ctis_status_disp  = coalesce(ctis_status, "—")
	  ) %>%
	  select(
	    `EUCTR base ID`           = euctr_base,
	    `EUCTR status`            = euctr_status_disp,
	    `CTIS ID (kept)`          = ctis_id_disp,
	    `CTIS title`              = ctis_title,
	    `CTIS status`             = ctis_status_disp,
	    `Match method`            = match_method
	  ) %>%
  datatable(
    rownames = FALSE,
    options  = list(pageLength = 20, scrollX = TRUE),
    caption  = paste0(
	      fmt_n(n_transitioned), " unique EUCTR base trials flagged 'transitioned' in DB. ",
	      fmt_n(n_matched), " matched to a CTIS counterpart by transition EudraCT number, base ID, or title_key; ",
	      fmt_n(n_unmatched), " unmatched by current fallbacks."
	    )
	  )
```

```{r dedup-cross-reg-examples}
# Residual cross-register duplicates: same title_key in both EUCTR and CTIS in the cache.
# These are trials that slipped through all dedup steps and exist in both registers.
euctr_by_tk <- df %>%
  filter(register == "EUCTR") %>%
  mutate(tk = make_title_key(Full_title)) %>%
  filter(!is.na(tk) & nchar(tk) >= 20) %>%
  select(euctr_id = `_id`, tk, euctr_title = Full_title, euctr_status = status)

cross_reg_tbl <- ctis_by_tk %>%
  inner_join(euctr_by_tk, by = "tk") %>%
  filter(ctis_id != euctr_id) %>%
  select(
    `EUCTR ID (still in cache)` = euctr_id,
    `CTIS ID (still in cache)`  = ctis_id,
    `Shared title_key`          = tk,
    `EUCTR status`              = euctr_status,
    `CTIS status`               = ctis_status
  ) %>%
  slice_head(n = 50)

if (nrow(cross_reg_tbl) > 0) {
  datatable(cross_reg_tbl,
    rownames = FALSE,
    options  = list(pageLength = 20, scrollX = TRUE),
    caption  = paste0(
      fmt_n(nrow(cross_reg_tbl)), " trial(s) where the same title_key appears in both ",
      "EUCTR and CTIS in the final cache — potential residual duplicates not caught by dedup."
    )
  )
} else {
  cat("No residual cross-register title_key duplicates in final cache — deduplication is complete.")
}
```

## Pre-2023 CTIS Analysis Label {.tabset}

CTIS launched in January 2023. Records in CTIS with a submission date before 2023 are migrated EudraCT trials; their source `register` remains `"CTIS"`, while `analysis_register` is set to `"EUCTR"` for timeline charts (the row count does not change).

```{r dedup-pre2023}
pre2023_tbl <- df %>%
  filter((register == "CTIS" & analysis_register == "EUCTR") |
         (register == "EUCTR" & str_detect(`_id`, "^\\d{4}-\\d{6}-\\d{2}-\\d"))) %>%
  mutate(yr = year(submission_date_parsed)) %>%
  filter(!is.na(yr) & yr < 2023) %>%
  select(`CTIS ID` = `_id`,
         `Submission year` = yr,
         `Trial title` = Full_title,
         `Sponsor` = sponsor_label,
         `Status` = status) %>%
  arrange(`Submission year`) %>%
  slice_head(n = 30)

if (nrow(pre2023_tbl) > 0) {
  datatable(pre2023_tbl,
    rownames = FALSE,
    options  = list(pageLength = 20, scrollX = TRUE),
    caption  = paste0(fmt_n(n_pre2023), " CTIS trials with submission year < 2023, labelled analysis_register='EUCTR'")
  )
} else {
  # Fallback: show by year
  df %>%
    filter((register == "CTIS" & analysis_register == "EUCTR") |
           (register == "EUCTR" & str_detect(`_id`, "^\\d{4}-\\d{6}-\\d{2}-\\d"))) %>%
    mutate(yr = year(submission_date_parsed)) %>%
    filter(!is.na(yr) & yr < 2023) %>%
    count(yr) %>%
    nk(cap = paste0(fmt_n(n_pre2023), " pre-2023 CTIS records labelled EUCTR-era, by year"))
}
```

```{r dedup-pre2023-overlap}
# Check: do any CTIS records labelled EUCTR-era share a title_key with an EUCTR-original trial?
# That would mean the same real-world trial appears twice in the cache.
euctr_orig_tk <- df %>%
  filter(register == "EUCTR",
         str_detect(`_id`, "^\\d{4}-\\d{6}-\\d{2}-[A-Z]")) %>%
  mutate(tk = make_title_key(Full_title)) %>%
  filter(!is.na(tk) & nchar(tk) >= 20) %>%
  select(euctr_orig_id = `_id`, tk, euctr_title = Full_title, euctr_status = status)

euctr_relab_tk <- df %>%
  filter((register == "CTIS" & analysis_register == "EUCTR") |
         (register == "EUCTR" & str_detect(`_id`, "^\\d{4}-\\d{6}-\\d{2}-\\d"))) %>%
  mutate(tk = make_title_key(Full_title)) %>%
  filter(!is.na(tk) & nchar(tk) >= 20) %>%
  select(relab_id = `_id`, tk, relab_title = Full_title, relab_yr = year)

overlap_tbl <- euctr_orig_tk %>%
  inner_join(euctr_relab_tk, by = "tk") %>%
  select(`EUCTR original ID`         = euctr_orig_id,
         `CTIS labelled EUCTR-era ID` = relab_id,
         `Shared title_key`          = tk,
         `Submission year (labelled)` = relab_yr,
         `EUCTR status`              = euctr_status)

n_overlap <- nrow(overlap_tbl)

if (n_overlap > 0) {
  overlap_tbl %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(
        fmt_n(n_overlap), " trial(s) appear as BOTH 'EUCTR (original)' AND ",
        "'CTIS labelled EUCTR-era' in the cache — same real-world trial counted ",
        "TWICE in the bar chart below. These are dedup failures."
      ))
} else {
  cat("No overlap between EUCTR-original and CTIS records labelled EUCTR-era — the bar chart below is not double-counting.")
}
```

```{r dedup-pre2023-bar, fig.height=3.5}
df %>%
  mutate(yr = year(submission_date_parsed)) %>%
  filter(!is.na(yr), yr >= 2000, yr <= 2024) %>%
  mutate(source = case_when(
    (register == "CTIS" & analysis_register == "EUCTR") |
      (register == "EUCTR" & str_detect(`_id`, "^\\d{4}-\\d{6}-\\d{2}-\\d")) ~
      "CTIS labelled EUCTR-era",
    register == "EUCTR" ~ "EUCTR (original)",
    register == "CTIS" ~ "CTIS (native)",
    TRUE ~ "Other"
  )) %>%
  count(yr, source) %>%
  ggplot(aes(yr, n, fill = source)) +
  geom_col() +
  scale_fill_manual(values = c(
    `EUCTR (original)`             = BLUE,
	    `CTIS labelled EUCTR-era`      = FROST,
    `CTIS (native)`                = TEAL
  )) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2025, 2)) +
  labs(title = "Trials by authorization year and data origin",
	       subtitle = "Frost = CTIS source records labelled EUCTR-era for analysis (pre-2023 migrated records)",
       x = NULL, y = "Trials") +
  theme_pre()
```

---

# 5 · Country Normalisation {.tabset}

`clean_member_state()` splits multi-country strings on `" / "` or `", "`, strips junk tokens (numerics, dates, UUIDs, single characters), resolves aliases via `COUNTRY_LOOKUP`, and rejoins sorted unique values.

```{r country-stats}
lc <- log_country %>% filter(!is.na(raw) & raw != "")
n_raw    <- nrow(lc)
n_uniq_r <- n_distinct(lc$raw)
n_uniq_n <- n_distinct(lc$normalised, na.rm = TRUE)
n_chg    <- sum(lc$changed, na.rm = TRUE)
```

<div style="display:flex;gap:16px;flex-wrap:wrap;margin:10px 0 16px;">
<div style="background:#ECEFF4;border-left:4px solid #5E81AC;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_raw)`</b><br><small>raw log entries</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #88C0D0;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_uniq_r)`</b><br><small>unique raw values</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #A3BE8C;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_uniq_n)`</b><br><small>unique normalised values</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #D08770;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_chg)`</b> (`r fmt_pct(n_chg/n_raw*100)`)<br><small>entries changed</small>
</div>
</div>

## Top Countries

```{r country-bar, fig.height=5}
df %>%
  filter(!is.na(Member_state)) %>%
  separate_rows(Member_state, sep = " / ") %>%
  filter(str_trim(Member_state) != "") %>%
  count(Member_state, sort = TRUE) %>%
  slice_head(n = 25) %>%
  mutate(Member_state = forcats::fct_reorder(Member_state, n)) %>%
  ggplot(aes(n, Member_state, fill = n)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = FROST, high = BLUE) +
  scale_x_continuous(labels = comma) +
  labs(title = "Top 25 countries after normalisation",
       subtitle = "Counts are trial-country combinations (one trial can appear in multiple countries)",
       x = "Trial–country combinations", y = NULL) +
  theme_pre() +
  theme(axis.text.y = element_text(size = 9))
```

## Changed Entries

```{r country-changed}
log_country %>%
  filter(changed == TRUE, !is.na(raw)) %>%
  arrange(desc(n_trials)) %>%
  select(register, raw, normalised, n_trials) %>%
  datatable(
    colnames = c("Register", "Raw value", "Normalised", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "Country entries that were changed during normalisation"
  )
```

## Alias Examples

Key alias resolutions hard-coded in `COUNTRY_LOOKUP`:

```{r country-aliases}
tribble(
  ~`Raw alias`,          ~`Canonical name`,
  "uk, u.k., great britain", "United Kingdom",
  "us, usa, u.s.a., united states of america", "United States",
  "czechia",              "Czech Republic",
  "republic of korea, korea", "South Korea",
  "russian federation",   "Russia",
  "türkiye",              "Turkey"
) %>% nk(cap = "Alias mappings in COUNTRY_LOOKUP")
```

---

# 6 · MedDRA Term Normalisation {.tabset}

`clean_meddra_term()` fixes American → British spelling variants and converts Roman numeral type suffixes to Arabic numerals, then deduplicates within multi-term strings.

```{r meddra-stats}
lm <- log_meddra %>% filter(!is.na(raw))
n_raw_m  <- nrow(lm)
n_uniq_rm <- n_distinct(lm$raw)
n_uniq_nm <- n_distinct(lm$normalised, na.rm = TRUE)
n_chg_m  <- sum(lm$changed, na.rm = TRUE)
```

<div style="display:flex;gap:16px;flex-wrap:wrap;margin:10px 0 16px;">
<div style="background:#ECEFF4;border-left:4px solid #5E81AC;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_raw_m)`</b><br><small>raw log entries</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #A3BE8C;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_uniq_rm)`</b> → <b>`r fmt_n(n_uniq_nm)`</b><br><small>unique raw → normalised</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #D08770;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_chg_m)`</b> (`r fmt_pct(n_chg_m/n_raw_m*100)`)<br><small>entries changed</small>
</div>
</div>

## Top MedDRA Terms

```{r meddra-bar, fig.height=5.5}
df %>%
  filter(!is.na(MEDDRA_term)) %>%
  separate_rows(MEDDRA_term, sep = " / ") %>%
  mutate(MEDDRA_term = str_trim(MEDDRA_term)) %>%
  filter(MEDDRA_term != "") %>%
  count(MEDDRA_term, sort = TRUE) %>%
  slice_head(n = 25) %>%
  mutate(MEDDRA_term = forcats::fct_reorder(MEDDRA_term, n)) %>%
  ggplot(aes(n, MEDDRA_term, fill = n)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = TEAL, high = BLUE) +
  scale_x_continuous(labels = comma) +
  labs(title = "Top 25 MedDRA condition terms after normalisation",
       x = "Trial count", y = NULL) +
  theme_pre() +
  theme(axis.text.y = element_text(size = 8.5))
```

## Spelling Fixes

```{r meddra-spelling}
log_meddra %>%
  filter(changed == TRUE, !is.na(raw)) %>%
  filter(!str_detect(normalised, "\\d")) %>%  # not Roman→Arabic
  arrange(desc(n_trials)) %>%
  select(register, raw, normalised, n_trials) %>%
  datatable(
    colnames = c("Register", "Raw term", "Normalised", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "MedDRA terms changed by spelling normalisation"
  )
```

## Roman → Arabic Numerals

```{r meddra-roman}
log_meddra %>%
  filter(changed == TRUE, !is.na(raw)) %>%
  filter(str_detect(raw, "Type (I|II|III|IV)|type (i|ii|iii|iv)")) %>%
  arrange(desc(n_trials)) %>%
  select(register, raw, normalised, n_trials) %>%
  datatable(
    colnames = c("Register", "Raw (Roman)", "Normalised (Arabic)", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "MedDRA terms where Roman numeral type suffixes were converted to Arabic"
  )
```

---

# 7 · Organ Class Normalisation {.tabset}

`clean_organ_class()` handles two distinct raw formats:

- **EUCTR**: `"10021881 - Infections and infestations"` → strip numeric prefix
- **CTIS**: `"10021881"` (code only) → lookup in `ctis_soc_lookup` table

```{r organ-stats}
lo <- log_organ %>% filter(!is.na(raw))
n_raw_o  <- nrow(lo)
n_uniq_ro <- n_distinct(lo$raw)
n_uniq_no <- n_distinct(lo$normalised, na.rm = TRUE)
n_chg_o  <- sum(lo$changed, na.rm = TRUE)
```

<div style="display:flex;gap:16px;flex-wrap:wrap;margin:10px 0 16px;">
<div style="background:#ECEFF4;border-left:4px solid #5E81AC;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_raw_o)`</b><br><small>raw log entries</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #A3BE8C;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_uniq_ro)`</b> → <b>`r fmt_n(n_uniq_no)`</b><br><small>unique raw → normalised</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #D08770;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_chg_o)`</b> (`r fmt_pct(n_chg_o/n_raw_o*100)`)<br><small>entries changed</small>
</div>
</div>

## Distribution by Organ Class

```{r organ-bar, fig.height=5}
df %>%
  filter(!is.na(MEDDRA_organ_class)) %>%
  separate_rows(MEDDRA_organ_class, sep = " / ") %>%
  mutate(MEDDRA_organ_class = str_trim(MEDDRA_organ_class)) %>%
  filter(MEDDRA_organ_class != "") %>%
  count(MEDDRA_organ_class, sort = TRUE) %>%
  slice_head(n = 20) %>%
  mutate(MEDDRA_organ_class = forcats::fct_reorder(MEDDRA_organ_class, n)) %>%
  ggplot(aes(n, MEDDRA_organ_class, fill = n)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = FROST, high = PURPLE) +
  scale_x_continuous(labels = comma) +
  labs(title = "Top 20 MedDRA System Organ Classes after normalisation",
       x = "Trial count", y = NULL) +
  theme_pre() +
  theme(axis.text.y = element_text(size = 9))
```

## Changed Entries

```{r organ-changed}
log_organ %>%
  filter(changed == TRUE, !is.na(raw)) %>%
  mutate(type = if_else(str_detect(raw, "^[0-9]+ - "), "EUCTR prefix strip",
                        if_else(str_detect(raw, "^[0-9]+$"), "CTIS code lookup", "Other"))) %>%
  arrange(type, desc(n_trials)) %>%
  select(register, type, raw, normalised, n_trials) %>%
  datatable(
    colnames = c("Register", "Change type", "Raw", "Normalised", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "Organ class entries changed during normalisation, by change type"
  )
```

---

# 8 · Sponsor Name Normalisation {.tabset}

The sponsor normalisation pipeline runs at **cache-build time** (not at session start).
It maps raw sponsor strings from EUCTR and CTIS to clean canonical labels via:
manual overrides → negative-alias reject → exact alias lookup → fuzzy match (JW ≥ 0.92) → abstain.

Output is stored in `data/trial_sponsor_labels.csv` and joined to `rv$data` by `_id` at
startup as `sponsor_label`. The tabs below audit every step.

```{r spon-data-load}
spon_log <- {
  p <- file.path(log_dir, "sponsor_normalisation_log.csv")
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}

spon_labels <- {
  p <- file.path(log_dir, "trial_sponsor_labels.csv")
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}

spon_queue <- {
  p <- file.path(dirname(log_dir), "config", "sponsor_norm_pipeline",
                 "sponsor_review_queue.csv")
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}
```

### Match status

```{r spon-status-kpis, results='asis', eval=!is.null(spon_log)}
n_spon_in  <- n_distinct(spon_log$raw_sponsor)
n_spon_acc <- n_distinct(spon_log$raw_sponsor[spon_log$match_status == "accepted"])
n_spon_rev <- n_distinct(spon_log$raw_sponsor[spon_log$match_status == "review"])
n_spon_rej <- n_distinct(spon_log$raw_sponsor[spon_log$match_status == "rejected"])
n_spon_unk <- n_distinct(spon_log$raw_sponsor[spon_log$match_status == "unknown"])
n_spon_lab <- if (!is.null(spon_labels)) n_distinct(spon_labels$`_id`) else NA_integer_

cat(paste0(
  '<div style="display:flex;gap:14px;flex-wrap:wrap;margin:10px 0 16px;">',
  kpi_box(fmt_n(n_spon_in),                          "unique raw sponsor strings",        BLUE),
  kpi_box(fmt_pct(n_spon_acc / n_spon_in * 100),     "accepted rate",                     GREEN),
  kpi_box(fmt_pct(n_spon_rev / n_spon_in * 100),     "review rate",                       YELLOW),
  kpi_box(fmt_pct(n_spon_rej / n_spon_in * 100),     "rejected rate",                     ORANGE),
  kpi_box(fmt_pct(n_spon_unk / n_spon_in * 100),     "unknown rate",                      RED),
  kpi_box(fmt_n(n_spon_lab),                         "trials with a sponsor label",       TEAL),
  '</div>'
))
```

```{r spon-status-chart, fig.height=3.2, eval=!is.null(spon_log)}
spon_log %>%
  distinct(raw_sponsor, match_status) %>%
  count(match_status) %>%
  mutate(
    match_status = factor(match_status,
                          levels = c("accepted", "review", "rejected", "unknown")),
    fill_col = c(accepted = GREEN, review = YELLOW,
                 rejected = ORANGE, unknown = RED)[as.character(match_status)]
  ) %>%
  ggplot(aes(n, match_status, fill = fill_col)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3.2, colour = GRAY) +
  scale_fill_identity() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Unique raw sponsor strings by normalisation outcome",
       x = "Unique raw sponsor strings", y = NULL) +
  theme_pre()
```

```{r spon-source-chart, fig.height=3.5, eval=!is.null(spon_log)}
spon_log %>%
  filter(match_status == "accepted") %>%
  distinct(raw_sponsor, match_source) %>%
  mutate(
    source_group = case_when(
      grepl("manual_override|override", match_source) ~ "Manual override",
      grepl("epar|ema",                 match_source, ignore.case = TRUE) ~ "EMA EPAR",
      grepl("ror",                      match_source, ignore.case = TRUE) ~ "ROR academic",
      grepl("manual|alias",             match_source, ignore.case = TRUE) ~ "Manual alias",
      grepl("fuzzy",                    match_source, ignore.case = TRUE) ~ "Fuzzy match",
      TRUE ~ coalesce(match_source, "Other")
    )
  ) %>%
  count(source_group, sort = TRUE) %>%
  mutate(source_group = factor(source_group, levels = rev(source_group))) %>%
  ggplot(aes(n, source_group)) +
  geom_col(fill = BLUE, width = 0.6) +
  geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3.2, colour = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Accepted strings by match source",
       x = "Unique raw sponsor strings", y = NULL) +
  theme_pre()
```

### Top sponsors

```{r spon-top-prep, eval=!is.null(spon_labels)}
cache_meta_spon <- tryCatch(
  readRDS(params$cache_path) %>%
    select(`_id`, register, year) %>%
    mutate(year = suppressWarnings(as.integer(year))),
  error = function(e) NULL
)

spon_by_trial <- if (!is.null(cache_meta_spon)) {
  spon_labels %>%
    filter(match_status == "accepted", !is.na(sponsor_clean), nchar(sponsor_clean) > 0) %>%
    distinct(`_id`, sponsor_clean) %>%
    left_join(cache_meta_spon, by = "_id")
} else NULL
```

```{r spon-top-chart, fig.height=5.5, eval=!is.null(spon_by_trial)}
top_spon <- spon_by_trial %>%
  count(sponsor_clean, register, name = "n_trials")

top20s <- top_spon %>%
  group_by(sponsor_clean) %>%
  summarise(total = sum(n_trials), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 20) %>%
  pull(sponsor_clean)

top_spon %>%
  filter(sponsor_clean %in% top20s) %>%
  mutate(sponsor_clean = factor(sponsor_clean, levels = rev(top20s))) %>%
  ggplot(aes(n_trials, sponsor_clean, fill = register)) +
  geom_col() +
  scale_fill_manual(values = REG_COLS) +
  labs(title = "Top 20 sponsors by trial count (accepted matches only)",
       x = "Trials", y = NULL) +
  theme_pre()
```

```{r spon-top-unknowns, fig.height=4.5, eval=!is.null(spon_log)}
spon_log %>%
  filter(match_status == "unknown") %>%
  group_by(raw_sponsor) %>%
  summarise(n_trials = sum(n_trials, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(n_trials)) %>%
  slice_head(n = 25) %>%
  mutate(raw_sponsor = forcats::fct_reorder(raw_sponsor, n_trials)) %>%
  ggplot(aes(n_trials, raw_sponsor)) +
  geom_col(fill = RED, width = 0.65) +
  geom_text(aes(label = n_trials), hjust = -0.1, size = 3, colour = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Top 25 unmatched sponsors by trial count",
       subtitle = "These drive the 'unknown' rate and are the highest-value curation targets",
       x = "Trials", y = NULL) +
  theme_pre()
```

```{r spon-trial-dist, fig.height=3.5, eval=!is.null(spon_log)}
spon_dist <- spon_log %>%
  group_by(raw_sponsor) %>%
  summarise(
    n_trials     = sum(n_trials, na.rm = TRUE),
    match_status = first(match_status),
    .groups = "drop"
  ) %>%
  mutate(
    bin = case_when(
      n_trials == 1  ~ "1",
      n_trials == 2  ~ "2",
      n_trials == 3  ~ "3",
      n_trials <= 5  ~ "4–5",
      n_trials <= 10 ~ "6–10",
      n_trials <= 20 ~ "11–20",
      TRUE           ~ "21+"
    ),
    bin = factor(bin, levels = c("1", "2", "3", "4–5", "6–10", "11–20", "21+"))
  )

pct_spon_one <- round(100 * mean(spon_dist$n_trials == 1))
n_spon_one   <- sum(spon_dist$n_trials == 1)

spon_dist %>%
  count(bin, match_status) %>%
  mutate(match_status = factor(match_status,
    levels = c("accepted", "review", "rejected", "unknown"))) %>%
  ggplot(aes(bin, n, fill = match_status)) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(accepted = GREEN, review = YELLOW, rejected = ORANGE, unknown = RED),
    name   = NULL
  ) +
  annotate("text", x = 1, y = Inf,
    label  = sprintf("%d%% of raw strings\nappear in only 1 trial", pct_spon_one),
    hjust  = 0.5, vjust = 1.3, size = 3.2, colour = GRAY) +
  labs(
    title    = "Sponsor string frequency distribution",
    subtitle = sprintf(
      "%s unique strings appear in exactly 1 trial — the main curation opportunity",
      fmt_n(n_spon_one)
    ),
    x = "Trials per raw sponsor string", y = "Unique raw strings"
  ) +
  theme_pre()
```

### Review queue

```{r spon-queue-kpis, results='asis', eval=!is.null(spon_queue)}
n_sq       <- nrow(spon_queue)
n_sq_dec   <- sum(!is.na(spon_queue$decision) & nzchar(spon_queue$decision), na.rm = TRUE)
n_sq_pend  <- n_sq - n_sq_dec
n_sq_rev   <- sum(spon_queue$match_status == "review", na.rm = TRUE)
n_sq_unk   <- sum(spon_queue$match_status == "unknown", na.rm = TRUE)

cat(paste0(
  '<div style="display:flex;gap:14px;flex-wrap:wrap;margin:10px 0 16px;">',
  kpi_box(fmt_n(n_sq),      "rows in review queue",   BLUE),
  kpi_box(fmt_n(n_sq_rev),  "review status",          YELLOW),
  kpi_box(fmt_n(n_sq_unk),  "unknown status",         RED),
  kpi_box(fmt_n(n_sq_dec),  "decisions recorded",     GREEN),
  kpi_box(fmt_n(n_sq_pend), "pending review",         ORANGE),
  '</div>'
))
```

```{r spon-queue-table, eval=!is.null(spon_queue)}
spon_queue %>%
  arrange(desc(n_trials)) %>%
  mutate(
    match_score = round(as.numeric(match_score), 1),
    decision    = coalesce(decision, "")
  ) %>%
  select(
    `Raw sponsor`       = raw_sponsor,
    `Candidate`         = candidate_sponsor,
    Status              = match_status,
    Score               = match_score,
    Source              = match_source,
    Reason              = match_reason,
    `N trials`          = n_trials,
    Decision            = decision,
    `Canonical`         = canonical_sponsor
  ) %>%
  datatable(
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 25, scrollX = TRUE),
    caption  = "Sponsor review queue — rows needing manual curation"
  )
```

```{r spon-queue-status-chart, fig.height=3, eval=!is.null(spon_queue)}
spon_queue %>%
  count(match_status, sort = TRUE) %>%
  mutate(
    match_status = factor(match_status, levels = c("review", "unknown")),
    fill_col = c(review = YELLOW, unknown = RED)[as.character(match_status)]
  ) %>%
  ggplot(aes(n, match_status, fill = fill_col)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3.2, colour = GRAY) +
  scale_fill_identity() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Queue rows by status", x = "Rows", y = NULL) +
  theme_pre()
```

### Full normalisation log

```{r spon-log-table, eval=!is.null(spon_log)}
spon_log %>%
  arrange(match_status, desc(match_score)) %>%
  mutate(match_score = round(as.numeric(match_score), 1)) %>%
  select(
    `Raw sponsor`    = raw_sponsor,
    `Clean label`    = sponsor_clean,
    Status           = match_status,
    Score            = match_score,
    Source           = match_source,
    Reason           = match_reason,
    `N trials`       = n_trials
  ) %>%
  datatable(
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 25, scrollX = TRUE),
    caption  = "Full sponsor normalisation log — one row per unique raw sponsor string"
  )
```


# 9 · Sponsor Type Derivation

`sponsor_type` is pattern-matched from free-text fields:

- **EUCTR**: `b1_sponsor.b31_and_b32_status_of_the_sponsor` — patterns `"non.commercial|non commercial|academic"` → Academic; `"commercial"` → Industry
- **CTIS**: `sponsors.commercial` — same patterns

```{r sponsor-type-bar, fig.height=3.5}
df %>%
  mutate(sponsor_type = replace_na(sponsor_type, "Unknown / NA")) %>%
  count(register, sponsor_type) %>%
  ggplot(aes(register, n, fill = sponsor_type)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c(Academic = BLUE, Industry = GREEN,
                               `Unknown / NA` = GRAY)) +
  labs(title = "Sponsor type by register",
       x = NULL, y = "Proportion of trials") +
  theme_pre()
```

```{r sponsor-type-coverage}
df %>%
  group_by(register) %>%
  summarise(
    Total    = n(),
    Academic = sum(sponsor_type == "Academic", na.rm = TRUE),
    Industry = sum(sponsor_type == "Industry", na.rm = TRUE),
    Unknown  = sum(is.na(sponsor_type)),
    `% classified` = fmt_pct((Academic + Industry) / Total * 100)
  ) %>%
  nk(cap = "Sponsor type coverage by register")
```

---

# 10 · Phase Derivation {.tabset}

Phase is derived differently for each register:

- **EUCTR**: four boolean flag columns (`e71_…_phase_i` through `e74_…_phase_iv`) — `TRUE/YES/1` → label; multiple flags possible → joined with `" / "`
- **CTIS**: free-text `trialPhase` field parsed by keyword matching (e.g., "Therapeutic confirmatory (Phase III)" → "Phase III")

```{r phase-stats}
lp <- log_phase %>% filter(!is.na(raw))
n_raw_p  <- nrow(lp)
n_chg_p  <- sum(lp$changed, na.rm = TRUE)
```

<div style="display:flex;gap:16px;flex-wrap:wrap;margin:10px 0 16px;">
<div style="background:#ECEFF4;border-left:4px solid #5E81AC;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_raw_p)`</b><br><small>raw log entries</small>
</div>
<div style="background:#ECEFF4;border-left:4px solid #D08770;border-radius:4px;padding:10px 18px;">
  <b>`r fmt_n(n_chg_p)`</b> (`r fmt_pct(n_chg_p/n_raw_p*100)`)<br><small>entries changed (CTIS text → label)</small>
</div>
</div>

## Phase Distribution

```{r phase-bar, fig.height=4}
df %>%
  filter(!is.na(phase)) %>%
  separate_rows(phase, sep = " / ") %>%
  mutate(phase = str_trim(phase)) %>%
  filter(str_detect(phase, "^Phase")) %>%
  count(register, phase) %>%
  ggplot(aes(phase, n, fill = register)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma) +
  labs(title = "Trial phase distribution by register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r phase-multiphase}
n_multi <- df %>% filter(str_detect(phase, " / ")) %>% nrow()
```

**Multi-phase EUCTR trials** (multiple boolean flags set simultaneously): **`r fmt_n(n_multi)`** trials (`r fmt_pct(n_multi/nrow(df)*100)`).

## CTIS Text → Phase Mapping

```{r phase-ctis-mapping}
log_phase %>%
  filter(register == "CTIS", !is.na(raw)) %>%
  arrange(desc(n_trials)) %>%
  select(raw, normalised, n_trials) %>%
  datatable(
    colnames = c("CTIS raw text", "Mapped phase", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 20, scrollX = TRUE),
    caption  = "CTIS trialPhase text values and their parsed phase labels"
  )
```

---

# 11 · Status Derivation {.tabset}

Status is derived in three layers:

1. **`status_raw_orig`** — coalesce of EUCTR end-of-trial status, CTIS status code, and EUCTR trial status; CTIS numeric codes (1–6) mapped to labels
2. **`status_raw`** — standardised display string: splits multi-country aggregations, removes numeric tokens, normalises casing
3. **`status`** — four-category classification: `Ongoing`, `Completed`, `Withdrawn`, `Other`

```{r status-dist, fig.height=3.5}
df %>%
  mutate(status = replace_na(status, "Other")) %>%
  count(register, status) %>%
  ggplot(aes(register, n, fill = status)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = STATUS_COLS) +
  labs(title = "Trial status distribution by register", x = NULL, y = "Proportion") +
  theme_pre()
```

## CTIS Numeric Code Mapping

```{r status-ctis-codes}
tribble(
  ~Code, ~Label,
  "1", "Authorised",
  "2", "In Progress → Ongoing",
  "3", "Completed",
  "4", "Terminated",
  "5", "Temporarily halted",
  "6", "Withdrawn",
  "8", "Other (unmapped code → 'Other' category)"
) %>% nk(cap = "CTIS numeric status codes and their label mappings")
```

## Status Category Log

```{r status-cat-log}
log_status_c %>%
  filter(changed == TRUE, !is.na(raw)) %>%
  arrange(desc(n_trials)) %>%
  select(register, raw, normalised, n_trials) %>%
  datatable(
    colnames = c("Register", "Raw (status_raw_orig)", "Category (status)", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "Status entries changed when mapping to Ongoing / Completed / Other"
  )
```

## Status Display Log

```{r status-disp-log}
log_status_d %>%
  filter(changed == TRUE, !is.na(raw)) %>%
  arrange(desc(n_trials)) %>%
  select(register, raw, normalised, n_trials) %>%
  datatable(
    colnames = c("Register", "Raw (status_raw_orig)", "Display (status_raw)", "N trials"),
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 15, scrollX = TRUE),
    caption  = "Status entries changed when standardising the display string"
  )
```

---

# 12 · Date Derivation & Days-to-Decision {.tabset}

Three date columns are derived:

| Column | Source | Notes |
|---|---|---|
| `submission_date_parsed` | `submission_date` (united) | Parsed via `lubridate::parse_date_time()` with `orders = c("ymd","ym","y","ymd HMS")` |
| `decision_date` | Register-specific field | EUCTR: `n_date_of_competent_authority_decision`; CTIS: `applicationInfo.decisionDate` |
| `start_date` | Multiple fields | Max of available dates across sources |
| `days_to_decision` | `decision_date − submission_date_parsed` | Set to `NA` if negative (data entry errors) |

## Trials by Year

```{r year-bar, fig.height=4}
df %>%
  filter(!is.na(year), year >= 2000, year <= year(Sys.Date())) %>%
  count(year, register) %>%
  ggplot(aes(year, n, fill = register)) +
  geom_col() +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2025, 2)) +
  labs(title = "Trials authorized per year by register",
       subtitle = "Based on submission_date_parsed year",
       x = NULL, y = "Trials") +
  theme_pre()
```

## Date Parse Coverage

```{r date-coverage}
df %>%
  group_by(register) %>%
  summarise(
    Total               = n(),
    `submission parsed` = fmt_pct(sum(!is.na(submission_date_parsed)) / Total * 100),
    `decision date`     = fmt_pct(sum(!is.na(decision_date)) / Total * 100),
    `start date`        = fmt_pct(sum(!is.na(start_date)) / Total * 100),
    `days_to_decision`  = fmt_pct(sum(!is.na(days_to_decision)) / Total * 100)
  ) %>%
  nk(cap = "Date parse success rate by register (% non-NA)")
```

## Days-to-Decision Distribution

```{r days-hist, fig.height=4}
d2d <- df %>%
  filter(!is.na(days_to_decision), days_to_decision > 0) %>%
  mutate(log_days = log10(days_to_decision))

tick_vals <- log10(c(1, 10, 30, 100, 365, 1000, 3650))
tick_lbls <- c("1", "10", "30", "100", "365", "1000", "3650")

d2d %>%
  ggplot(aes(log_days, fill = register)) +
  geom_histogram(bins = 50, alpha = 0.75, position = "identity") +
  scale_x_continuous(breaks = tick_vals, labels = tick_lbls) +
  scale_fill_manual(values = REG_COLS) +
  labs(title = "Days from submission to decision (log₁₀ scale)",
       subtitle = "Negative values set to NA; zero-day records excluded from this view",
       x = "Days", y = "Trials") +
  theme_pre()
```

```{r days-summary}
df %>%
  filter(!is.na(days_to_decision)) %>%
  group_by(register) %>%
  summarise(
    N        = fmt_n(n()),
    Median   = round(median(days_to_decision), 0),
    Q1       = round(quantile(days_to_decision, 0.25), 0),
    Q3       = round(quantile(days_to_decision, 0.75), 0),
    Min      = min(days_to_decision),
    Max      = max(days_to_decision),
    `% NA`   = fmt_pct(sum(is.na(df$days_to_decision[df$register == register[1]])) /
                         nrow(df[df$register == register[1],]) * 100)
  ) %>%
  nk(cap = "Days-to-decision summary statistics by register")
```

---

# 13 · Substance Name Normalisation {.tabset}

The substance normalisation pipeline runs at **cache-build time** (not at session start).
It maps raw `DIMP_inn_name` / `DIMP_product_name` strings to clean canonical INN labels
via a multi-step entity-resolution pipeline: placebo rule → manual overrides →
negative-alias reject → pattern reject → canonical lookup → brand/alias lookup →
conservative fuzzy match → abstain.

Output is stored in `data/trial_substance_labels.csv` and joined to `rv$data` by `_id`
at startup. The tabs below audit every step.

```{r sub-data-load}
sub_log <- {
  p <- file.path(log_dir, "substance_normalisation_log.csv")
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}

sub_labels <- {
  p <- file.path(log_dir, "trial_substance_labels.csv")
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}

sub_queue <- {
  p <- file.path(dirname(log_dir), "config", "substance_norm_pipeline",
                 "substance_review_queue.csv")
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else NULL
}

# Trial metadata for joins (register, year, age_group)
cache_meta <- tryCatch(
  readRDS(params$cache_path) %>%
    select(`_id`, register, year, age_group) %>%
    mutate(year = suppressWarnings(as.integer(year))),
  error = function(e) NULL
)
```

### Match status

```{r sub-status-kpis, results='asis', eval=!is.null(sub_log)}
n_unique_in  <- n_distinct(sub_log$raw_substance)
n_accepted   <- n_distinct(sub_log$raw_substance[sub_log$match_status == "accepted"])
n_review     <- n_distinct(sub_log$raw_substance[sub_log$match_status == "review"])
n_rejected   <- n_distinct(sub_log$raw_substance[sub_log$match_status == "rejected"])
n_unknown    <- n_distinct(sub_log$raw_substance[sub_log$match_status == "unknown"])
n_trials_lab <- if (!is.null(sub_labels)) n_distinct(sub_labels$`_id`) else NA_integer_

cat(paste0(
  '<div style="display:flex;gap:14px;flex-wrap:wrap;margin:10px 0 16px;">',
  kpi_box(fmt_n(n_unique_in),                      "unique raw strings normalised",    BLUE),
  kpi_box(fmt_pct(n_accepted / n_unique_in * 100), "accepted rate",                   GREEN),
  kpi_box(fmt_pct(n_review   / n_unique_in * 100), "review rate",                     YELLOW),
  kpi_box(fmt_pct(n_rejected / n_unique_in * 100), "rejected rate",                   ORANGE),
  kpi_box(fmt_pct(n_unknown  / n_unique_in * 100), "unknown rate",                    RED),
  kpi_box(fmt_n(n_trials_lab),                     "trials with a substance label",   TEAL),
  '</div>'
))
```

```{r sub-status-chart, fig.height=3.2, eval=!is.null(sub_log)}
sub_log %>%
  distinct(raw_substance, match_status) %>%
  count(match_status) %>%
  mutate(
    match_status = factor(match_status,
                          levels = c("accepted", "review", "rejected", "unknown")),
    fill_col = c(accepted = GREEN, review = YELLOW,
                 rejected = ORANGE, unknown = RED)[as.character(match_status)]
  ) %>%
  ggplot(aes(n, match_status, fill = fill_col)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3.2, colour = GRAY) +
  scale_fill_identity() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Unique raw strings by normalisation outcome",
       x = "Unique raw substance strings", y = NULL) +
  theme_pre()
```

```{r sub-source-chart, fig.height=3.5, eval=!is.null(sub_log)}
sub_log %>%
  filter(match_status == "accepted") %>%
  distinct(raw_substance, match_source) %>%
  mutate(
    source_group = case_when(
      grepl("canonical",    match_source) ~ "Canonical INN list",
      grepl("manual",       match_source) ~ "Manual brand table",
      grepl("epar|ema",     match_source, ignore.case = TRUE) ~ "EMA EPAR",
      grepl("chembl",       match_source, ignore.case = TRUE) ~ "ChEMBL",
      grepl("fuzzy",        match_source) ~ "Fuzzy match",
      grepl("placebo",      match_source) ~ "Placebo rule",
      grepl("override",     match_source) ~ "Manual override",
      TRUE ~ match_source
    )
  ) %>%
  count(source_group, sort = TRUE) %>%
  mutate(source_group = factor(source_group, levels = rev(source_group))) %>%
  ggplot(aes(n, source_group)) +
  geom_col(fill = BLUE, width = 0.6) +
  geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3.2, colour = GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Accepted strings by match source",
       x = "Unique raw substance strings", y = NULL) +
  theme_pre()
```

### Top substances

```{r sub-top-prep, eval=!is.null(sub_log) && !is.null(cache_meta)}
# Join log → cache for trial-level counts by register
active_by_trial <- sub_log %>%
  filter(match_status == "accepted",
         !is.na(active_substance_clean), nchar(active_substance_clean) > 0) %>%
  distinct(`_id`, active_substance_clean) %>%
  left_join(cache_meta, by = "_id")
```

```{r sub-top-chart, fig.height=5.5, eval=!is.null(sub_log) && !is.null(cache_meta)}
top_sub <- active_by_trial %>%
  count(active_substance_clean, register, name = "n_trials")

top20 <- top_sub %>%
  group_by(active_substance_clean) %>%
  summarise(total = sum(n_trials), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 20) %>%
  pull(active_substance_clean)

top_sub %>%
  filter(active_substance_clean %in% top20) %>%
  mutate(active_substance_clean = factor(active_substance_clean, levels = rev(top20))) %>%
  ggplot(aes(n_trials, active_substance_clean, fill = register)) +
  geom_col() +
  scale_fill_manual(values = REG_COLS) +
  labs(title = "Top 20 active substances by trial count",
       subtitle = "Accepted matches only; one trial counted once per substance",
       x = "Trials", y = NULL) +
  theme_pre()
```

```{r sub-trial-dist, fig.height=3.5, eval=!is.null(sub_log)}
sub_dist <- sub_log %>%
  group_by(raw_substance) %>%
  summarise(
    n_trials     = n_distinct(`_id`),
    match_status = first(match_status),
    .groups = "drop"
  ) %>%
  mutate(
    bin = case_when(
      n_trials == 1  ~ "1",
      n_trials == 2  ~ "2",
      n_trials == 3  ~ "3",
      n_trials <= 5  ~ "4–5",
      n_trials <= 10 ~ "6–10",
      n_trials <= 20 ~ "11–20",
      TRUE           ~ "21+"
    ),
    bin = factor(bin, levels = c("1", "2", "3", "4–5", "6–10", "11–20", "21+"))
  )

pct_sub_one <- round(100 * mean(sub_dist$n_trials == 1))
n_sub_one   <- sum(sub_dist$n_trials == 1)

sub_dist %>%
  count(bin, match_status) %>%
  mutate(match_status = factor(match_status,
    levels = c("accepted", "review", "rejected", "unknown"))) %>%
  ggplot(aes(bin, n, fill = match_status)) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(accepted = GREEN, review = YELLOW, rejected = ORANGE, unknown = RED),
    name   = NULL
  ) +
  annotate("text", x = 1, y = Inf,
    label  = sprintf("%d%% of raw strings\nappear in only 1 trial", pct_sub_one),
    hjust  = 0.5, vjust = 1.3, size = 3.2, colour = GRAY) +
  labs(
    title    = "Substance string frequency distribution",
    subtitle = sprintf(
      "%s unique strings appear in exactly 1 trial — the main curation opportunity",
      fmt_n(n_sub_one)
    ),
    x = "Trials per raw substance string", y = "Unique raw strings"
  ) +
  theme_pre()
```

### Timeline

```{r sub-timeline, fig.height=4.8, eval=!is.null(sub_log) && !is.null(cache_meta)}
top8 <- active_by_trial %>%
  count(active_substance_clean, name = "total") %>%
  arrange(desc(total)) %>%
  slice_head(n = 8) %>%
  pull(active_substance_clean)

active_by_trial %>%
  filter(active_substance_clean %in% top8, !is.na(year)) %>%
  count(year, active_substance_clean, name = "n_trials") %>%
  ggplot(aes(year, n_trials, colour = active_substance_clean)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  scale_colour_manual(
    values = setNames(rep(NORD8, length.out = length(top8)), top8)) +
  scale_x_continuous(breaks = pretty_breaks()) +
  labs(title = "Trial volume for top 8 substances over time",
       x = "Submission year", y = "Trials") +
  theme_pre()
```

### Register & age mix

```{r sub-reg-age, fig.height=4.8, eval=!is.null(sub_log) && !is.null(cache_meta)}
top10 <- active_by_trial %>%
  count(active_substance_clean, name = "total") %>%
  arrange(desc(total)) %>%
  slice_head(n = 10) %>%
  pull(active_substance_clean)

reg_mix <- active_by_trial %>%
  filter(active_substance_clean %in% top10) %>%
  count(active_substance_clean, register, name = "n_trials") %>%
  group_by(active_substance_clean) %>%
  mutate(pct = n_trials / sum(n_trials)) %>%
  ungroup() %>%
  mutate(active_substance_clean = factor(active_substance_clean, levels = rev(top10)))

age_mix <- active_by_trial %>%
  filter(active_substance_clean %in% top10) %>%
  mutate(age_group = coalesce(age_group, "Unknown")) %>%
  count(active_substance_clean, age_group, name = "n_trials") %>%
  group_by(active_substance_clean) %>%
  mutate(pct = n_trials / sum(n_trials)) %>%
  ungroup() %>%
  mutate(active_substance_clean = factor(active_substance_clean, levels = rev(top10)))

p1 <- ggplot(reg_mix, aes(pct, active_substance_clean, fill = register)) +
  geom_col() +
  scale_x_continuous(labels = percent) +
  scale_fill_manual(values = REG_COLS) +
  labs(title = "Register mix", x = "% of trials", y = NULL) +
  theme_pre()

p2 <- ggplot(age_mix, aes(pct, active_substance_clean, fill = age_group)) +
  geom_col() +
  scale_x_continuous(labels = percent) +
  scale_fill_manual(values = c(Paediatric = BLUE, Adult = ORANGE,
                               `Paediatric & Adult` = GREEN, Unknown = GRAY)) +
  labs(title = "Age-group profile", x = "% of trials", y = NULL) +
  theme_pre()

subplot(ggplotly(p1), ggplotly(p2), nrows = 1, margin = 0.05, shareY = TRUE) %>%
  layout(showlegend = TRUE)
```

### Review queue

```{r sub-queue-kpis, results='asis', eval=!is.null(sub_queue)}
n_q       <- nrow(sub_queue)
n_decided <- sum(!is.na(sub_queue$decision) & nzchar(sub_queue$decision), na.rm = TRUE)
n_pending <- n_q - n_decided

cat(paste0(
  '<div style="display:flex;gap:14px;flex-wrap:wrap;margin:10px 0 16px;">',
  kpi_box(fmt_n(n_q),       "rows in review queue (n_occurrences ≥ 2)", BLUE),
  kpi_box(fmt_n(n_decided), "decisions recorded",                             GREEN),
  kpi_box(fmt_n(n_pending), "pending review",                                 ORANGE),
  '</div>'
))
```

```{r sub-queue-table, eval=!is.null(sub_queue)}
sub_queue %>%
  arrange(desc(n_occurrences)) %>%
  mutate(
    match_score = round(match_score, 1),
    decision    = coalesce(decision, "")
  ) %>%
  select(
    `Raw substance`    = raw_substance,
    `Suggested clean`  = active_substance_clean,
    Status             = match_status,
    Score              = match_score,
    Source             = match_source,
    Reason             = match_reason,
    `N trials`         = n_occurrences,
    Decision           = decision
  ) %>%
  datatable(
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 25, scrollX = TRUE),
    caption  = "Substance review queue — rows needing manual curation (singletons excluded)"
  )
```

```{r sub-queue-status-chart, fig.height=3, eval=!is.null(sub_queue)}
sub_queue %>%
  count(match_status, sort = TRUE) %>%
  mutate(
    match_status = factor(match_status, levels = c("review", "unknown")),
    fill_col = c(review = YELLOW, unknown = RED)[as.character(match_status)]
  ) %>%
  ggplot(aes(n, match_status, fill = fill_col)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3.2, colour = GRAY) +
  scale_fill_identity() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Queue rows by status", x = "Rows", y = NULL) +
  theme_pre()
```

### Full normalisation log

```{r sub-log-table, eval=!is.null(sub_log)}
sub_log %>%
  arrange(match_status, desc(match_score)) %>%
  mutate(match_score = round(match_score, 1)) %>%
  select(
    `Raw substance`   = raw_substance,
    `Clean label`     = active_substance_clean,
    Status            = match_status,
    Score             = match_score,
    Source            = match_source,
    Reason            = match_reason
  ) %>%
  datatable(
    rownames = FALSE, filter = "top",
    options  = list(pageLength = 25, scrollX = TRUE),
    caption  = "Full substance normalisation log — one row per accepted trial-substance pair"
  )
```

---

# 14 · PIP Status, Results, and Orphan Status {.tabset}

| Column | EUCTR source | CTIS source | Categories |
|---|---|---|---|
| `has_PIP` | `a7_…_paediatric_investigation_plan` text | PIP record presence | Yes / No / Unknown |
| `DIMP_inn_name` | `dimp.d38_…_d38_inn__proposed_inn`; fallback `dimp.d391_inn_generic_name` | NA | INN / generic name |
| `has_results` | `endPoints.endPoint.readyForValues` (non-NA) | `resultsFirstReceived` (= "true") | TRUE / FALSE |
| `is_orphan` | `dimp.d25_…_orphan_drug_…` ("yes"/"no") | `orphanDrugDesigNumber` (non-empty) | Yes / No / Unknown |

## PIP Status

```{r pip-bar, fig.height=3.5}
df %>%
  count(register, has_PIP) %>%
  ggplot(aes(register, n, fill = has_PIP)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c(Yes = GREEN, No = RED, Unknown = YELLOW)) +
  labs(title = "PIP (Paediatric Investigation Plan) status by register",
       x = NULL, y = "Proportion") +
  theme_pre()
```

## Results Posted

```{r results-bar, fig.height=3.5}
df %>%
  mutate(has_results = if_else(has_results, "Results posted", "No results")) %>%
  count(register, has_results) %>%
  ggplot(aes(register, n, fill = has_results)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c(`Results posted` = GREEN, `No results` = RED)) +
  labs(title = "Results posted by register", x = NULL, y = "Proportion") +
  theme_pre()
```

```{r results-by-status}
df %>%
  filter(!is.na(status)) %>%
  group_by(status) %>%
  summarise(
    Trials         = fmt_n(n()),
    `With results` = fmt_n(sum(has_results, na.rm = TRUE)),
    `% with results` = fmt_pct(mean(has_results, na.rm = TRUE) * 100)
  ) %>%
  nk(cap = "Results posted rate by trial status")
```

## Orphan Status

```{r orphan-bar, fig.height=3.5}
df %>%
  count(register, is_orphan) %>%
  ggplot(aes(register, n, fill = is_orphan)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c(Yes = PURPLE, No = BLUE, Unknown = GRAY)) +
  labs(title = "Orphan drug designation by register", x = NULL, y = "Proportion") +
  theme_pre()
```

```{r orphan-x-sponsor}
df %>%
  filter(!is.na(sponsor_type), !is.na(is_orphan)) %>%
  count(sponsor_type, is_orphan) %>%
  group_by(sponsor_type) %>%
  mutate(pct = n / sum(n)) %>%
  select(-n) %>%
  pivot_wider(names_from = is_orphan, values_from = pct, values_fill = 0) %>%
  mutate(across(where(is.numeric), fmt_pct)) %>%
  nk(cap = "Orphan designation distribution by sponsor type (% within sponsor type)")
```

---

# 14 · Final Dataset Completeness

```{r completeness-table}
comp_df <- tibble(
  Column   = names(df),
  Type     = sapply(df, function(x) class(x)[1]),
  NonMiss  = sapply(df, function(x) sum(!is.na(x))),
  Pct      = round(NonMiss / nrow(df) * 100, 1)
) %>% arrange(Pct)

datatable(comp_df,
  colnames  = c("Column", "Type", "Non-missing (n)", "Complete (%)"),
  rownames  = FALSE,
  filter    = "top",
  options   = list(pageLength = 25, scrollX = TRUE),
  caption   = paste0("Completeness of all ", ncol(df), " columns in the final cache (",
                     fmt_n(nrow(df)), " trials total)")
)
```

## Missingness Heatmap by Register

```{r missingness-heatmap, fig.height=8, fig.width=10}
key_cols <- c("CT_number", "Full_title", "MEDDRA_term", "MEDDRA_organ_class",
              "Member_state", "sponsor_label", "sponsor_type", "phase", "status",
              "status_raw", "has_PIP", "has_results", "is_orphan",
              "submission_date_parsed", "decision_date", "start_date",
              "days_to_decision", "year", "DIMP_product_name")

key_cols <- key_cols[key_cols %in% names(df)]

miss_df <- df %>%
  select(register, all_of(key_cols)) %>%
  group_by(register) %>%
  summarise(across(all_of(key_cols), ~ mean(is.na(.x)) * 100), .groups = "drop") %>%
  pivot_longer(-register, names_to = "column", values_to = "pct_missing")

ggplot(miss_df, aes(register, column, fill = pct_missing)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.0f%%", pct_missing)), size = 3, color = DARK) +
  scale_fill_gradient2(low = GREEN, mid = YELLOW, high = RED,
                       midpoint = 50, name = "% missing") +
  labs(title = "Missingness (%) by column and register",
       subtitle = "Green = complete, red = fully missing",
       x = NULL, y = NULL) +
  theme_pre() +
  theme(axis.text.y = element_text(size = 9),
        legend.position = "right")
```

---

# 15 · Data Quality Issues {.tabset}

All anomalies detected during the pipeline audit, ranked by severity. Each subsection gives detail and a concrete suggested fix for `app.R`.

```{r issues-compute, include=FALSE}
pct_na   <- function(x) mean(is.na(x) | (is.character(x) & x == "")) * 100
count_na <- function(x) sum(is.na(x)  | (is.character(x) & x == ""))

# 1. Residual cross-register duplicates (computed in dedup-cross-reg-examples)
n_resid_dupes <- nrow(cross_reg_tbl)

# 1b. Within-analysis duplicates: EUCTR-original + CTIS labelled EUCTR-era sharing title_key
#     (these cause double-counting in the bar chart)
n_within_euctr_overlap <- nrow(overlap_tbl)

# 2. n_unmatched already set in dedup-transitioned chunk

# 3. CTIS status codes outside the 1-6 mapping
unmapped_status_codes <- log_status_c %>%
  filter(register == "CTIS", !is.na(raw)) %>%
  filter(str_detect(raw, "^\\d+$") & !(raw %in% c("1","2","3","4","5","6"))) %>%
  arrange(desc(n_trials))
n_ctis_unmapped_status <- unmapped_status_codes %>% summarise(n = sum(n_trials)) %>% pull(n)

# 4. Completed trials with no decision_date
n_completed_no_date <- df %>% filter(status == "Completed", is.na(decision_date)) %>% nrow()

# 5. Country strings that normalise to NA
country_fail_examples <- log_country %>%
  filter(!is.na(raw) & raw != "" & is.na(normalised)) %>%
  arrange(desc(n_trials)) %>% slice_head(n = 30)
n_country_fail <- log_country %>%
  filter(!is.na(raw) & raw != "" & is.na(normalised)) %>%
  summarise(n = sum(n_trials)) %>% pull(n)

# 6. CTIS organ class codes not in lookup (normalised still numeric)
organ_unresolved_examples <- log_organ %>%
  filter(!is.na(normalised) & str_detect(normalised, "^\\d+$")) %>%
  arrange(desc(n_trials)) %>% slice_head(n = 20)
n_organ_unresolved <- log_organ %>%
  filter(!is.na(normalised) & str_detect(normalised, "^\\d+$")) %>%
  summarise(n = sum(n_trials)) %>% pull(n)

# 7. Duplicate canonical trial identities in final cache
ct_dupes_examples <- df %>%
  filter(!is.na(canonical_trial_id) & canonical_trial_id != "") %>%
  add_count(canonical_trial_id) %>%
  filter(n > 1) %>%
  arrange(canonical_trial_id) %>%
  select(`_id`, register, canonical_trial_id, CT_number, transition_eudract_number,
         sponsor_label, status)
n_ct_dupes <- n_distinct(ct_dupes_examples$canonical_trial_id)

# 8. Sponsor labels still containing legal suffixes (pipeline fallback / unknown path)
junk_pat <- regex("\\bGmbH\\b|\\bInc\\.?\\b|\\bLtd\\.?\\b|\\bS\\.A\\.\\b|\\bAG\\b|\\bB\\.V\\.\\b|\\bLLC\\b|\\bPLC\\b")
sponsor_junk_ex <- df %>%
  filter(!is.na(sponsor_label), str_detect(sponsor_label, junk_pat)) %>%
  count(sponsor_label, sort = TRUE) %>% slice_head(n = 20)
n_sponsor_junk <- df %>%
  filter(!is.na(sponsor_label), str_detect(sponsor_label, junk_pat)) %>% nrow()

# 9. Extreme days_to_decision (> 10 years)
days_extreme_examples <- df %>%
  filter(!is.na(days_to_decision), days_to_decision > 3650) %>%
  arrange(desc(days_to_decision)) %>%
  select(`_id`, register, sponsor_label, submission_date_parsed,
         decision_date, days_to_decision) %>% slice_head(n = 20)
n_days_extreme <- nrow(days_extreme_examples)

# 10. Implausible submission year
cur_yr <- year(Sys.Date())
year_implaus_ex <- df %>%
  filter(!is.na(year), year < 1990 | year > cur_yr) %>%
  select(`_id`, register, submission_date, submission_date_parsed, year, sponsor_label) %>%
  slice_head(n = 20)
n_year_implaus <- nrow(year_implaus_ex)

# 11. Status "Other" and orphan "Unknown" — counts for summary table
n_status_other   <- sum(df$status    == "Other",   na.rm = TRUE)
n_orphan_unknown <- sum(df$is_orphan == "Unknown",  na.rm = TRUE)

# Summary table
issue_tbl <- tribble(
  ~Severity,   ~Category,       ~Issue,                                                                          ~`N affected`,                ~`Suggested fix in app.R`,
  "High",      "Deduplication", "EUCTR-original + CTIS labelled EUCTR-era share same title_key (double-counted in chart)", fmt_n(n_within_euctr_overlap),"Re-check analysis-register CTIS records against EUCTR originals using canonical_trial_id and transition aliases",
  "High",      "Deduplication", "Residual cross-register duplicates (same title_key in EUCTR + CTIS cache)",     fmt_n(n_resid_dupes),         "Treat title-only matches as candidates unless transition/base identifiers also agree",
  "High",      "Deduplication", "Transitioned EUCTR with no CTIS counterpart found by current fallbacks",        fmt_n(n_unmatched),           "Re-fetch CTIS and inspect missing transition EudraCT identifiers",
  "High",      "Status",        "CTIS status codes outside the 1-6 mapping -> classified as 'Other'",            fmt_n(n_ctis_unmapped_status),"Add missing codes to ctis_status_map (~line 1023)",
  "High",      "Dates",         "Completed trials with no decision_date (breaks compliance calculations)",        fmt_n(n_completed_no_date),   "Add fallback date field for CTIS completed trials (~line 1091)",
  "Medium",    "Country",       "Country strings that normalise to NA (location data silently lost)",             fmt_n(n_country_fail),        "Add unresolved strings to COUNTRY_LOOKUP (~line 386)",
  "Medium",    "Organ class",   "CTIS organ class codes not in ctis_soc_lookup (stored as raw numbers)",          fmt_n(n_organ_unresolved),    "Extend ctis_soc_lookup (~line 22) with missing MedDRA SOC codes",
  "Medium",    "Identifiers",   "Duplicate canonical trial IDs in final cache",                                  fmt_n(n_ct_dupes),            "Review canonical_trial_id deduplication and transition aliases",
  "Medium",    "Normalisation", "Sponsor labels still contain legal-entity tokens (pipeline fallback / unknown)",  fmt_n(n_sponsor_junk),        "Add raw_sponsor string to manual_sponsor_aliases.csv and rebuild sponsor index",
  "Medium",    "Orphan",        "Orphan designation classified as 'Unknown' (unresolvable from source data)",    fmt_n(n_orphan_unknown),      "Improve CTIS orphan detection; extend EUCTR No-pattern",
  "Low",       "Status",        "Trials with status='Other' (no Ongoing/Completed pattern match)",               fmt_n(n_status_other),        "Review and extend ongoing_pat / completed_pat (~line 1012)",
  "Low",       "Dates",         "days_to_decision > 10 years (likely source data entry error)",                  fmt_n(n_days_extreme),        "Add upper cap: if_else(d > 3650, NA_real_, d) alongside negative guard (~line 1100)",
  "Low",       "Dates",         paste0("Implausible submission year (< 1990 or > ", cur_yr, ")"),                fmt_n(n_year_implaus),        "Set year to NA when outside [1990, current year] after parsing (~line 1070)"
)
```

## Summary

```{r issues-summary}
issue_tbl %>%
  datatable(
    rownames = FALSE,
    escape   = FALSE,
    filter   = "top",
    options  = list(pageLength = 20, scrollX = TRUE),
    caption  = "All data quality issues detected during pipeline audit"
  ) %>%
  formatStyle("Severity",
    backgroundColor = styleEqual(
      c("High", "Medium", "Low"),
      c("#FFEBEB", "#FFF3E0", "#FFFDE7")
    )
  )
```

## High: Residual Duplicates

```{r issue-residual-dupes}
if (nrow(cross_reg_tbl) > 0) {
  cross_reg_tbl %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(nrow(cross_reg_tbl)),
        " trials with matching title_key in both EUCTR and CTIS — should have been collapsed by dedup."))
} else {
  cat("No residual cross-register duplicates detected.")
}
```

**Root cause:** Title_key matching uses a 20-char minimum and exact 80-char match. Short or register-divergent titles escape detection, while generic long prefixes can also create false positives.

**Fix:** Keep title-only matches as review candidates unless a transition EudraCT number, CTIS base ID, protocol code, or other identifier also agrees.

## High: Unmatched Transitioned

```{r issue-unmatched-trans}
trans_matched %>%
  filter(is.na(ctis_id)) %>%
  select(`EUCTR base ID` = euctr_base,
         `Trial status`  = trial_status,
         `End status`    = end_status) %>%
  slice_head(n = 50) %>%
  datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
	    caption = paste0(fmt_n(n_unmatched),
	      " EUCTR trials marked transitioned in the DB with no CTIS counterpart found by transition number, base ID, or title_key."))
```

**Root cause:** Matching CTIS record (a) not yet fetched, (b) lacks a public transition EudraCT identifier, (c) has divergent title, or (d) EUCTR flag is a false positive.

**Fix:** Keep the fallback lookup by CTIS transition EudraCT number and run fresh `update_data.R` to ensure CTIS is current.

## High: Unmapped Status Codes

```{r issue-ctis-status}
if (nrow(unmapped_status_codes) > 0) {
  unmapped_status_codes %>%
    select(register, `Raw code` = raw, `Mapped to` = normalised, `N trials` = n_trials) %>%
    nk(cap = paste0("CTIS numeric status codes not in ctis_status_map, classified as Other. Affects ",
                    fmt_n(n_ctis_unmapped_status), " trials."))
} else {
  cat("All CTIS numeric status codes are mapped.")
}
```

**Fix in `app.R`:** Extend `ctis_status_map` around line 1023 with the codes shown above.

## High: Completed — No Decision Date

```{r issue-completed-no-date}
df %>%
  filter(status == "Completed", is.na(decision_date)) %>%
  count(register) %>%
  nk(cap = paste0(fmt_n(n_completed_no_date),
    " completed trials have no decision_date. Compliance overdue calculations are broken for these."))
```

**Fix in `app.R`:** In `decision_date` derivation (~line 1091), add a fallback for CTIS using the `decisionDate` field; also try deriving from the trial status change timestamp.

## Medium: Country Failures

```{r issue-country-fail}
if (nrow(country_fail_examples) > 0) {
  country_fail_examples %>%
    select(register, `Raw string` = raw, `N trials` = n_trials) %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(n_country_fail),
        " trial-country strings normalised to NA — location data silently lost."))
} else {
  cat("No country normalisation failures.")
}
```

**Fix in `app.R`:** Add the raw strings to `COUNTRY_LOOKUP` (~line 386).

## Medium: Unresolved Organ Codes

```{r issue-organ-codes}
if (nrow(organ_unresolved_examples) > 0) {
  organ_unresolved_examples %>%
    select(register, `Raw code` = raw, `Stored as` = normalised, `N trials` = n_trials) %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(n_organ_unresolved),
        " CTIS organ class codes not in ctis_soc_lookup — stored as raw numbers, breaking filters."))
} else {
  cat("All CTIS organ class codes resolved.")
}
```

**Fix in `app.R`:** Extend `ctis_soc_lookup` (~line 22) with missing MedDRA SOC code-label pairs.

## Medium: Duplicate CT Numbers

```{r issue-ct-dupes}
if (nrow(ct_dupes_examples) > 0) {
  ct_dupes_examples %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(n_ct_dupes), " canonical trial IDs appear more than once in the final cache."))
} else {
  cat("No duplicate CT numbers found.")
}
```

## Medium: Sponsor Name Junk

```{r issue-sponsor-junk}
if (nrow(sponsor_junk_ex) > 0) {
  sponsor_junk_ex %>%
    rename(`Sponsor label` = sponsor_label, `N trials` = n) %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(n_sponsor_junk),
        " trials with legal-entity tokens in sponsor_label (unknown/fallback path)."))
} else {
  cat("No residual legal suffixes found.")
}
```

**Fix:** Add the raw sponsor string to `config/sponsor_norm_pipeline/manual_sponsor_aliases.csv` and rebuild the sponsor index (`build_sponsor_index.R --no-ror` + `build_sponsor_labels.R --write-queue`).

## Low: Extreme Days-to-Decision

```{r issue-days-extreme}
if (nrow(days_extreme_examples) > 0) {
  days_extreme_examples %>%
    mutate(days_to_decision = fmt_n(round(days_to_decision))) %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(n_days_extreme),
        " trials with days_to_decision > 10 years — likely source data entry errors."))
} else {
  cat("No extreme days_to_decision values.")
}
```

**Fix in `app.R`:** Extend guard at ~line 1100: `if_else(!is.na(d) & d >= 0 & d <= 3650, d, NA_real_)`.

## Low: Implausible Year

```{r issue-year}
if (nrow(year_implaus_ex) > 0) {
  year_implaus_ex %>%
    datatable(rownames = FALSE, options = list(pageLength = 20, scrollX = TRUE),
      caption = paste0(fmt_n(n_year_implaus),
        " trials with submission year before 1990 or after ", cur_yr, "."))
} else {
  cat("No implausible submission years found.")
}
```

**Fix in `app.R`:** After `year = year(submission_date_parsed)` (~line 1070), add `year = if_else(year < 1990 | year > year(Sys.Date()), NA_integer_, year)`.

---

*Generated by `rmarkdown/preprocessing.Rmd` — `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`*
