---
title: "Data Preprocessing Report"
subtitle: "EU Paediatric Clinical Trials - Pipeline Audit"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    self_contained: true
    fig_width: 9
    fig_height: 4.5
params:
  cache_path: "../trials_cache.rds"
  log_dir:    "../data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 120
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(stringr)
  library(lubridate)
  library(scales)
  library(knitr)
  library(kableExtra)
  library(readr)
  library(DT)
})

has_pkg <- function(pkg) requireNamespace(pkg, quietly = TRUE)

# Palette
BLUE   <- "#5E81AC"; FROST  <- "#81A1C1"; TEAL   <- "#88C0D0"
GREEN  <- "#A3BE8C"; YELLOW <- "#EBCB8B"; ORANGE <- "#D08770"
RED    <- "#BF616A"; PURPLE <- "#B48EAD"; GRAY   <- "#4C566A"
LIGHT  <- "#ECEFF4"; DARK   <- "#2E3440"; PALE <- "#F8FAFC"
NORD8  <- c(BLUE, GREEN, YELLOW, RED, PURPLE, ORANGE, TEAL, FROST)

REG_COLS <- c(EUCTR = BLUE, CTIS = TEAL)
STATUS_COLS <- c(Ongoing = GREEN, Completed = BLUE, Withdrawn = ORANGE, Other = RED)

theme_pre <- function(base = 11) {
  theme_minimal(base_size = base, base_family = "sans") %+replace% theme(
    plot.title       = element_text(size = base + 2, face = "bold", color = DARK, hjust = 0),
    plot.subtitle    = element_text(size = base, color = GRAY, hjust = 0, margin = margin(t = 2, b = 6)),
    axis.title       = element_text(size = base, color = GRAY),
    axis.text        = element_text(size = base - 1, color = GRAY),
    axis.text.y      = element_text(margin = margin(r = 4)),
    panel.grid.major = element_line(color = LIGHT, linewidth = 0.45),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    legend.key.size  = unit(0.4, "cm"),
    legend.text      = element_text(size = base - 1, color = GRAY),
    plot.margin      = margin(8, 12, 8, 8),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.text       = element_text(size = base, face = "bold", color = DARK)
  )
}

fmt_n <- function(x) {
  if (length(x) == 0 || all(is.na(x))) return("NA")
  format(round(as.numeric(x)), big.mark = ",", scientific = FALSE, trim = TRUE)
}
fmt_pct <- function(x, digits = 1) {
  ifelse(is.na(x) | is.nan(x) | is.infinite(x), "NA", sprintf(paste0("%.", digits, "f%%"), x))
}
safe_div <- function(num, den) ifelse(is.na(den) | den == 0, NA_real_, num / den)
co_chr <- function(x, default = NA_character_) ifelse(is.na(x) | x == "", default, as.character(x))
has_cols <- function(x, cols) !is.null(x) && all(cols %in% names(x))
changed_flag <- function(x) if ("changed" %in% names(x)) x$changed %in% TRUE else rep(FALSE, nrow(x))
first_existing <- function(paths) {
  paths <- unique(paths[nzchar(paths)])
  hit <- paths[file.exists(paths)]
  if (length(hit) == 0) NA_character_ else normalizePath(hit[[1]], mustWork = TRUE)
}

find_project_root <- function() {
  starts <- unique(c(getwd(), dirname(normalizePath(knitr::current_input() %||% ".", mustWork = FALSE))))
  starts <- starts[nzchar(starts)]
  for (start in starts) {
    cur <- normalizePath(start, mustWork = FALSE)
    for (i in seq_len(8)) {
      if (file.exists(file.path(cur, "app.R")) && dir.exists(file.path(cur, "data"))) return(cur)
      parent <- dirname(cur)
      if (identical(parent, cur)) break
      cur <- parent
    }
  }
  normalizePath(getwd(), mustWork = TRUE)
}

`%||%` <- function(x, y) if (is.null(x) || length(x) == 0 || is.na(x) || !nzchar(x)) y else x

project_root <- find_project_root()
rmd_dir <- dirname(normalizePath(knitr::current_input() %||% "rmarkdown/preprocessing.Rmd", mustWork = FALSE))

resolve_file <- function(path, fallback = NULL) {
  if (is.null(path) || is.na(path) || !nzchar(path)) path <- fallback
  first_existing(c(
    path,
    file.path(getwd(), path),
    file.path(rmd_dir, path),
    file.path(project_root, path),
    file.path(project_root, basename(path)),
    fallback,
    file.path(project_root, fallback %||% "")
  ))
}

resolve_dir <- function(path, fallback = "data") {
  candidates <- unique(c(
    path,
    file.path(getwd(), path),
    file.path(rmd_dir, path),
    file.path(project_root, path),
    file.path(project_root, basename(path)),
    file.path(project_root, fallback)
  ))
  hit <- candidates[dir.exists(candidates)]
  if (length(hit) == 0) NA_character_ else normalizePath(hit[[1]], mustWork = TRUE)
}

cache_file <- resolve_file(params$cache_path, "trials_cache.rds")
log_dir <- resolve_dir(params$log_dir, "data")
db_path <- resolve_file("data/trials.sqlite", file.path("data", "trials.sqlite"))

if (is.na(cache_file)) stop("Cache not found. Tried param cache_path and project-root trials_cache.rds.")

t0 <- proc.time()
df <- readRDS(cache_file)
load_sec <- round((proc.time() - t0)[["elapsed"]], 1)
df <- tibble::as_tibble(df)

required_defaults <- list(
  DIMP_inn_name = NA_character_,
  transition_eudract_number = NA_character_,
  trial_identifiers = if ("CT_number" %in% names(df)) df$CT_number else NA_character_,
  analysis_register = if ("register" %in% names(df)) df$register else NA_character_,
  sponsor_label = if ("sponsor_name" %in% names(df)) df$sponsor_name else NA_character_,
  canonical_trial_id = if ("_id" %in% names(df)) df$`_id` else NA_character_
)
for (nm in names(required_defaults)) {
  if (!nm %in% names(df)) df[[nm]] <- required_defaults[[nm]]
}

read_log <- function(name) {
  p <- file.path(log_dir, paste0(name, "_normalisation_log.csv"))
  if (!is.na(log_dir) && file.exists(p)) readr::read_csv(p, show_col_types = FALSE) else NULL
}
read_optional_csv <- function(...) {
  p <- file.path(...)
  if (file.exists(p)) readr::read_csv(p, show_col_types = FALSE) else NULL
}

log_country  <- read_log("country")
log_meddra   <- read_log("meddra_term")
log_organ    <- read_log("organ_class")
log_phase    <- read_log("phase")
log_status_c <- read_log("status_category")
log_status_d <- read_log("status_display")
spon_log_raw <- read_log("sponsor")
sub_log      <- read_log("substance")
spon_labels  <- read_optional_csv(log_dir, "trial_sponsor_labels.csv")
sub_labels   <- read_optional_csv(log_dir, "trial_substance_labels.csv")
spon_raw     <- read_optional_csv(log_dir, "trial_sponsors_raw.csv")
sub_raw      <- read_optional_csv(log_dir, "trial_substances_raw.csv")
spon_queue   <- read_optional_csv(project_root, "config", "sponsor_norm_pipeline", "sponsor_review_queue.csv")
sub_queue    <- read_optional_csv(project_root, "config", "substance_norm_pipeline", "substance_review_queue.csv")

normalise_sponsor_log <- function(x) {
  if (is.null(x)) return(NULL)
  out <- x
  if (!"raw_sponsor" %in% names(out) && "raw" %in% names(out)) out$raw_sponsor <- out$raw
  if (!"sponsor_clean" %in% names(out) && "normalised" %in% names(out)) out$sponsor_clean <- out$normalised
  if (!"match_status" %in% names(out)) {
    out$match_status <- dplyr::case_when(
      "sponsor_clean" %in% names(out) & !is.na(out$sponsor_clean) & out$sponsor_clean != "" ~ "accepted",
      TRUE ~ "unknown"
    )
  }
  if (!"match_score" %in% names(out)) out$match_score <- NA_real_
  if (!"match_source" %in% names(out)) {
    out$match_source <- if ("changed" %in% names(out)) ifelse(out$changed, "normalised", "unchanged") else NA_character_
  }
  if (!"match_reason" %in% names(out)) out$match_reason <- NA_character_
  if (!"n_trials" %in% names(out)) out$n_trials <- 1L
  out
}
spon_log <- normalise_sponsor_log(spon_log_raw)

waterfall_plot <- function(wf, title, subtitle = NULL) {
  wf <- wf |>
    mutate(step = factor(step, levels = step))
  ggplot(wf, aes(step, n, fill = type)) +
    geom_col(width = 0.58, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.25, color = DARK) +
    geom_text(
      aes(label = ifelse(!is.na(delta) & delta > 0, paste0("-", fmt_n(delta)), "")),
      y = 0,
      vjust = 1.3,
      size = 3.05,
      color = RED
    ) +
    scale_fill_manual(values = c(start = BLUE, step = RED, end = GREEN, label = TEAL), na.value = GRAY) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0.08, 0.13))) +
    labs(title = title, subtitle = subtitle, x = NULL, y = "Rows / strings remaining") +
    theme_pre() +
    theme(axis.text.x = element_text(angle = 16, hjust = 1))
}

dt_tbl <- function(x, caption = NULL, page = 20, filter = "none") {
  DT::datatable(
    x,
    rownames = FALSE,
    filter = filter,
    options = list(pageLength = page, scrollX = TRUE, autoWidth = TRUE),
    caption = caption
  )
}

nk <- function(x, cap = NULL, cn = colnames(x)) {
  knitr::kable(x, booktabs = TRUE, linesep = "", caption = cap, col.names = cn, escape = FALSE) |>
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left",
      font_size = 13
    )
}

kpi_box <- function(value, label, colour = BLUE) {
  sprintf(
    '<div class="kpi" style="border-left-color:%s"><div class="kpi-value">%s</div><div class="kpi-label">%s</div></div>',
    colour, value, label
  )
}

kpi_row <- function(...) {
  paste0('<div class="kpi-row">', paste0(..., collapse = ""), '</div>')
}

notice <- function(title, text, type = "note") {
  cls <- paste("audit-notice", type)
  sprintf('<div class="%s"><strong>%s</strong><br>%s</div>', cls, title, text)
}

make_title_key <- function(title) {
  title |>
    as.character() |>
    tolower() |>
    str_replace_all("[^a-z0-9 ]", " ") |>
    str_squish() |>
    substr(1, 80)
}

is_euctr_id <- function(x) str_detect(coalesce(as.character(x), ""), "^[0-9]{4}-[0-9]{6}-[0-9]{2}-[A-Z]")
is_ctis_id <- function(x) !is_euctr_id(x)
```

```{css, echo=FALSE}
body {
  color: #2E3440;
}
.main-container {
  max-width: 1180px;
}
h1, h2, h3 {
  letter-spacing: 0;
}
.kpi-row {
  display: flex;
  flex-wrap: wrap;
  gap: 12px;
  margin: 12px 0 18px;
}
.kpi {
  background: #F8FAFC;
  border: 1px solid #E5E9F0;
  border-left: 5px solid #5E81AC;
  border-radius: 6px;
  min-width: 150px;
  padding: 10px 14px;
}
.kpi-value {
  font-size: 1.45rem;
  font-weight: 700;
  line-height: 1.15;
  color: #2E3440;
}
.kpi-label {
  margin-top: 3px;
  font-size: 0.88rem;
  color: #4C566A;
}
.audit-notice {
  border-radius: 6px;
  padding: 12px 14px;
  margin: 12px 0 16px;
  border: 1px solid #D8DEE9;
  background: #F8FAFC;
}
.audit-notice.warn {
  border-color: #EBCB8B;
  background: #FFF8E6;
}
.audit-notice.fail {
  border-color: #BF616A;
  background: #FFF0F0;
}
.section-note {
  color: #4C566A;
  max-width: 880px;
}
.dataTables_wrapper {
  margin-bottom: 22px;
}
```

# 1. Dataset Overview

```{r overview-kpis, results='asis'}
n_tot <- nrow(df)
n_cols <- ncol(df)
reg_counts <- df |> count(register, name = "n")
n_euctr <- reg_counts$n[match("EUCTR", reg_counts$register)] %||% 0
n_ctis  <- reg_counts$n[match("CTIS", reg_counts$register)] %||% 0
n_analysis_euctr <- sum(df$analysis_register == "EUCTR", na.rm = TRUE)
n_analysis_ctis  <- sum(df$analysis_register == "CTIS", na.rm = TRUE)

cat(kpi_row(
  kpi_box(fmt_n(n_tot), "trials in final cache", BLUE),
  kpi_box(fmt_n(n_euctr), "source EUCTR", FROST),
  kpi_box(fmt_n(n_ctis), "source CTIS", TEAL),
  kpi_box(fmt_n(n_cols), "columns", GREEN),
  kpi_box(paste0(load_sec, " s"), "cache load time", PURPLE)
))
```

The report audits the cache at `r cache_file`. Logs are read from `r log_dir`; the SQLite source database is `r ifelse(is.na(db_path), "not available", db_path)`.

## Register Mix

```{r overview-registers, fig.height=3.6}
df |>
  count(register, name = "n") |>
  ggplot(aes(register, n, fill = register)) +
  geom_col(width = 0.58, show.legend = FALSE) +
  geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.5, color = DARK) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.14))) +
  labs(title = "Trials by source register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r overview-analysis-registers, fig.height=3.6}
df |>
  count(analysis_register, name = "n") |>
  mutate(analysis_register = coalesce(analysis_register, "Missing")) |>
  ggplot(aes(analysis_register, n, fill = analysis_register)) +
  geom_col(width = 0.58, show.legend = FALSE) +
  geom_text(aes(label = fmt_n(n)), vjust = -0.35, size = 3.5, color = DARK) +
  scale_fill_manual(values = c(REG_COLS, Missing = GRAY), na.value = GRAY) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.14))) +
  labs(title = "Trials by analysis register", subtitle = "Migrated CTIS records with EudraCT transition numbers are analysed as EUCTR-era.", x = NULL, y = "Trials") +
  theme_pre()
```

## Key Field Completeness

```{r overview-key-completeness}
key_cols <- c(
  "_id", "CT_number", "Full_title", "register", "analysis_register",
  "transition_eudract_number", "canonical_trial_id", "Member_state",
  "age_group", "sponsor_label", "sponsor_type", "phase", "status",
  "submission_date_parsed", "decision_date", "days_to_decision",
  "has_PIP", "has_results", "is_orphan", "substance_label"
)
key_cols <- intersect(key_cols, names(df))

tibble(
  Column = key_cols,
  Type = vapply(df[key_cols], function(x) class(x)[1], character(1)),
  `Non-missing` = vapply(df[key_cols], function(x) sum(!(is.na(x) | (is.character(x) & x == ""))), integer(1)),
  `Complete (%)` = round(`Non-missing` / nrow(df) * 100, 1)
) |>
  arrange(`Complete (%)`) |>
  dt_tbl(caption = "Completeness of key fields in the final cache", page = 25)
```

# 2. Deduplication Pipeline

The dedup audit compares raw SQLite row identifiers with the final cache. EUCTR country variants and CTIS amendments are counted directly from raw IDs; cross-register and final removals are the remaining difference between the deduplicated raw universe and the cache.

```{r dedup-compute}
db_available <- !is.na(db_path) && file.exists(db_path) && has_pkg("DBI") && has_pkg("RSQLite")

if (db_available) {
  con <- DBI::dbConnect(RSQLite::SQLite(), db_path)
  all_ids <- DBI::dbGetQuery(con, 'SELECT "_id" FROM trials')[["_id"]]
  trans_df <- DBI::dbGetQuery(con, '
    SELECT "_id",
      json_extract(json, "$.x5_trial_status") AS trial_status,
      json_extract(json, "$.p_end_of_trial_status") AS end_status,
      json_extract(json, "$.a3_full_title_of_the_trial") AS title
    FROM trials
    WHERE lower(json_extract(json, "$.x5_trial_status")) LIKE "%transition%"
       OR lower(json_extract(json, "$.p_end_of_trial_status")) LIKE "%transition%"
  ')
  DBI::dbDisconnect(con)
} else {
  all_ids <- character(0)
  trans_df <- tibble(`_id` = character(), trial_status = character(), end_status = character(), title = character())
}

euctr_all <- all_ids[is_euctr_id(all_ids)]
ctis_all <- all_ids[is_ctis_id(all_ids)]
euctr_base_all <- str_replace(euctr_all, "-[A-Z]{2,3}$", "")
ctis_base_all <- str_replace(ctis_all, "-[0-9]{2}$", "")

n_raw <- length(all_ids)
n_euctr_base <- n_distinct(euctr_base_all)
n_ctis_base <- n_distinct(ctis_base_all)
n_ctis_amended <- if (length(ctis_base_all) > 0) sum(table(ctis_base_all) > 1) else 0
n_step1 <- n_euctr_base + length(ctis_all)
n_step2 <- n_euctr_base + n_ctis_base
n_final <- nrow(df)
n_drop_country <- n_raw - n_step1
n_drop_ctis <- n_step1 - n_step2
n_drop_final <- n_step2 - n_final
n_migrated <- sum(df$register == "CTIS" & df$analysis_register == "EUCTR", na.rm = TRUE)

wf <- tibble(
  step = factor(
    c("Raw DB rows", "After EUCTR country collapse", "After CTIS latest amendment", "After cross-register/final dedup", "Final cache"),
    levels = c("Raw DB rows", "After EUCTR country collapse", "After CTIS latest amendment", "After cross-register/final dedup", "Final cache")
  ),
  remaining = c(n_raw, n_step1, n_step2, n_final, n_final),
  removed = c(NA_integer_, n_drop_country, n_drop_ctis, n_drop_final, 0L),
  type = c("start", "step", "step", "step", "end")
)

wf_ok <- db_available && n_raw > 0 && all(!is.na(wf$remaining)) && all(diff(wf$remaining) <= 0) && all(wf$remaining >= 0)
```

## Waterfall

```{r dedup-waterfall, fig.height=4.6, results='asis'}
if (!db_available) {
  cat(notice("SQLite database unavailable", "The waterfall needs data/trials.sqlite. The final cache can still be audited, but raw-row dedup counts are skipped.", "warn"))
} else if (!wf_ok) {
  cat(notice("Waterfall counts are inconsistent", "The report will show the raw stage table instead of a plot because counts are not monotonic.", "fail"))
  print(nk(wf, cap = "Raw waterfall stage counts"))
} else {
  ggplot(wf, aes(step, remaining, fill = type)) +
    geom_col(width = 0.58, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(remaining)), vjust = -0.35, size = 3.4, color = DARK) +
    geom_text(
      data = wf |> filter(!is.na(removed) & removed > 0),
      aes(label = paste0("-", fmt_n(removed))),
      y = 0,
      vjust = 1.3,
      size = 3.2,
      color = RED,
      inherit.aes = TRUE
    ) +
    scale_fill_manual(values = c(start = BLUE, step = RED, end = GREEN)) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0.08, 0.12))) +
    labs(
      title = "Deduplication waterfall: rows remaining after each step",
      subtitle = "Red labels at the baseline show rows removed at that stage.",
      x = NULL,
      y = "Rows remaining"
    ) +
    theme_pre() +
    theme(axis.text.x = element_text(angle = 18, hjust = 1))
}
```

```{r dedup-table}
dedup_table <- tibble(
  Step = c("Raw DB", "EUCTR country collapse", "CTIS latest amendment", "Cross-register / final dedup", "Analysis label"),
  Logic = c(
    "All raw records fetched from SQLite",
    "Strip EUCTR member-state suffix and keep one base trial",
    "Strip CTIS amendment suffix and keep the latest version",
    "Prefer CTIS over duplicate EUCTR records and apply final cache exclusions",
    "Set analysis_register to EUCTR for migrated CTIS records"
  ),
  `Rows removed` = c("NA", fmt_n(n_drop_country), fmt_n(n_drop_ctis), fmt_n(n_drop_final), "0"),
  `Rows remaining` = c(fmt_n(n_raw), fmt_n(n_step1), fmt_n(n_step2), fmt_n(n_final), fmt_n(n_final))
)
nk(dedup_table, cap = "Deduplication stages")
```

## EUCTR Country Variants

```{r dedup-euctr-table}
if (!db_available) {
  cat("Raw DB not available.")
} else {
  cache_euctr_base <- df |>
    filter(is_euctr_id(`_id`)) |>
    transmute(base = str_replace(`_id`, "-[A-Z]{2,3}$", "")) |>
    pull(base)

  euctr_variant_tbl <- tibble(raw_id = euctr_all) |>
    mutate(base = str_replace(raw_id, "-[A-Z]{2,3}$", ""),
           country = str_extract(raw_id, "[A-Z]{2,3}$")) |>
    count(base, country, name = "n") |>
    group_by(base) |>
    summarise(
      `Countries in raw DB` = n_distinct(country),
      `Country variants` = paste(sort(unique(country)), collapse = ", "),
      `Kept in cache` = base %in% cache_euctr_base,
      .groups = "drop"
    ) |>
    filter(`Kept in cache`) |>
    arrange(desc(`Countries in raw DB`)) |>
    slice_head(n = 20) |>
    rename(`Base trial ID` = base)
  dt_tbl(euctr_variant_tbl, caption = "EUCTR base trials with the most country-specific raw rows", page = 20)
}
```

```{r dedup-euctr-dist, fig.height=3.5}
if (db_available && length(euctr_all) > 0) {
  tibble(raw_id = euctr_all) |>
    mutate(base = str_replace(raw_id, "-[A-Z]{2,3}$", "")) |>
    count(base, name = "country_rows") |>
    count(country_rows, name = "trials") |>
    filter(country_rows <= 30) |>
    ggplot(aes(country_rows, trials)) +
    geom_col(fill = BLUE, width = 0.8) +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = pretty_breaks()) +
    labs(title = "Distribution of EUCTR country rows per trial", x = "Country-specific rows", y = "Trials") +
    theme_pre()
}
```

## CTIS Amendment Versions

```{r dedup-ctis-table}
if (!db_available) {
  cat("Raw DB not available.")
} else {
  ctis_version_tbl <- tibble(raw_id = ctis_all) |>
    mutate(base = str_replace(raw_id, "-[0-9]{2}$", ""),
           version = suppressWarnings(as.integer(str_extract(raw_id, "[0-9]{2}$")))) |>
    group_by(base) |>
    filter(n() > 1) |>
    summarise(
      `Versions in DB` = n(),
      `Latest version` = max(version, na.rm = TRUE),
      `Latest raw ID` = raw_id[which.max(version)],
      `Latest kept in cache` = `Latest raw ID` %in% df$`_id`,
      `All versions` = paste(sort(raw_id), collapse = " -> "),
      .groups = "drop"
    ) |>
    arrange(desc(`Versions in DB`)) |>
    slice_head(n = 25) |>
    rename(`Base CTIS ID` = base)
  dt_tbl(ctis_version_tbl, caption = paste0("CTIS trials with multiple amendment versions (", fmt_n(n_ctis_amended), " base trials)"), page = 25)
}
```

```{r dedup-ctis-dist, fig.height=3.4}
if (db_available && length(ctis_all) > 0) {
  tibble(raw_id = ctis_all) |>
    mutate(base = str_replace(raw_id, "-[0-9]{2}$", "")) |>
    count(base, name = "versions") |>
    count(versions, name = "trials") |>
    ggplot(aes(versions, trials)) +
    geom_col(fill = TEAL, width = 0.7) +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = pretty_breaks()) +
    labs(title = "Distribution of CTIS amendment versions", x = "Versions in raw DB", y = "Base trials") +
    theme_pre()
}
```

## Cross-Register Checks

```{r dedup-cross-register}
ctis_by_tk <- df |>
  filter(register == "CTIS") |>
  mutate(title_key_check = make_title_key(Full_title)) |>
  filter(!is.na(title_key_check), nchar(title_key_check) >= 20) |>
  select(ctis_id = `_id`, title_key_check, ctis_title = Full_title, ctis_status = status)

euctr_by_tk <- df |>
  filter(register == "EUCTR") |>
  mutate(title_key_check = make_title_key(Full_title)) |>
  filter(!is.na(title_key_check), nchar(title_key_check) >= 20) |>
  select(euctr_id = `_id`, title_key_check, euctr_title = Full_title, euctr_status = status)

cross_reg_tbl <- ctis_by_tk |>
  inner_join(euctr_by_tk, by = "title_key_check") |>
  filter(ctis_id != euctr_id) |>
  select(
    `EUCTR ID` = euctr_id,
    `CTIS ID` = ctis_id,
    `Shared title key` = title_key_check,
    `EUCTR status` = euctr_status,
    `CTIS status` = ctis_status
  ) |>
  slice_head(n = 100)

trans_matched <- tibble()
if (db_available && nrow(trans_df) > 0) {
  ctis_transition <- df |>
    filter(register == "CTIS", !is.na(transition_eudract_number), transition_eudract_number != "") |>
    separate_rows(transition_eudract_number, sep = " / ") |>
    mutate(transition_eudract_number = str_trim(transition_eudract_number)) |>
    distinct(transition_eudract_number, ctis_id = `_id`, ctis_title = Full_title, ctis_status = status)

  trans_matched <- trans_df |>
    as_tibble() |>
    rename(euctr_id = `_id`) |>
    mutate(
      euctr_base = str_replace(euctr_id, "-[A-Z]{2,3}$", ""),
      title_key_check = make_title_key(title)
    ) |>
    distinct(euctr_base, title_key_check, trial_status, end_status) |>
    left_join(ctis_transition, by = c("euctr_base" = "transition_eudract_number")) |>
    left_join(ctis_by_tk, by = "title_key_check") |>
    mutate(
      matched_ctis_id = coalesce(ctis_id.x, ctis_id.y),
      match_method = case_when(
        !is.na(ctis_id.x) ~ "transition EudraCT number",
        !is.na(ctis_id.y) ~ "title key",
        TRUE ~ NA_character_
      )
    ) |>
    distinct(euctr_base, .keep_all = TRUE)
}
n_resid_dupes <- nrow(cross_reg_tbl)
n_unmatched_transitioned <- if (nrow(trans_matched) > 0) sum(is.na(trans_matched$matched_ctis_id)) else NA_integer_
```

```{r dedup-cross-table}
if (nrow(cross_reg_tbl) > 0) {
  dt_tbl(cross_reg_tbl, caption = "Potential residual cross-register duplicates in the final cache", page = 25)
} else {
  cat("No residual EUCTR/CTIS title-key duplicates were detected in the final cache.")
}
```

```{r dedup-transition-table}
if (!db_available) {
  cat("Raw DB not available.")
} else if (nrow(trans_matched) == 0) {
  cat("No transitioned EUCTR rows were found in the raw DB status fields.")
} else {
  trans_matched |>
    mutate(
      `EUCTR status` = coalesce(end_status, trial_status),
      `CTIS match` = coalesce(matched_ctis_id, "No CTIS match found")
    ) |>
    select(`EUCTR base ID` = euctr_base, `EUCTR status`, `CTIS match`, `Match method` = match_method) |>
    arrange(`CTIS match` == "No CTIS match found", `EUCTR base ID`) |>
    dt_tbl(caption = "EUCTR rows marked as transitioned and their CTIS match status", page = 25)
}
```

# 3. Register, Age, Phase, Status, and Dates

## Age Group

```{r age-kpis, results='asis'}
age_counts <- df |> count(age_group, name = "n") |> mutate(pct = n / sum(n) * 100)
n_paediatric <- sum(df$age_group %in% c("Paediatric", "Paediatric & Adult"), na.rm = TRUE)
n_adult <- sum(df$age_group %in% c("Adult", "Paediatric & Adult"), na.rm = TRUE)
cat(kpi_row(
  kpi_box(fmt_n(n_paediatric), "eligible under 18", BLUE),
  kpi_box(fmt_n(n_adult), "eligible 18 or older", GREEN),
  kpi_box(fmt_n(sum(df$age_group == "Unknown", na.rm = TRUE)), "unknown age group", RED)
))
```

```{r age-chart, fig.height=3.8}
df |>
  count(register, age_group, name = "n") |>
  mutate(age_group = coalesce(age_group, "Unknown")) |>
  ggplot(aes(register, n, fill = age_group)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = ifelse(n > max(n) * 0.05, fmt_n(n), "")), position = position_stack(vjust = 0.5), size = 3, color = "white") +
  scale_fill_manual(values = c(Paediatric = BLUE, Adult = ORANGE, `Paediatric & Adult` = GREEN, Unknown = GRAY), na.value = GRAY) +
  scale_y_continuous(labels = comma) +
  labs(title = "Age-group classification by register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r age-table}
age_counts |>
  mutate(`Trials` = fmt_n(n), `% of cache` = fmt_pct(pct)) |>
  select(`Age group` = age_group, Trials, `% of cache`) |>
  nk(cap = "Adult / paediatric classification")
```

## Phase

```{r phase-chart, fig.height=4}
df |>
  mutate(phase = coalesce(phase, "Unknown")) |>
  count(register, phase, name = "n") |>
  ggplot(aes(phase, n, fill = register)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.62) +
  scale_fill_manual(values = REG_COLS) +
  scale_y_continuous(labels = comma) +
  labs(title = "Trial phase by register", x = NULL, y = "Trials") +
  theme_pre() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r phase-log-table}
if (!is.null(log_phase) && has_cols(log_phase, c("raw", "normalised", "n_trials"))) {
  log_phase |>
    arrange(desc(n_trials)) |>
    select(Register = register, Raw = raw, Normalised = normalised, `N trials` = n_trials) |>
    slice_head(n = 30) |>
    dt_tbl(caption = "Most common raw phase values and parsed labels", page = 30)
} else {
  cat("Phase normalisation log not available.")
}
```

## Status

```{r status-chart, fig.height=3.8}
df |>
  mutate(status = coalesce(status, "Other")) |>
  count(register, status, name = "n") |>
  ggplot(aes(register, n, fill = status)) +
  geom_col(width = 0.62) +
  scale_fill_manual(values = STATUS_COLS, na.value = GRAY) +
  scale_y_continuous(labels = comma) +
  labs(title = "Derived trial status by register", x = NULL, y = "Trials") +
  theme_pre()
```

```{r status-log-table}
if (!is.null(log_status_c) && has_cols(log_status_c, c("raw", "normalised", "n_trials"))) {
  log_status_c |>
    filter(register == "CTIS" | changed_flag(log_status_c)) |>
    arrange(register, desc(n_trials)) |>
    select(Register = register, Raw = raw, Normalised = normalised, `N trials` = n_trials) |>
    slice_head(n = 40) |>
    dt_tbl(caption = "Status category mapping audit", page = 25)
} else {
  cat("Status normalisation log not available.")
}
```

## Dates and Decision Lag

```{r date-kpis, results='asis'}
n_decision_missing <- sum(is.na(df$decision_date))
n_completed_no_decision <- sum(df$status == "Completed" & is.na(df$decision_date), na.rm = TRUE)
median_days <- median(df$days_to_decision, na.rm = TRUE)
cat(kpi_row(
  kpi_box(fmt_n(sum(!is.na(df$submission_date_parsed))), "parsed submission dates", BLUE),
  kpi_box(fmt_n(sum(!is.na(df$decision_date))), "decision dates", GREEN),
  kpi_box(fmt_n(n_completed_no_decision), "completed with no decision date", RED),
  kpi_box(fmt_n(median_days), "median days to decision", PURPLE)
))
```

```{r year-chart, fig.height=4.2}
df |>
  filter(!is.na(year), year >= 1990, year <= year(Sys.Date())) |>
  count(year, analysis_register, name = "n") |>
  ggplot(aes(year, n, fill = analysis_register)) +
  geom_col(width = 0.9) +
  scale_fill_manual(values = REG_COLS, na.value = GRAY) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(labels = comma) +
  labs(title = "Trials by submission/analysis year", x = NULL, y = "Trials") +
  theme_pre()
```

```{r days-chart, fig.height=4}
df |>
  filter(!is.na(days_to_decision), days_to_decision >= 0, days_to_decision <= 3650) |>
  mutate(log_days = log10(days_to_decision + 1)) |>
  ggplot(aes(log_days, fill = register)) +
  geom_histogram(bins = 45, alpha = 0.72, position = "identity") +
  scale_fill_manual(values = REG_COLS) +
  scale_x_continuous(labels = function(x) fmt_n(10^x - 1)) +
  labs(title = "Days from submission to decision", subtitle = "X-axis is log-scaled for readability.", x = "Days to decision", y = "Trials") +
  theme_pre()
```

# 4. Field Normalisation Audits

## Country

```{r country-kpis, results='asis'}
if (!is.null(log_country) && has_cols(log_country, c("raw", "normalised", "n_trials"))) {
  n_country_rows <- sum(log_country$n_trials, na.rm = TRUE)
  n_country_raw <- n_distinct(log_country$raw)
  n_country_norm <- n_distinct(log_country$normalised, na.rm = TRUE)
  n_country_changed <- sum(log_country$n_trials[changed_flag(log_country)], na.rm = TRUE)
  cat(kpi_row(
    kpi_box(fmt_n(n_country_rows), "country entries", BLUE),
    kpi_box(fmt_n(n_country_raw), "unique raw values", FROST),
    kpi_box(fmt_n(n_country_norm), "unique normalised values", GREEN),
    kpi_box(fmt_n(n_country_changed), "entries changed", ORANGE)
  ))
} else {
  cat(notice("Country log unavailable", "No country normalisation log was found for this render.", "warn"))
}
```

```{r country-chart, fig.height=5}
if ("Member_state" %in% names(df)) {
  df |>
    filter(!is.na(Member_state), Member_state != "") |>
    separate_rows(Member_state, sep = " / ") |>
    mutate(Member_state = str_squish(Member_state)) |>
    filter(Member_state != "") |>
    count(Member_state, sort = TRUE, name = "n") |>
    slice_head(n = 20) |>
    mutate(Member_state = reorder(Member_state, n)) |>
    ggplot(aes(n, Member_state, fill = n)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    scale_fill_gradient(low = FROST, high = BLUE) +
    labs(title = "Top countries in final cache", x = "Trials", y = NULL) +
    theme_pre()
}
```

## MedDRA Terms

```{r meddra-kpis, results='asis'}
if (!is.null(log_meddra) && has_cols(log_meddra, c("raw", "normalised", "n_trials"))) {
  cat(kpi_row(
    kpi_box(fmt_n(sum(log_meddra$n_trials, na.rm = TRUE)), "MedDRA term entries", BLUE),
    kpi_box(fmt_n(n_distinct(log_meddra$raw)), "unique raw values", FROST),
    kpi_box(fmt_n(n_distinct(log_meddra$normalised, na.rm = TRUE)), "unique normalised values", GREEN),
    kpi_box(fmt_n(sum(log_meddra$n_trials[changed_flag(log_meddra)], na.rm = TRUE)), "entries changed", ORANGE)
  ))
} else {
  cat(notice("MedDRA term log unavailable", "No MedDRA term normalisation log was found for this render.", "warn"))
}
```

```{r meddra-chart, fig.height=5}
if ("MEDDRA_term" %in% names(df)) {
  df |>
    filter(!is.na(MEDDRA_term), MEDDRA_term != "") |>
    separate_rows(MEDDRA_term, sep = " / ") |>
    mutate(MEDDRA_term = str_squish(MEDDRA_term)) |>
    count(MEDDRA_term, sort = TRUE, name = "n") |>
    slice_head(n = 20) |>
    mutate(MEDDRA_term = reorder(MEDDRA_term, n)) |>
    ggplot(aes(n, MEDDRA_term, fill = n)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    scale_fill_gradient(low = FROST, high = BLUE) +
    labs(title = "Most common MedDRA terms", x = "Trials", y = NULL) +
    theme_pre()
}
```

## Organ Class

```{r organ-kpis, results='asis'}
if (!is.null(log_organ) && has_cols(log_organ, c("raw", "normalised", "n_trials"))) {
  cat(kpi_row(
    kpi_box(fmt_n(sum(log_organ$n_trials, na.rm = TRUE)), "organ-class entries", BLUE),
    kpi_box(fmt_n(n_distinct(log_organ$raw)), "unique raw values", FROST),
    kpi_box(fmt_n(n_distinct(log_organ$normalised, na.rm = TRUE)), "unique normalised values", GREEN),
    kpi_box(fmt_n(sum(log_organ$n_trials[changed_flag(log_organ)], na.rm = TRUE)), "entries changed", ORANGE)
  ))
} else {
  cat(notice("Organ-class log unavailable", "No organ-class normalisation log was found for this render.", "warn"))
}
```

```{r organ-chart, fig.height=5}
if ("MEDDRA_organ_class" %in% names(df)) {
  df |>
    filter(!is.na(MEDDRA_organ_class), MEDDRA_organ_class != "") |>
    separate_rows(MEDDRA_organ_class, sep = " / ") |>
    mutate(MEDDRA_organ_class = str_squish(MEDDRA_organ_class)) |>
    count(MEDDRA_organ_class, sort = TRUE, name = "n") |>
    slice_head(n = 18) |>
    mutate(MEDDRA_organ_class = reorder(MEDDRA_organ_class, n)) |>
    ggplot(aes(n, MEDDRA_organ_class, fill = n)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = fmt_n(n)), hjust = -0.1, size = 3, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    scale_fill_gradient(low = FROST, high = BLUE) +
    labs(title = "Most common MedDRA organ classes", x = "Trials", y = NULL) +
    theme_pre()
}
```

## Sponsor Names

The sponsor audit follows the cache-build pipeline: exported trial-sponsor pairs, unique raw sponsor strings, normalised unique strings, accepted/review labels joined back to trials, and optional review queue rows.

```{r sponsor-audit-compute}
spon_has_rich_log <- !is.null(spon_log_raw) && has_cols(
  spon_log_raw,
  c("raw_sponsor", "sponsor_clean", "match_status", "match_score", "match_source")
)

spon_unique <- if (!is.null(spon_log) && "raw_sponsor" %in% names(spon_log)) {
  spon_log |>
    mutate(changed_for_audit = changed_flag(spon_log)) |>
    group_by(raw_sponsor) |>
    summarise(
      sponsor_clean = first(na.omit(sponsor_clean)),
      match_status = first(match_status),
      match_score = suppressWarnings(max(as.numeric(match_score), na.rm = TRUE)),
      match_source = first(na.omit(match_source)),
      match_reason = first(na.omit(match_reason)),
      n_trials = sum(n_trials, na.rm = TRUE),
      changed = any(changed_for_audit, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(match_score = ifelse(is.infinite(match_score), NA_real_, match_score))
} else {
  tibble()
}

spon_raw_pairs <- if (!is.null(spon_raw) && has_cols(spon_raw, c("_id", "raw_sponsor"))) nrow(spon_raw) else NA_integer_
spon_raw_trials <- if (!is.null(spon_raw) && "_id" %in% names(spon_raw)) n_distinct(spon_raw$`_id`) else NA_integer_
spon_unique_raw <- if (nrow(spon_unique) > 0) n_distinct(spon_unique$raw_sponsor) else NA_integer_
spon_accepted_strings <- if (nrow(spon_unique) > 0) sum(spon_unique$match_status %in% c("accepted", "review"), na.rm = TRUE) else NA_integer_
spon_label_rows <- if (!is.null(spon_labels) && "_id" %in% names(spon_labels)) nrow(spon_labels) else NA_integer_
spon_label_trials <- if (!is.null(spon_labels) && "_id" %in% names(spon_labels)) n_distinct(spon_labels$`_id`) else NA_integer_
spon_queue_rows <- if (!is.null(spon_queue)) nrow(spon_queue) else NA_integer_

spon_wf <- tibble(
  step = c("Raw trial-sponsor pairs", "Unique raw strings", "Accepted/review strings", "Trial label rows", "Labelled trials"),
  n = c(spon_raw_pairs, spon_unique_raw, spon_accepted_strings, spon_label_rows, spon_label_trials),
  type = c("start", "step", "step", "step", "end")
) |>
  mutate(delta = lag(n) - n)
```

```{r sponsor-kpis, results='asis'}
if (nrow(spon_unique) > 0) {
  cat(kpi_row(
    kpi_box(fmt_n(spon_raw_pairs), "raw trial-sponsor pairs", BLUE),
    kpi_box(fmt_n(spon_unique_raw), "unique raw strings", FROST),
    kpi_box(fmt_n(spon_label_trials), "trials with sponsor label", GREEN),
    kpi_box(fmt_n(spon_queue_rows), "review queue rows", ORANGE),
    kpi_box(ifelse(spon_has_rich_log, "rich", "compact"), "log schema", PURPLE)
  ))
} else {
  cat(notice("Sponsor log unavailable", "No sponsor normalisation log was found for this render.", "warn"))
}
```

```{r sponsor-waterfall, fig.height=4.3}
if (nrow(spon_unique) > 0 && all(!is.na(spon_wf$n))) {
  waterfall_plot(
    spon_wf,
    "Sponsor normalisation waterfall",
    "Counts move from raw trial-sponsor pairs to unique strings and final per-trial labels."
  )
}
```

```{r sponsor-waterfall-table}
if (nrow(spon_unique) > 0) {
  spon_wf |>
    transmute(Step = step, Count = fmt_n(n), `Removed / collapsed from previous step` = ifelse(is.na(delta), "NA", fmt_n(delta))) |>
    nk(cap = "Sponsor normalisation stage counts")
}
```

```{r sponsor-status-source-charts, fig.height=4.2}
if (nrow(spon_unique) > 0) {
  p_status <- spon_unique |>
    count(match_status, wt = n_trials, name = "trial_weighted_n") |>
    mutate(match_status = factor(match_status, levels = c("accepted", "review", "rejected", "unknown"))) |>
    ggplot(aes(match_status, trial_weighted_n, fill = match_status)) +
    geom_col(width = 0.62, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), vjust = -0.35, size = 3.1, color = DARK) +
    scale_fill_manual(values = c(accepted = GREEN, review = YELLOW, rejected = ORANGE, unknown = RED), na.value = GRAY) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    labs(title = "Sponsor outcomes weighted by trial count", x = NULL, y = "Trial occurrences") +
    theme_pre()

  print(p_status)
}
```

```{r sponsor-source-chart, fig.height=3.8}
if (nrow(spon_unique) > 0 && "match_source" %in% names(spon_unique)) {
  spon_unique |>
    mutate(match_source = coalesce(as.character(match_source), "not recorded")) |>
    count(match_source, wt = n_trials, name = "trial_weighted_n", sort = TRUE) |>
    slice_head(n = 12) |>
    mutate(match_source = reorder(match_source, trial_weighted_n)) |>
    ggplot(aes(trial_weighted_n, match_source)) +
    geom_col(fill = BLUE, width = 0.62) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), hjust = -0.1, size = 3.1, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    labs(title = "Sponsor normalisation source", x = "Trial occurrences", y = NULL) +
    theme_pre()
}
```

```{r sponsor-score-chart, fig.height=3.6}
if (nrow(spon_unique) > 0 && any(!is.na(spon_unique$match_score))) {
  spon_unique |>
    filter(!is.na(match_score)) |>
    ggplot(aes(match_score, weight = n_trials, fill = match_status)) +
    geom_histogram(binwidth = 5, boundary = 0, color = "white") +
    scale_fill_manual(values = c(accepted = GREEN, review = YELLOW, rejected = ORANGE, unknown = RED), na.value = GRAY) +
    scale_y_continuous(labels = comma) +
    labs(title = "Sponsor match-score distribution", x = "Match score", y = "Trial occurrences") +
    theme_pre()
} else {
  cat("Sponsor match scores are not available in the compact log schema.")
}
```

```{r sponsor-changes-table}
if (nrow(spon_unique) > 0) {
  spon_unique |>
    filter(!is.na(sponsor_clean), raw_sponsor != sponsor_clean | changed %in% TRUE) |>
    arrange(desc(n_trials)) |>
    select(
      `Raw sponsor` = raw_sponsor,
      `Normalised sponsor` = sponsor_clean,
      Status = match_status,
      Source = match_source,
      `N trials` = n_trials
    ) |>
    slice_head(n = 50) |>
    dt_tbl(caption = "Largest sponsor-name changes by trial count", page = 25, filter = "top")
}
```

```{r sponsor-top-chart, fig.height=5.5}
if ("sponsor_label" %in% names(df)) {
  df |>
    filter(!is.na(sponsor_label), sponsor_label != "") |>
    count(sponsor_label, register, name = "n_trials") |>
    group_by(sponsor_label) |>
    mutate(total = sum(n_trials)) |>
    ungroup() |>
    slice_max(total, n = 40, with_ties = FALSE) |>
    mutate(sponsor_label = reorder(sponsor_label, total)) |>
    ggplot(aes(n_trials, sponsor_label, fill = register)) +
    geom_col() +
    scale_fill_manual(values = REG_COLS) +
    scale_x_continuous(labels = comma) +
    labs(title = "Top sponsors by trial count", x = "Trials", y = NULL) +
    theme_pre()
}
```

```{r sponsor-label-detail}
if (!is.null(spon_labels)) {
  spon_labels |>
    count(sponsor_clean, sponsor_parent, sponsor_group, sponsor_type, match_status, name = "label_rows", sort = TRUE) |>
    slice_head(n = 100) |>
    dt_tbl(caption = "Sponsor labels written to trial_sponsor_labels.csv", page = 25, filter = "top")
}
```

```{r sponsor-queue-table}
if (!is.null(spon_queue)) {
  spon_queue |>
    arrange(desc(n_trials)) |>
    select(any_of(c("raw_sponsor", "candidate_sponsor", "sponsor_type", "match_status", "match_score", "match_source", "match_reason", "n_trials", "decision", "canonical_sponsor", "comment"))) |>
    dt_tbl(caption = "Sponsor review queue", page = 25, filter = "top")
} else {
  cat("Sponsor review queue not available.")
}
```

```{r sponsor-log-table}
if (nrow(spon_unique) > 0) {
  spon_unique |>
    arrange(match_status, desc(n_trials)) |>
    select(
      `Raw sponsor` = raw_sponsor,
      `Clean label` = sponsor_clean,
      Status = match_status,
      Score = match_score,
      Source = match_source,
      Reason = match_reason,
      `N trials` = n_trials
    ) |>
    dt_tbl(caption = "Full sponsor normalisation log, collapsed to unique raw strings", page = 25, filter = "top")
} else {
  cat("Sponsor log not available.")
}
```

## Substance Names

The substance audit follows the cache-build pipeline: exported trial-substance pairs, pre-filtered/normalised exploratory substances, accepted/review labels, per-trial `substance_label`, and optional review queue rows.

```{r substance-audit-compute}
sub_unique <- if (!is.null(sub_log) && has_cols(sub_log, c("raw_substance", "active_substance_clean", "match_status"))) {
  sub_log |>
    group_by(raw_substance) |>
    summarise(
      active_substance_clean = first(na.omit(active_substance_clean)),
      match_status = first(match_status),
      match_score = suppressWarnings(max(as.numeric(match_score), na.rm = TRUE)),
      match_source = first(na.omit(match_source)),
      match_reason = first(na.omit(match_reason)),
      trial_occurrences = n_distinct(`_id`),
      log_rows = n(),
      .groups = "drop"
    ) |>
    mutate(match_score = ifelse(is.infinite(match_score), NA_real_, match_score))
} else {
  tibble()
}

sub_raw_pairs <- if (!is.null(sub_raw) && has_cols(sub_raw, c("_id", "raw_substance"))) nrow(sub_raw) else NA_integer_
sub_raw_unique <- if (!is.null(sub_raw) && "raw_substance" %in% names(sub_raw)) n_distinct(sub_raw$raw_substance, na.rm = TRUE) else NA_integer_
sub_log_rows <- if (!is.null(sub_log)) nrow(sub_log) else NA_integer_
sub_unique_raw <- if (nrow(sub_unique) > 0) n_distinct(sub_unique$raw_substance) else NA_integer_
sub_accepted_strings <- if (nrow(sub_unique) > 0) sum(sub_unique$match_status %in% c("accepted", "review"), na.rm = TRUE) else NA_integer_
sub_label_trials <- if (!is.null(sub_labels) && "_id" %in% names(sub_labels)) n_distinct(sub_labels$`_id`) else NA_integer_
sub_queue_rows <- if (!is.null(sub_queue)) nrow(sub_queue) else NA_integer_

sub_wf <- tibble(
  step = c("Raw trial-substance pairs", "Unique raw strings", "Exploratory log rows", "Accepted/review strings", "Labelled trials"),
  n = c(sub_raw_pairs, sub_raw_unique, sub_log_rows, sub_accepted_strings, sub_label_trials),
  type = c("start", "step", "step", "step", "end")
) |>
  mutate(delta = lag(n) - n)

active_by_trial <- if (!is.null(sub_log) && has_cols(sub_log, c("_id", "active_substance_clean", "match_status"))) {
  sub_log |>
    filter(match_status %in% c("accepted", "review"), !is.na(active_substance_clean), active_substance_clean != "") |>
    distinct(`_id`, active_substance_clean) |>
    left_join(df |> select(`_id`, register, year, age_group), by = "_id")
} else {
  tibble()
}
```

```{r substance-kpis, results='asis'}
if (nrow(sub_unique) > 0) {
  cat(kpi_row(
    kpi_box(fmt_n(sub_raw_pairs), "raw trial-substance pairs", BLUE),
    kpi_box(fmt_n(sub_unique_raw), "unique logged raw strings", FROST),
    kpi_box(fmt_n(sub_label_trials), "trials with substance label", GREEN),
    kpi_box(fmt_n(sub_queue_rows), "review queue rows", ORANGE),
    kpi_box(fmt_n(n_distinct(sub_log$`_id`)), "trials in substance log", PURPLE)
  ))
} else {
  cat(notice("Substance log unavailable", "No substance normalisation log was found for this render.", "warn"))
}
```

```{r substance-waterfall, fig.height=4.3}
if (nrow(sub_unique) > 0 && all(!is.na(sub_wf$n))) {
  waterfall_plot(
    sub_wf,
    "Substance normalisation waterfall",
    "Counts move from raw trial-substance pairs to exploratory normalised log rows and labelled trials."
  )
}
```

```{r substance-waterfall-table}
if (nrow(sub_unique) > 0) {
  sub_wf |>
    transmute(Step = step, Count = fmt_n(n), `Removed / collapsed from previous step` = ifelse(is.na(delta), "NA", fmt_n(delta))) |>
    nk(cap = "Substance normalisation stage counts")
}
```

```{r substance-status-chart, fig.height=3.5}
if (nrow(sub_unique) > 0) {
  sub_unique |>
    count(match_status, wt = trial_occurrences, name = "trial_weighted_n") |>
    mutate(match_status = factor(match_status, levels = c("accepted", "review", "rejected", "unknown"))) |>
    ggplot(aes(match_status, trial_weighted_n, fill = match_status)) +
    geom_col(width = 0.62, show.legend = FALSE) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), vjust = -0.35, size = 3.1, color = DARK) +
    scale_fill_manual(values = c(accepted = GREEN, review = YELLOW, rejected = ORANGE, unknown = RED), na.value = GRAY) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    labs(title = "Substance outcomes weighted by trial count", x = NULL, y = "Trial occurrences") +
    theme_pre()
}
```

```{r substance-source-chart, fig.height=4}
if (nrow(sub_unique) > 0) {
  sub_unique |>
    mutate(match_source = coalesce(as.character(match_source), "not recorded")) |>
    count(match_source, wt = trial_occurrences, name = "trial_weighted_n", sort = TRUE) |>
    slice_head(n = 14) |>
    mutate(match_source = reorder(match_source, trial_weighted_n)) |>
    ggplot(aes(trial_weighted_n, match_source)) +
    geom_col(fill = BLUE, width = 0.62) +
    geom_text(aes(label = fmt_n(trial_weighted_n)), hjust = -0.1, size = 3.1, color = GRAY) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.16))) +
    labs(title = "Substance normalisation source", x = "Trial occurrences", y = NULL) +
    theme_pre()
}
```

```{r substance-score-chart, fig.height=3.6}
if (nrow(sub_unique) > 0 && any(!is.na(sub_unique$match_score))) {
  sub_unique |>
    filter(!is.na(match_score)) |>
    ggplot(aes(match_score, weight = trial_occurrences, fill = match_status)) +
    geom_histogram(binwidth = 5, boundary = 0, color = "white") +
    scale_fill_manual(values = c(accepted = GREEN, review = YELLOW, rejected = ORANGE, unknown = RED), na.value = GRAY) +
    scale_y_continuous(labels = comma) +
    labs(title = "Substance match-score distribution", x = "Match score", y = "Trial occurrences") +
    theme_pre()
}
```

```{r substance-top-chart, fig.height=5.5}
if (nrow(active_by_trial) > 0) {
  active_by_trial |>
    count(active_substance_clean, register, name = "n_trials") |>
    group_by(active_substance_clean) |>
    mutate(total = sum(n_trials)) |>
    ungroup() |>
    slice_max(total, n = 40, with_ties = FALSE) |>
    mutate(active_substance_clean = reorder(active_substance_clean, total)) |>
    ggplot(aes(n_trials, active_substance_clean, fill = register)) +
    geom_col() +
    scale_fill_manual(values = REG_COLS) +
    scale_x_continuous(labels = comma) +
    labs(title = "Top active substances by trial count", subtitle = "Accepted/review substance matches only.", x = "Trials", y = NULL) +
    theme_pre()
}
```

```{r substance-register-age, fig.height=4.8}
if (nrow(active_by_trial) > 0) {
  top_subs <- active_by_trial |>
    count(active_substance_clean, name = "total") |>
    slice_max(total, n = 12, with_ties = FALSE) |>
    pull(active_substance_clean)

  active_by_trial |>
    filter(active_substance_clean %in% top_subs) |>
    mutate(age_group = coalesce(age_group, "Unknown")) |>
    count(active_substance_clean, age_group, name = "n_trials") |>
    group_by(active_substance_clean) |>
    mutate(pct = n_trials / sum(n_trials)) |>
    ungroup() |>
    mutate(active_substance_clean = factor(active_substance_clean, levels = rev(top_subs))) |>
    ggplot(aes(pct, active_substance_clean, fill = age_group)) +
    geom_col() +
    scale_x_continuous(labels = percent) +
    scale_fill_manual(values = c(Paediatric = BLUE, Adult = ORANGE, `Paediatric & Adult` = GREEN, Unknown = GRAY), na.value = GRAY) +
    labs(title = "Age-group mix for top substances", x = "% of trials", y = NULL) +
    theme_pre()
}
```

```{r substance-queue-table}
if (!is.null(sub_queue)) {
  sub_queue |>
    arrange(desc(n_occurrences)) |>
    select(any_of(c("raw_substance", "active_substance_clean", "match_status", "match_score", "match_source", "match_reason", "n_occurrences", "decision", "canonical_substance", "comment"))) |>
    dt_tbl(caption = "Substance review queue", page = 25, filter = "top")
} else {
  cat("Substance review queue not available.")
}
```

```{r substance-log-table}
if (!is.null(sub_log) && has_cols(sub_log, c("raw_substance", "active_substance_clean", "match_status"))) {
  sub_log |>
    arrange(match_status, desc(match_score)) |>
    mutate(match_score = if ("match_score" %in% names(sub_log)) round(as.numeric(match_score), 2) else NA_real_) |>
    select(
      `Trial ID` = `_id`,
      `Raw substance` = raw_substance,
      `Clean label` = active_substance_clean,
      Status = match_status,
      Score = match_score,
      Source = match_source,
      Reason = match_reason
    ) |>
    dt_tbl(caption = "Full substance normalisation log", page = 25, filter = "top")
} else {
  cat("Substance log not available.")
}
```

# 5. PIP, Results, and Orphan Flags

```{r boolean-flag-charts, fig.height=4}
flag_cols <- intersect(c("has_PIP", "has_results", "is_orphan"), names(df))
if (length(flag_cols) > 0) {
  flag_df <- df |>
    select(register, all_of(flag_cols)) |>
    pivot_longer(-register, names_to = "flag", values_to = "value", values_transform = list(value = as.character)) |>
    mutate(value = coalesce(as.character(value), "Unknown"))

  ggplot(flag_df |> count(flag, register, value, name = "n"), aes(register, n, fill = value)) +
    geom_col(width = 0.62) +
    facet_wrap(~ flag, scales = "free_y") +
    scale_y_continuous(labels = comma) +
    labs(title = "PIP, results, and orphan-derived flags", x = NULL, y = "Trials") +
    theme_pre()
}
```

```{r flag-tables}
if (length(flag_cols) > 0) {
  df |>
    select(register, all_of(flag_cols)) |>
    pivot_longer(-register, names_to = "Flag", values_to = "Value", values_transform = list(Value = as.character)) |>
    mutate(Value = coalesce(as.character(Value), "Unknown")) |>
    count(Flag, register, Value, name = "Trials") |>
    group_by(Flag, register) |>
    mutate(`% within register` = fmt_pct(Trials / sum(Trials) * 100), Trials = fmt_n(Trials)) |>
    ungroup() |>
    dt_tbl(caption = "Flag distributions by register", page = 30)
}
```

# 6. Completeness

```{r completeness-table}
comp_df <- tibble(
  Column = names(df),
  Type = vapply(df, function(x) class(x)[1], character(1)),
  `Non-missing` = vapply(df, function(x) sum(!(is.na(x) | (is.character(x) & x == ""))), integer(1)),
  `Complete (%)` = round(`Non-missing` / nrow(df) * 100, 1)
) |>
  arrange(`Complete (%)`)

dt_tbl(comp_df, caption = paste0("Completeness of all ", ncol(df), " cache columns"), page = 25, filter = "top")
```

```{r missingness-heatmap, fig.height=8, fig.width=10}
heat_cols <- intersect(c(
  "CT_number", "Full_title", "MEDDRA_term", "MEDDRA_organ_class", "Member_state",
  "sponsor_label", "sponsor_type", "phase", "status", "has_PIP", "has_results",
  "is_orphan", "submission_date_parsed", "decision_date", "start_date",
  "days_to_decision", "year", "DIMP_product_name", "substance_label"
), names(df))

if (length(heat_cols) > 0) {
  miss_df <- df |>
    select(register, all_of(heat_cols)) |>
    group_by(register) |>
    summarise(across(all_of(heat_cols), ~ mean(is.na(.x) | (is.character(.x) & .x == "")) * 100), .groups = "drop") |>
    pivot_longer(-register, names_to = "column", values_to = "pct_missing")

  ggplot(miss_df, aes(register, column, fill = pct_missing)) +
    geom_tile(color = "white", linewidth = 0.45) +
    geom_text(aes(label = sprintf("%.0f%%", pct_missing)), size = 2.8, color = DARK) +
    scale_fill_gradient2(low = GREEN, mid = YELLOW, high = RED, midpoint = 50, name = "% missing") +
    labs(title = "Missingness by register", x = NULL, y = NULL) +
    theme_pre() +
    theme(legend.position = "right", axis.text.y = element_text(size = 8.5))
}
```

# 7. Data Quality Issues

```{r issue-compute, include=FALSE}
count_blank <- function(x) sum(is.na(x) | (is.character(x) & x == ""), na.rm = TRUE)

unmapped_status_codes <- if (!is.null(log_status_c) && has_cols(log_status_c, c("register", "raw", "normalised", "n_trials"))) {
  log_status_c |>
    filter(register == "CTIS", !is.na(raw), str_detect(as.character(raw), "^[0-9]+$"), !(as.character(raw) %in% as.character(1:6))) |>
    arrange(desc(n_trials))
} else tibble()
n_ctis_unmapped_status <- if (nrow(unmapped_status_codes) > 0) sum(unmapped_status_codes$n_trials, na.rm = TRUE) else 0

country_fail_examples <- if (!is.null(log_country) && has_cols(log_country, c("raw", "normalised", "n_trials"))) {
  log_country |>
    filter(!is.na(raw), raw != "", is.na(normalised) | normalised == "") |>
    arrange(desc(n_trials)) |>
    slice_head(n = 30)
} else tibble()
n_country_fail <- if (nrow(country_fail_examples) > 0) sum(country_fail_examples$n_trials, na.rm = TRUE) else 0

organ_unresolved_examples <- if (!is.null(log_organ) && has_cols(log_organ, c("raw", "normalised", "n_trials"))) {
  log_organ |>
    filter(!is.na(normalised), str_detect(as.character(normalised), "^[0-9]+$")) |>
    arrange(desc(n_trials)) |>
    slice_head(n = 30)
} else tibble()
n_organ_unresolved <- if (nrow(organ_unresolved_examples) > 0) sum(organ_unresolved_examples$n_trials, na.rm = TRUE) else 0

ct_dupes_examples <- df |>
  filter(!is.na(canonical_trial_id), canonical_trial_id != "") |>
  add_count(canonical_trial_id) |>
  filter(n > 1) |>
  arrange(canonical_trial_id) |>
  select(any_of(c("_id", "register", "canonical_trial_id", "CT_number", "transition_eudract_number", "sponsor_label", "status")))
n_ct_dupes <- n_distinct(ct_dupes_examples$canonical_trial_id)

junk_pat <- regex("\\bGmbH\\b|\\bInc\\.?\\b|\\bLtd\\.?\\b|\\bS\\.A\\.\\b|\\bAG\\b|\\bB\\.V\\.\\b|\\bLLC\\b|\\bPLC\\b", ignore_case = FALSE)
sponsor_junk_ex <- df |>
  filter(!is.na(sponsor_label), str_detect(sponsor_label, junk_pat)) |>
  count(sponsor_label, sort = TRUE) |>
  slice_head(n = 30)
n_sponsor_junk <- sum(!is.na(df$sponsor_label) & str_detect(df$sponsor_label, junk_pat))

days_extreme_examples <- df |>
  filter(!is.na(days_to_decision), days_to_decision > 3650) |>
  arrange(desc(days_to_decision)) |>
  select(any_of(c("_id", "register", "sponsor_label", "submission_date_parsed", "decision_date", "days_to_decision"))) |>
  slice_head(n = 30)
n_days_extreme <- nrow(days_extreme_examples)

year_implaus_ex <- df |>
  filter(!is.na(year), year < 1990 | year > year(Sys.Date())) |>
  select(any_of(c("_id", "register", "submission_date", "submission_date_parsed", "year", "sponsor_label"))) |>
  slice_head(n = 30)
n_year_implaus <- nrow(year_implaus_ex)

issue_tbl <- tibble(
  Severity = c("High", "High", "High", "High", "High", "Medium", "Medium", "Medium", "Medium", "Low", "Low", "Low"),
  Category = c("Deduplication", "Deduplication", "Deduplication", "Status", "Dates", "Country", "Organ class", "Identifiers", "Sponsor", "Status", "Dates", "Dates"),
  Issue = c(
    "Residual cross-register duplicates by shared title key",
    "Transitioned EUCTR trials without a CTIS match in current fallbacks",
    "Duplicate canonical trial IDs in final cache",
    "CTIS numeric status codes outside the current mapping",
    "Completed trials with no decision_date",
    "Country strings normalised to missing",
    "CTIS organ-class codes still stored as raw numbers",
    "Canonical trial IDs appearing more than once",
    "Sponsor labels still contain legal-entity tokens",
    "Trials with status 'Other'",
    "days_to_decision greater than 10 years",
    "Submission year outside [1990, current year]"
  ),
  `N affected` = c(
    fmt_n(n_resid_dupes),
    fmt_n(n_unmatched_transitioned),
    fmt_n(n_ct_dupes),
    fmt_n(n_ctis_unmapped_status),
    fmt_n(sum(df$status == "Completed" & is.na(df$decision_date), na.rm = TRUE)),
    fmt_n(n_country_fail),
    fmt_n(n_organ_unresolved),
    fmt_n(n_ct_dupes),
    fmt_n(n_sponsor_junk),
    fmt_n(sum(df$status == "Other", na.rm = TRUE)),
    fmt_n(n_days_extreme),
    fmt_n(n_year_implaus)
  ),
  Suggested_action = c(
    "Review title-only matches and require corroborating identifiers before automated collapse.",
    "Refresh CTIS data and inspect missing transition EudraCT identifiers.",
    "Review canonical_trial_id derivation for transition aliases and latest-version preference.",
    "Extend CTIS status-code mapping if these codes are valid source values.",
    "Add or audit decision-date fallbacks for completed trials.",
    "Add unresolved country strings to the country lookup or preserve raw values for review.",
    "Extend the MedDRA SOC lookup for unresolved numeric CTIS codes.",
    "Review duplicate canonical identities and remove duplicate real-world trials.",
    "Add sponsor aliases for fallback labels that still include legal suffixes.",
    "Review raw status values and extend status category patterns where warranted.",
    "Treat implausible lag values as missing or audit source dates.",
    "Set implausible parsed years to missing after date parsing."
  )
)
```

## Summary

```{r issues-summary}
issue_tbl |>
  dt_tbl(caption = "Data-quality issues detected during preprocessing audit", page = 20, filter = "top") |>
  DT::formatStyle(
    "Severity",
    backgroundColor = DT::styleEqual(c("High", "Medium", "Low"), c("#FFEBEB", "#FFF3E0", "#FFFDE7"))
  )
```

## Detail Tables

```{r issue-detail-tabs, results='asis'}
cat("The tables below show examples for issues where example rows are available.")
```

```{r issue-residual-dupes}
if (nrow(cross_reg_tbl) > 0) {
  dt_tbl(cross_reg_tbl, caption = "Residual cross-register duplicate candidates", page = 25)
} else {
  cat("No residual cross-register duplicate candidates.")
}
```

```{r issue-unmapped-status}
if (nrow(unmapped_status_codes) > 0) {
  unmapped_status_codes |>
    select(Register = register, `Raw code` = raw, Normalised = normalised, `N trials` = n_trials) |>
    dt_tbl(caption = "Unmapped CTIS status codes", page = 20)
} else {
  cat("No unmapped CTIS status codes were detected.")
}
```

```{r issue-country-fail}
if (nrow(country_fail_examples) > 0) {
  country_fail_examples |>
    select(Register = register, `Raw country` = raw, `N trials` = n_trials) |>
    dt_tbl(caption = "Country strings normalised to missing", page = 20)
} else {
  cat("No country strings normalised to missing.")
}
```

```{r issue-organ-unresolved}
if (nrow(organ_unresolved_examples) > 0) {
  organ_unresolved_examples |>
    select(Register = register, Raw = raw, Normalised = normalised, `N trials` = n_trials) |>
    dt_tbl(caption = "Unresolved numeric organ-class values", page = 20)
} else {
  cat("No unresolved numeric organ-class values.")
}
```

```{r issue-canonical-dupes}
if (nrow(ct_dupes_examples) > 0) {
  dt_tbl(ct_dupes_examples, caption = "Duplicate canonical trial IDs", page = 25)
} else {
  cat("No duplicate canonical trial IDs.")
}
```

```{r issue-sponsor-junk}
if (nrow(sponsor_junk_ex) > 0) {
  sponsor_junk_ex |>
    rename(`Sponsor label` = sponsor_label, `N trials` = n) |>
    dt_tbl(caption = "Sponsor labels with legal suffix tokens", page = 20)
} else {
  cat("No sponsor labels with legal suffix tokens.")
}
```

```{r issue-days-extreme}
if (nrow(days_extreme_examples) > 0) {
  days_extreme_examples |>
    mutate(days_to_decision = fmt_n(days_to_decision)) |>
    dt_tbl(caption = "Extreme days_to_decision values", page = 20)
} else {
  cat("No days_to_decision values above 10 years.")
}
```

```{r issue-year-implaus}
if (nrow(year_implaus_ex) > 0) {
  dt_tbl(year_implaus_ex, caption = "Implausible submission years", page = 20)
} else {
  cat("No implausible submission years.")
}
```

---

*Generated by `rmarkdown/preprocessing.Rmd` at `r format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")`.*
